### 2021-06-07 Draft/Supplementary_Material/Traits_Table.Rmd
EDITS

---
title: ""
output:
  pdf_document: default
  html_document: default
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=12, fig.width=12}
setwd("~/R_Projects/Reproductive Traits")

library(kableExtra)

library(dplyr)
library(readxl) #read excel file (trait data)
library(dplyr) #data manipulation
library(visdat) #VISUALIZE MISSING DATA
library(naniar)
library(tidyverse)
########################################################################################################################################################
#1) READ TRAIT DATA
########################################################################################################################################################

Trait <- c("Plant height (m)", "Flower width (mm)", "Flower length (mm)", "Inflorescence width (mm)", "Style length (mm)", "Ovules per flower",
           "Flowers per plant",  "Autonomous selfing (fruit set)")

Trait1 <- c("Lifepan", "Life form", "Flower shape", "Flower symmetry", "Autonomous selfing", "Compatibility system", "Breeding system", "")

T.cat <- linebreak(c("Short-lived \n Perennial", "Herb \n Shrub \n Tree", "Brush \n Campanulate \n Capitulum \n Open \n Papilionaceous \n Tube",  "Actinomorphic \n Zygomorphic", "None \n Low \n Medium \n High", "Self-incomp. \n Partially self-comp. \n Self-comp.", "Hermaphrodite \n Monoecious \n Dioecious",""),align = c("l"))

Type <- c("Vegetative", "Floral", "Floral", "Floral", "Floral", "Floral", "Floral","Reproductive")

Type_1 <- c("Vegetative", "Vegetative", "Floral", "Floral", "Reproductive", "Reproductive", "Reproductive", "")

dat <- data.frame(Type, Trait,Type_1, Trait1, T.cat)

colnames(dat) <- c("Type","Traits","Type", "Traits", "Categories")

kbl(dat, booktabs = T, linesep = "\\addlinespace", escape = FALSE, align = c("c","l","c","c","l","l","c")) %>%
kable_styling(full_width = T) %>%
add_header_above(c("Quantitative traits" = 2,"Categorical traits" = 3),bold=T)  %>%
    column_spec(c(1,4), bold = T) %>% row_spec(0, bold=T)

```