### 2021-06-15 Draft/trial.Rmd
edits on heatmap

---
title: "Untitled"
output:
  html_document: default
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=13, fig.cap="Figure S2. Hierarchical clustering dendrogram with the branches coloured by the optimal number of clusters (5). The labels of the subgroup of species (N = 524) used in this study are coloured in black in order to show the evenness of the distribution of the species across clusters. The rest of species labels are omitted for visualization purposes (N = 982)."}

library(ggplot2)
library(dendextend)
library(dplyr)

e.clust_5 <- readRDS("../Data/RData/e.clust_5.rds")

dendro <- as.dendrogram(e.clust_5)
dendro.col <- dendro %>%
  set("branches_k_color", k = 5, value =   c("lightsteelblue2", "thistle", "sandybrown", "#7BB0A3", "lightgoldenrod2")) %>%
  set("branches_lwd", 0.7) %>%
  set("labels_colors", 
      value = c("darkslategray")) %>% 
  set("labels_cex", 0.065)

labels_colors(dendro.col) <- get_leaves_branches_col(dendro.col)

# We use "darkgrey" to color some species
final_d_1 <- read.csv("../Data/Csv/data_for_motifs_analysis_1.csv")

plant_species <- unique(final_d_1$Plant_species)

index_species_out <- which(names(labels_colors(dendro.col)) %in% plant_species)

'%ni%' <- Negate('%in%')

index_species_in <- which(names(labels_colors(dendro.col)) %ni% plant_species)

labels_colors(dendro.col)[index_species_out] <- "black"
labels_colors(dendro.col)[index_species_in] <- "white"

ggd1 <- as.ggdend(dendro.col)

ggd1$labels$angle <- 90
ggd1$labels$vjust <- 1
ggd1$labels$hjust <- 1

ggplot(ggd1,offset_labels=-0.05) + theme(panel.grid.major = element_blank(),
 axis.text = element_blank(),axis.title = element_blank()) +
  ggtitle("Plant functional groups")+
  theme(plot.title = element_text(hjust = 0.5, vjust = -2,size = 20,face="bold"))+
  geom_text(aes(x = x, y = y, label = label, angle = -90, hjust = 0), data= label(ddata))

```