### 2021-08-05 Draft/Draft.Rmd
fig 5 added thinking of headings for methods and results...

---
title: Plant-pollintor networks worldwide are composed by the same specific building
  blocks
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
output:
  html_document: default
  pdf_document: default
  word_document: default
always_allow_html: true
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Abstract

Ecological processes leave distinct structural imprints on species interactions shaping the topology of mutualistic networks. Detecting those relationships is not trivial since they go beyond pair-wise interactions, but may get blurred when considering full network descriptors. However, recent work has shown the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of interactions between a small number of species (i.e. motifs) and those constitute the building blocks of the whole network. Here, we have compiled 60 networks from 18 different studies and show that some motifs are consitently over-represented worldwide, suggesting that the building blocks of plant-pollintor networks are not random and are associated to ... Second, we show that the position of pollinator guilds and plant reproductive strategies is not random with respect to the positions occupied within each motif. ... Hence, we show that species ecology is shaping the building blocks that conform the web of life.

##INTRODUCTION

The interaction between plants and pollinators can be studied at different scales, from species level interactions (micro-scale) to the full network structure (macro-scale). Research of the plant-pollinator network structure have proven common invariant structural properties across them including a degree distribution that decays as a power law [@jordano1987], nestedness [@bascompte2003], or modularity in large networks [@olesen2007]. In addition, species phenological overlap, morphological matching and species abundances have been shown to be determinant for the understanding of pairwise plant-pollinator interactions [@bartomeus2016; @stang2006; @peralta2020]. However, both species level information and the holistic view of the full network involves missing relevant information for the understanding of ecological processes [@cirtwill2018; @simmons2019].

Traditionally, plant-pollinator research has focused on direct interactions and overlooked indirect interactions such as facilitative or competitive interactions between plants for pollinators [@moeller2004; @sargent2008]. Despite the widespread nature of indirect interactions in ecological communities [@strauss1991], plant-pollinator research often fails to finely capture those indirect interactions with the conventional analytical tools that condense the information either by species (e.g. interaction frequency) or in single topological indices (e.g. nestedness). Nonetheless, the emerging framework of network motifs in plant-pollinator research, the building blocks of a network that depict subsets of interactions [@milo2002], allows to consider both direct and indirect interactions [@simmons2019]. The analysis of motifs (meso-scale) in plant-pollinator networks have revealed that the different ecological processes that govern species interactions (e.g., species abundances versus trait-matching) can lead to different patterns of indirect interactions [@simmons2020]. Yet, the global patterns of indirect interactions in plant-pollinator networks (over- and under- representation of motifs) are still unknown.

Motifs are abstract representations often decoupled from species ecology. Thus, linking the structural properties of the meso-scale with the species ecology can help the understanding of ecological processes. For instance, different motifs can have different ecological meanings [@simmons2019] and the position within a motif can determine the species functional role [@stouffer2012; @baker2015]. However, it is unclear how the species ecology and life history traits determines the species functional role within the network of interactions [@coux2016]. For example, large pollinators can forage larger distances [@greenleaf2007], deposit greater pollen quantities [@foldesi2021] and handle complex zygomorphic flowers in comparison with small pollinators that are restricted to lower floral complexity [@gong2009]. How this different pollinator behaviors translate into their interaction topology is unknown. Similarly, recent empirical findings indicate that the meso-scale is the best descriptor of plant reproductive success [@allen2021], but little is known on how plants reproductive strategies shape their position within the network of interactions. Although some studies have evaluated plant reproductive strategies in plant-pollinator networks [@tur2013; @lazaro2020], they are often overlooked in a community context [@devaux2014] and rarely incorporated into plant-pollinator network studies. Hence, exploring how the main plant reproductive strategies integrate with the emergent motif framework can shed light on key aspects of ecosystem functioning.

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside a detailed grouping of plants into reproductive strategies and pollinator species into functional groups. Plants were grouped based on a comprehensive dataset that included floral, reproductive and vegetative traits compiled in Lanuza et al., (unpublished) on a larger set of plant-pollinator networks. Pollinators were grouped into the main taxonomical groups that differed in life form and behaviour. Then, we explored which motifs up to five nodes in these set of plant-pollinator networks were over and under-represented. Finally, we explored if the different plants and pollinators functional groups were over- or under-represented in certain positions.

##METHODS

**Plant-pollinator studies**

We have compiled 60 plant-pollinator networks from 18 different studies (Table S1). All studies sampled plant-pollinator interactions in natural systems and were selected based on wide geographical coverage and the presence of interaction frequency as a measure of interaction strength. In total, there were 503 plant species, 1,111 pollinator species and 6248 of pairwise interactions registered. For ease of data manipulation plant and pollinator species names were standardize with the help of the package _taxize_ version 0.9.99 [@chamberlain2020].

**Plant and pollinator functional groups**

First, plant species were grouped into the optimal number of functional groups that summarized the main plant reproductive strategies. This was done with the help of hierarchical cluster analysis by using the trait dataset collated in Lanuza et al., (unpublished) that comprised 1,506 plant species and contained the same 60 plant-pollinator networks used in this study plus some non-weighted and weighted metawebs (see Table S1 Lanuza et al., unpublished). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits (Table S2). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting the functional groups with that many variables @dolnicar2014. For this, we calculated the distance of the different functional traits with the function _gowdis_ from the package _FD_ version 1.0-12 [@laliberte2014] with the method ward.D2 used for non-squared distances [@murtagh2014]. All the numerical traits were previously scaled. Finally, we applied a hierarchical cluster analysis with the function _hclust_ from the _R_ stats package version 4.0.5 and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version 1.4-7 [@white2009].

Second, pollinators were grouped into functional groups based on taxonomic rank. We opted to divide pollinators on the taxonomic rank level and not with functional traits because (i) the main taxonomic orders differ in form and behaviour and (ii) the lower complexity of higher taxonomic ranks (i.e., pollinators had 6 orders versus plants that had 38). Hence, we grouped pollinators into 6 functional groups: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi) Coleoptera.

**Meso-scale analysis**

Following Simmons et al. [@simmons2019; @simmons2020], we broke down the plant-pollinator networks into their constituent motifs. Prior to analyses, we turned the quantitative networks into qualitative (or binary) ones, where interactions are present or absent.

We calculated the frequency of all motifs up to five nodes (see Figure 1) for each empirical network, by using the _bmotif_ package [@bmotifsimmons2019]. To control for variation in network size and for the fact that smaller motifs can be nested within larger motifs, the frequencies were normalised as a proportion of the total number of motifs within each motif class (i.e., the number of nodes a motif contains). Like [@simmons2020], we just used five-node motifs in our analyses for visualisation, interpretation and computational reasons. In addition, we also excluded two-node motifs (or links) from our analyses because their normalised frequencies would always equal one.

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.cap="Figure 1. Adapted figure of Simmons et al., 2019 with all the possible motifs from two to five species in bipartite networks. There is a total of 17 possible motifs with 46 different positions denoted within each node.", fig.width=10,fig.height=8}
library(tidyverse)
library(RColorBrewer)
source("../Scripts/Scripts_Alfonso/plot_motif_positionsV2.R")

motif_number <- 1
p1 <- plot_motif_positions(motif_number)

motif_number <- 2
p2 <- plot_motif_positions(motif_number)

motif_number <- 3
p3 <- plot_motif_positions(motif_number)

motif_number <- 4
p4 <- plot_motif_positions(motif_number)

motif_number <- 5
p5 <- plot_motif_positions(motif_number)

motif_number <- 6
p6 <- plot_motif_positions(motif_number)

motif_number <- 7
p7 <- plot_motif_positions(motif_number)

motif_number <- 8   
p8 <- plot_motif_positions(motif_number)

motif_number <- 9
p9 <- plot_motif_positions(motif_number)

motif_number <- 10
p10 <- plot_motif_positions(motif_number)

motif_number <- 11
p11 <- plot_motif_positions(motif_number)

motif_number <- 12
p12 <- plot_motif_positions(motif_number)

motif_number <- 13
p13 <- plot_motif_positions(motif_number)

motif_number <- 14
p14 <- plot_motif_positions(motif_number)

motif_number <- 15
p15 <- plot_motif_positions(motif_number)

motif_number <- 16
p16 <- plot_motif_positions(motif_number)

motif_number <- 17
p17 <- plot_motif_positions(motif_number)

library(patchwork)

(p1 & ylab("2 species")| p2 & ylab("3 species") | p3 )/ (p4 & ylab("4 species")| p5 | p6 | p7)/ (p8  & ylab("5 species")| p9 | p10 | p11 | p12) / (p13 & ylab("5 species")| p14 | p15 | p16 | p17) & theme(axis.title.y = element_text(color="black", size=14, face="bold")) #+
 #plot_annotation(tag_levels = '1')

```

To assess the significance of the observed frequencies, we created 1,000 simulated networks for each binary network using the _nullmodel_ function and the _vaznull_ model in the _bipartite_ package [@dormann2009]. Generated networks had the same number of plants and pollinators, as well as the same connectance of their corresponding empirical networks. After extracting the motif frequencies from the simulated networks, for each motif type and empirical network, we calculated the percentage of simulated networks whose frequencies were smaller than the ones observed, that is, we estimated the percentile of the observed motif frequencies. Motifs whose percentile is close to zero or 100 are under- or over-represented in the empirical networks, respectively, and, thus, they cannot be predicted by connectance and the number of species alone. To summarize general patterns across networks, we used an intercept-only linear mixed model (LMM) per motif, where the response variable was the observed motif percentile per network. In these models, we used the study identifiers in Table S1 as a random intercept. By doing so, we obtained estimates of the average motif frequency, in which we controlled the variation at the study level.

Next, we calculated which functional groups were over or under-represented in different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. We extracted the position frequencies of all motifs from three to five nodes for each network and species by using the _bmotif_ package [@bmotifsimmons2019]. To estimate the position frequencies of each functional group in a given network, we added the frequencies of those species that belong to the group, and then, we normalised the resulting frequencies by dividing the position measure for each group by the total number of times that a group appears in any position within the same motif size class. Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did motif frequencies. To outline the general patterns of position frequencies across networks and functional groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the functional group identifier as an explanatory variable and the study identifiers as a random intercept. By adding the functional group estimates to the model intercept, we assessed the average motif frequency, after controlling the variation at the study level. Finally, we visualized with the help of the package ComplexHeatmap version 2.6.2 [@gu2016] over- and under- representation of plant and pollinator functional groups on the different motif positions.

To test the effect of singletones on the frequencies of motif and those of functional groups' positions, in Appendices XXX, we applied the previous analyzes to the robust versions of the networks in table S1, namely: networks that, prior to binarization, only contained interactions whose frequency was greater than one (64.98% of interactions).

Finally, we studied which motif combinations of functional groups (up to five nodes) are over or under-represented in 95% of our empirical networks (57 out of 60 arrangements, due to computational limitations to identify all the nodes in the motifs of the three networks with the highest number of links). To do so, for each of the 53,250 possible motif combinations, we estimated the observed and the expected probability of finding that combination in empirical networks, respectively. Then, we determined whether the observed probabilities are likely to come from the expected probabilities or not. To calculate the observed probability of a given combination $i$, $p_i^O$, we divided the number of times that $i$ appears in our set of empirical networks ($n_i^O$) by the sum of the number of times that each possible combination appeared: $p_i^O = n_i^O / \sum_{k=1}^{53,250} n_k^O$, and, consequently, $\sum_{k=1}^{53,250} p_k^O = 1$. To estimate the expected probability of a given motif combination $i$, $p_i^E$, firstly, we calculated the probability of finding a given functional group $x$ in the position $\alpha$ of $i$, $p_i(x,\alpha)$. Then, by assuming the independency of $p_i(x,\alpha)$, we computed the expected probability of the combination $i$ as the product of the probability of its pairs $(x,\alpha)$, that is, $p_i^E = \prod _{(x,\alpha )\in i} p_i(x,\alpha)$. To obtain $p_i(x,\alpha)$, we proceeded as follows. First, we used the _bmotif_ package [@bmotifsimmons2019] to calculate the number of times (or absolute frequency) that the functional group $x$ appears in the position $\alpha$ in each empirical network $\eta$, denoted as $n_i^{\eta}(x,\alpha)$. Then, by controlling the variation at the network level, we assessed the average value of the absolute frequency of functional group $x$ appears in the position $\alpha$, $n_i(x,\alpha) = E\left [ n_i^{\eta}(x,\alpha) \right ]$, and estimated $p_i(x,\alpha)$ as $p_i(x,\alpha) = n_i(x,\alpha)/\sum_{k}n_i(k,\alpha)$. To obtain $n_i(x,\alpha)$, we fitted a LMM per motif position, where the response variable was the number of times that a given position was observed per network, the explanatory variable was functional group identifier, and the random intercept was given by network identifiers nested within the study identifiers. We used the ‘lmer’ package [@bates2015] to fit the LMMs models in our analyses.

Once we obtained $p_i^O$ and $p_i^E$, we used simulation to determine whether the former is likely to come from the latter or not, since the large number of possible motif combinations and the small probabilities for some of them advise against using an Exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contained 10 million elements and, for each combination, the probability of being selected was equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. Following the usual interpretation for z-scores, combinations with $z_i^O > 1.96$ are over-represented, whereas those with $z_i^O < -1.96$ are under-represented, at the 95%  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).

##RESULTS

**Functional groups**

The hierarchical cluster analysis divided the dataset with 1506 species and 15 traits into five different clusters with different and overlapping characteristics (Figure S1 and Figure S2). The subset of plant species used in this study (N = 503) were distributed evenly across these different larger five groups (see subset of species labels within cluster in Figure S1). The first cluster was dominated by herbs with hermaphrodite flowers with high levels of autonomous selfing and we we refer to this group as "selfing herbs". The second cluster was characterized by small perennial species with a mixed of life forms (trees shrubs and herbs) with outcrossing hermaphroditic flowers and we refer to this group of species as "small outcrossing perennials". The third cluster was dominated by also perennial species with a mixed of life forms too and had large self-incompatible hermaphroditic flowers with high number of ovules and we named this group "self-incompatible perennials with large flowers". The fourth cluster had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding system with small flowers but the highest number of flowers per plant and inflorescence and we refer to this set of species as "tall plants with small unisexual flowers". Finally, the last cluster of species was dominated by small perennial and shot-lived herbs with long self-compatible zygomorphic flowers unable to self-pollinate and we refer to this group as "short-lived outcrossers with long zygomorphic flowers".

**Overall meso-scale patterns**

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.cap="Figure 2. Comparison of the motif frequencies between empirical and simulated networks. Average percentages of motifs close to 0 and 100 indicate under- and over- representation in empirical networks. The different motifs are coloured by the mean path length as done in @simmons2020.", fig.width=9,fig.height=5}
library(lme4)
library(nlme)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models/Motifs_frequency_percentile.csv")
d <- add_study_id(d)
#str(d)
#head(d)

#ggplot(d %>% filter(motif != 1), aes(percentil_sizeclass))+
#  geom_histogram(color="black", fill="white")+
#  facet_wrap(~motif)+
#  theme_bw()+
#  labs(x="Sizeclass percentile")
#
# Motifs tend to be over- and under-represensented in real networks

# Mean percentile and SE taking into account the study system
motif_codes <- d %>% filter(motif != 1) %>% # We remove links (motif code: 1)
  select(motif) %>% unique() %>% pull()
motif_means <- tibble(motif = motif_codes)
motif_means$mean <- NA
motif_means$SE <- NA

for(i.motif in 1:length(motif_codes)){

  m <- lmer(percentil_sizeclass ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(d, motif == motif_codes[i.motif]))

  motif_means$mean[i.motif] <- fixed.effects(m) %>% unname()
  motif_means$SE[i.motif] <- sqrt(diag(vcov(m)))

}

#Add categories based on path length set by Simmons 2021 et al., Functional Ecology
d$Broad_categories[d$motif==2 | d$motif==7 | d$motif==17 | d$motif==3|d$motif==4|d$motif==8] <- "Fan"

d$Broad_categories[d$motif==15 | d$motif==11] <- "Medium-weak"

d$Broad_categories[d$motif==6 | d$motif==16 |  d$motif==12] <- "Strong"

d$Broad_categories[d$motif==5 | d$motif==9 |  d$motif==14|  d$motif==13|  d$motif==10] <- "Weak"

#order the categories from short to long path length
d$Broad_categories <- factor(d$Broad_categories, levels = c("Strong", "Fan", "Medium-weak", "Weak"))

ggplot(NULL) + 
  geom_point(data = d %>% filter(motif != 1),
             aes(y=as.factor(motif), x=100*percentil_sizeclass, 
                 color = as.factor(Broad_categories)),
             position = "jitter",alpha=0.5)+
  geom_errorbar(data = motif_means,aes(y = as.factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=1.0,size=1)+
  geom_point(data = motif_means,aes(y = as.factor(motif), x=100*mean), 
                size=2)+
  labs(y="Motif", x = "Percentile (%)",colour="Path length classification")+
  theme_bw() + scale_color_manual(labels = c('Strong' = expression(paste("Complete (", bar(x), " = 1.38)")),'Fan' = expression(paste("Fan (", bar(x), " = 1.48)")), 'Medium-weak' = expression(paste("Asymmetric complete (", bar(x), " = 1.60)")), "Weak" =expression(paste("Core-peripheral (", bar(x), " = 1.85)"))), name = "Path length classification",values = c("grey25","springgreen4","sienna2","goldenrod2"),guide = guide_legend(title.position = "top",nrow = 2)) +
theme(legend.text.align = 0)+
  theme(legend.position="bottom")

#Trying to maintain a gradient from low to high in colors

#Fig 1 Should be a dot plot (or forest plot) of the mean +- SE of the 15 motifs.
#NICE! Can we put motifs in Y axes, and % in X?
```

**Meso-scale functional groups position**

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE,fig.align = 'center',fig.width=6,fig.height=3.5}
library(lme4)
library(nlme)
library(tidyverse)
library(stringi)
library(ggpubr)
source("../Scripts/Scripts_Alfonso/add_study_id.R")

all_position_percentiles <- read_csv("../Data/Csv/Motifs_positions_and_null_models/GF_positions_frequency_percentile.csv")

all_position_percentiles$position <- stri_extract_first_regex(all_position_percentiles$position,
                                                              "[0-9]+") %>% as.numeric()

all_position_percentiles <- add_study_id(all_position_percentiles)
#str(all_position_percentiles)
#head(all_position_percentiles)

list_plants_FG <- as.character(1:10)

plant_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF)) #ME DA ERROR
pollinator_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(!Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF))

#ggplot(plant_position_percentiles_filtered, aes(percentil_its_GF))+
#  geom_histogram(color="black", fill="white")+
#  facet_wrap(~position)+
#  theme_bw()+
#  labs(x="Sizeclass percentile", title = "Histograms: Percentiles for plant positions")
#
#ggplot(pollinator_position_percentiles_filtered, aes(percentil_its_GF))+
#  geom_histogram(color="black", fill="white")+
#  facet_wrap(~position)+
#  theme_bw()+
#  labs(x="Sizeclass percentile",title = "Histograms: Percentiles for pollinator positions")
#

# estimated mean taking into account the random str is 0.XX +- 0.XX (look at the (Intercept) estimate)

# Plants-------------

# Mean percentile and SE taking into account the study system
plant_codes <- plant_position_percentiles_filtered$position %>% unique()
plant_means <- plant_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
plant_means$mean <- NA
plant_means$SE <- NA

for(i.pos in 1:length(plant_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id),
            data = plant_position_percentiles_filtered %>%
                          filter(position == plant_codes[i.pos]))

  temp <- fixed.effects(m) %>% unname()
  plant_means$mean[plant_means$position == plant_codes[i.pos]] <- c(temp[1],
                                                                     temp[1]+temp[2],
                                                                     temp[1]+temp[3],
                                                                     temp[1]+temp[4],
                                                                     temp[1]+temp[5]) ## to get real means per FG
  temp <- sqrt(diag(vcov(m)))
  plant_means$SE[plant_means$position == plant_codes[i.pos]] <- c(temp[1], 
                                                                  temp[1]+temp[2],
                                                                  temp[1]+temp[3],
                                                                  temp[1]+temp[4],
                                                                  temp[1]+temp[5])

}

GF_plant_pos <- ggplot(NULL) + 
  geom_point(data = plant_position_percentiles_filtered,
             aes(x=as.factor(position), y=percentil_its_GF, 
                 color = as.factor(position)),
             position = "jitter",alpha=0.5)+
  geom_errorbar(data = plant_means,aes(x = as.factor(position), ymin=mean-SE, ymax=mean+SE), 
                width=1.0,size=1)+
  facet_wrap(~Node_FG,ncol = 5)+
  geom_point(data = plant_means,aes(x = as.factor(position), y=mean), 
             size=2)+
  labs(x="Position", y = "Percentile",title = "Plants")+
  theme_bw()+
  theme(legend.position = "none")+#,axis.text.x=element_text(angle=90,vjust=0.5, hjust=1))+
  coord_flip()

# Pollinators--------

# Mean percentile and SE taking into account the study system
pollinator_codes <- pollinator_position_percentiles_filtered$position %>% unique()
pollinator_means <- pollinator_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
pollinator_means$mean <- NA
pollinator_means$SE <- NA
pollinator_means_reordered <- NULL

for(i.pos in 1:length(pollinator_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = pollinator_position_percentiles_filtered %>%
                          filter(position == pollinator_codes[i.pos]),
            control = lmerControl(optimizer ="Nelder_Mead")) # I changed the optimizer because motif 8 (i.pos = 3) caused convergence problems, max gradient = 0.0022 > 0.002.

  temp <- fixed.effects(m) %>% unname()
  pollinator_means_aux <- pollinator_means[pollinator_means$position == pollinator_codes[i.pos],] %>% arrange(Node_FG)

  pollinator_means_aux$mean <- c(temp[1],
                                 temp[1]+temp[2],
                                 temp[1]+temp[3],
                                 temp[1]+temp[4],
                                 temp[1]+temp[5],
                                 temp[1]+temp[6],
                                 temp[1]+temp[7],
                                 temp[1]+temp[8],
                                 temp[1]+temp[9]) ## to get real means per FG

  temp <- sqrt(diag(vcov(m)))
  pollinator_means_aux$SE <- c(temp[1],
                               temp[1]+temp[2],
                               temp[1]+temp[3],
                               temp[1]+temp[4],
                               temp[1]+temp[5],
                               temp[1]+temp[6],
                               temp[1]+temp[7],
                               temp[1]+temp[8],
                               temp[1]+temp[9])

  pollinator_means_reordered <- bind_rows(pollinator_means_reordered,pollinator_means_aux)

}

# There are lower limits (mean - SE) smaller than 0 (see Other insects)
# We set the lower limit of those error bars to zero

pollinator_means_reordered <- pollinator_means_reordered %>% mutate(lower = mean-SE)
pollinator_means_reordered$lower[pollinator_means_reordered$lower < 0] <- 0

GF_poll_pos <- ggplot(NULL) + 
  geom_point(data = pollinator_position_percentiles_filtered,
             aes(x=as.factor(position), y=percentil_its_GF, 
                 color = as.factor(position)),
             position = "jitter",alpha=0.5)+
  geom_errorbar(data = pollinator_means_reordered, aes(x = as.factor(position), ymin=lower,
                                                       ymax=mean+SE), 
                width=1.0,size=1)+
  facet_wrap(~Node_FG,ncol = 5)+
  geom_point(data = pollinator_means_reordered,aes(x = as.factor(position), y=mean), 
             size=2)+
  labs(x="Position", y = "Percentile", title = "Pollinators")+
  theme_bw()+
  theme(legend.position = "none")+#,axis.text.x=element_text(angle=90,vjust=0.5, hjust=1))+
  coord_flip()

# Plant and Pollinators--------
#
#ggarrange(GF_plant_pos,GF_poll_pos,
#          ncol = 1, nrow = 2,heights = c(.90, 1.4))
#
#Fig 2 may be the same, but for the each positions-FG combination? Maybe heatmap (but we lose the SE's)

```

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=8,fig.width=8,fig.align = 'center', fig.cap="Figure 3. Heatmap indicating under- and over- representation of pollinator and plant functional groups in the different motif positions. The different motif positions are dividied by the average path length clasification by @simmons2020."}

library(lme4)
library(nlme)
library(tidyverse)
library(stringi)
source("../Scripts/Scripts_Alfonso/add_study_id.R")

#############################################
# EXTRACT MEAN VALUES
#############################################

all_position_percentiles <- read_csv("../Data/Csv/Motifs_positions_and_null_models/GF_positions_frequency_percentile.csv")

all_position_percentiles$position <- stri_extract_first_regex(all_position_percentiles$position,
                                                              "[0-9]+") %>% as.numeric()

all_position_percentiles <- add_study_id(all_position_percentiles)
#str(all_position_percentiles)
#head(all_position_percentiles)

list_plants_FG <- as.character(1:10)

plant_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF)) #ME DA ERROR
pollinator_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(!Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF))

# Plants-------------

# Mean percentile and SE taking into account the study system
plant_codes <- plant_position_percentiles_filtered$position %>% unique()
plant_means <- plant_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
plant_means$mean <- NA
plant_means$SE <- NA

for(i.pos in 1:length(plant_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id),
            data = plant_position_percentiles_filtered %>%
              filter(position == plant_codes[i.pos]))

  temp <- fixed.effects(m) %>% unname()
  plant_means$mean[plant_means$position == plant_codes[i.pos]] <- c(temp[1],
                                                                    temp[1]+temp[2],
                                                                    temp[1]+temp[3],
                                                                    temp[1]+temp[4],
                                                                    temp[1]+temp[5]) ## to get real means per FG
  temp <- sqrt(diag(vcov(m)))
  plant_means$SE[plant_means$position == plant_codes[i.pos]] <- c(temp[1], 
                                                                  temp[1]+temp[2],
                                                                  temp[1]+temp[3],
                                                                  temp[1]+temp[4],
                                                                  temp[1]+temp[5])

}

# Pollinators--------

# Mean percentile and SE taking into account the study system
pollinator_codes <- pollinator_position_percentiles_filtered$position %>% unique()
pollinator_means <- pollinator_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
pollinator_means$mean <- NA
pollinator_means$SE <- NA
pollinator_means_reordered <- NULL

for(i.pos in 1:length(pollinator_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = pollinator_position_percentiles_filtered %>%
              filter(position == pollinator_codes[i.pos]),
            control = lmerControl(optimizer ="Nelder_Mead")) # I changed the optimizer because motif 8 (i.pos = 3) caused convergence problems, max gradient = 0.0022 > 0.002.

  temp <- fixed.effects(m) %>% unname()
  pollinator_means_aux <- pollinator_means[pollinator_means$position == pollinator_codes[i.pos],] %>% arrange(Node_FG)

  pollinator_means_aux$mean <- c(temp[1],
                                 temp[1]+temp[2],
                                 temp[1]+temp[3],
                                 temp[1]+temp[4],
                                 temp[1]+temp[5],
                                 temp[1]+temp[6],
                                 temp[1]+temp[7],
                                 temp[1]+temp[8],
                                 temp[1]+temp[9]) ## to get real means per FG

  temp <- sqrt(diag(vcov(m)))
  pollinator_means_aux$SE <- c(temp[1],
                               temp[1]+temp[2],
                               temp[1]+temp[3],
                               temp[1]+temp[4],
                               temp[1]+temp[5],
                               temp[1]+temp[6],
                               temp[1]+temp[7],
                               temp[1]+temp[8],
                               temp[1]+temp[9])

  pollinator_means_reordered <- bind_rows(pollinator_means_reordered,pollinator_means_aux)

}

# There are lower limits (mean - SE) smaller than 0 (see Other insects)
# We set the lower limit of those error bars to zero

pollinator_means_reordered <- pollinator_means_reordered %>% mutate(lower = mean-SE)
pollinator_means_reordered$lower[pollinator_means_reordered$lower < 0] <- 0

#############################################
# ADDING BROAD MOTIF CATEGORIES TO POSITIONS
#############################################

motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

motifs_raw_data$Broad_categories <- NA

motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(5,9,14,13,10)] <- "Core-peripherical"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(6,16,12)] <- "Complete"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(2,3,7,4,17,8)] <- "Fan"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(15,11)] <- "Asymmetric complete"

motifs_raw_data$plant <- gsub("[^0-9.-]", "", motifs_raw_data$plant)
motifs_raw_data$pollinator <- gsub("[^0-9.-]", "", motifs_raw_data$pollinator)
motifs_raw_data$plant <- as.numeric(motifs_raw_data$plant)
motifs_raw_data$pollinator <- as.numeric(motifs_raw_data$pollinator)
motifs_raw_data <- motifs_raw_data %>% filter(!is.na(Broad_categories))

fan_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

pollinator_means_reordered$Broad_categories <- NA

pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% fan_pollinator] <- "Fan" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% asymm_pollinator] <- "Medium-weak" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% complete_pollinator] <- "Strong" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% core_per_pollinator] <- "Weak"

fan_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

plant_means$Broad_categories <- NA

plant_means$Broad_categories[plant_means$position %in% fan_plant] <- "Fan" 
plant_means$Broad_categories[plant_means$position %in% asymm_plant] <- "Medium-weak" 
plant_means$Broad_categories[plant_means$position %in% complete_plant] <- "Strong" 
plant_means$Broad_categories[plant_means$position %in% core_per_plant] <- "Weak"

########################################
# CREATE HEATMAPS
########################################

library(ComplexHeatmap)

# PLANTS-----

# Prepare data for heatmap
heat_plants_aux <- plant_means %>% spread(Node_FG,mean) %>% select(-SE)

heat_plants <- heat_plants_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_plants_fan <- heat_plants %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_fan) <- heat_plants_fan[,1]

heat_plants_Weak <- heat_plants %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Weak) <- heat_plants_Weak[,1]

heat_plants_Strong <- heat_plants %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Strong) <- heat_plants_Strong[,1]

heat_plants_Medium_weak <- heat_plants %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Medium_weak) <- heat_plants_Medium_weak[,1]

# POLLINATORS-----------

filter_rare_groups <- T

if(filter_rare_groups == T){
  heat_pollinators_aux <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds",
                                                                                "Lizards",
                                                                                "Other_insects")) %>% 
    spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee",            
                                      "Coleoptera","Lepidoptera","Non-bee\nHymenoptera", 
                                      "Non-syrphids\ndiptera","Syrphids")
}else{
  heat_pollinators_aux <- pollinator_means_reordered %>% spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee","Birds",            
                                      "Coleoptera","Lepidoptera","Lizards","Non-bee\nHymenoptera", 
                                      "Non-syrphids\ndiptera","Other insects","Syrphids")
}

heat_pollinators <- heat_pollinators_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_pollinators_fan <- heat_pollinators %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_fan) <- heat_pollinators_fan[,1]

heat_pollinators_Weak <- heat_pollinators %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Weak) <- heat_pollinators_Weak[,1]

heat_pollinators_Strong <- heat_pollinators %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Strong) <- heat_pollinators_Strong[,1]

heat_pollinators_Medium_weak <- heat_pollinators %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Medium_weak) <- heat_pollinators_Medium_weak[,1]

##################

#Do not know how to handle this multiplot for creating a panel

# I'm going to create the 4 plots in one, 
# I'm not sure that the grouping of the functional group is correct
# when is done separately

#sorry for the mess of libraries that I do below

#Prepare first pollinators

library(tidyr)
library(data.table)
long <- melt(setDT(heat_pollinators), id.vars = c("position","Broad_categories"), variable.name = "guilds")

long_1 <- long %>% select(-Broad_categories)

library(reshape2)
long_m <- acast(long_1, position~guilds, value.var="value")

#set manually the divisions, so far this is the best way I have seen to split the rows by categories

r_split <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 4),
             rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 3),rep(c("Asymmetric complete"), 2),
             "Complete", "Fan")

#set order of labels that avoids overlapping
r_split <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

#plat a bit with colors, a bit of a more complex set but looks nicer
library(circlize)
col_fun = colorRamp2(c(0.05644663,0.25, 0.4673256,0.75, 0.8782047), c("blue","cadetblue3", "azure2", "gold","red"))

h_pol <- Heatmap(long_m, name = "Mean\npercentile", row_split=r_split,cluster_row_slices = FALSE, 
                 column_title = "Pollinator functional groups",row_names_gp = gpar(fontsize = 10),
                 col = col_fun,column_names_rot = 45,  column_names_gp = grid::gpar(fontsize = 10),
                 column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                 row_title_gp = gpar(fontsize = 10, fontface = "bold"))

#Prepare plants

#long format
long_plants <- melt(setDT(heat_plants), id.vars = c("position","Broad_categories"), variable.name = "Functional_groups")

#filter for creating matrix
long_plants_1 <- long_plants %>% select(-Broad_categories)

#set better colnames
long_plants_1$Functional_groups <- as.character(long_plants_1$Functional_groups)

long_plants_1$Functional_groups[long_plants_1$Functional_groups==1] <- "Selfing herbs" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==2] <- "Small outcrossing\nperennials" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==3] <- "Self-incomp. perennials\nwith large flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==4] <- "Tall plants with small\nunisexual flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==5] <- "Short-lived outcrossers with\nlong zygomorphic flowers"

long_plants_1$Functional_groups <- factor(long_plants_1$Functional_groups, levels=c("Selfing herbs",
"Small outcrossing\nperennials" ,"Self-incomp. perennials\nwith large flowers",
"Tall plants with small\nunisexual flowers" ,"Short-lived outcrossers with\nlong zygomorphic flowers" ))

#create matrix with reshape
long_plants_m <- acast(long_plants_1, position~Functional_groups, value.var="value")

r_split_plants <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 3),
                    rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 4),rep(c("Asymmetric complete"), 2),
                    "Complete", "Fan")

r_split_plants <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

h_plan <- Heatmap(long_plants_m, name = "Mean\npercentile", row_split=r_split_plants,cluster_row_slices = F, 
                  column_title = "Plant functional groups",row_names_gp = gpar(fontsize = 10),col = col_fun,column_names_rot = 45,  
                  column_names_gp = grid::gpar(fontsize = 10),column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                  row_title_gp = gpar(fontsize = 10, fontface = "bold"))

library(patchwork)
ht_list <- h_pol+h_plan

draw(ht_list, auto_adjust = FALSE, cluster_rows = F)

```

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=8, fig.width=12, fig.cap="Figure 4. Graphical representation of the probability of finding a given functional group $x$ in the position $\\alpha$ of motif $i$, $p_i(x,\\alpha)$, for all the possible motifs from two to five species in bipartite networks. The slices in the nodes for a given functional group $x$ are proportional to the corresponding value of $p_i(x,\\alpha)$."}

plant_means_reordered <- read_csv("../Data/Csv/plant_abs_freq_means.csv")
pollinator_means_reordered <- read_csv("../Data/Csv/pollinator_abs_freq_means.csv")

library(ggplot2)
library(scatterpie)
library(RColorBrewer)

# Pollinator--------
pollinator_pies <- pollinator_means_reordered %>% select(-mean, -SE) %>% 
  spread(Node_FG,mean_natural_units)

pollinator_pies$x <- 0
pollinator_pies$y <- 0

#ggplot() + geom_scatterpie(aes(x=x, y=y, group=position,r=5), 
#                           data = pollinator_pies,
#                           cols=as.character(c("Bee","Birds","Coleoptera","Lepidoptera",
#                                               "Lizards","Non-bee-Hymenoptera",
#                                               "Non-syrphids-diptera","Other_insects",
#                                               "Syrphids")))+
#  coord_equal()+
#  scale_fill_brewer(palette = "Paired")+
#  facet_wrap(~position,nrow = 3)+
#  theme_bw()+
#    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#          panel.background = element_blank(),
#          axis.title.x=element_blank(),axis.text.x=element_blank(),
#          axis.ticks.x=element_blank(),axis.title.y=element_blank(),
#          axis.text.y=element_blank(),axis.ticks.y=element_blank(),
#          legend.position="bottom")+
#  labs(title = "Pollinator positions",fill=NULL)
#    
# plant--------
plant_pies <- plant_means_reordered %>% select(-mean, -SE) %>% 
  spread(Node_FG,mean_natural_units)

plant_pies$x <- 0
plant_pies$y <- 0

#ggplot() + geom_scatterpie(aes(x=x, y=y, group=position,r=5), 
#                           data = plant_pies,
#                           cols=as.character(c("1","2","3","4","5")))+
#  coord_equal()+
#  scale_fill_brewer(palette = "Dark2")+
#  facet_wrap(~position,nrow = 3)+
#  theme_bw()+
#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
#        axis.title.x=element_blank(),axis.text.x=element_blank(),
#        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
#        axis.text.y=element_blank(),axis.ticks.y=element_blank(),
#        legend.position="bottom")+
#  labs(title = "Plant positions",fill=NULL)
#

#################################
# CREATE PIE-MOTIFS
#################################

equidistribution <- function(number_nodes){

  x_width <- 5.5/(number_nodes+1)

  x_result <- rep(-1.5,number_nodes)

  for(i in 1:number_nodes){

    x_result[i] <- x_result[i] + i*x_width

  }

  return(x_result)

}

plot_pie_motif_positions <- function(motif_number){
  motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

  motif_data <- motifs_raw_data %>% filter(motif_id == motif_number)

  # Create dataframe with node positions

  labdown = motif_data$plant %>% unique()

  xdown = equidistribution(length(labdown))

  ydown = rep(0,length(labdown))

  labup = motif_data$pollinator %>% unique()

  xup = equidistribution(length(labup))

  yup = rep(2,length(labup))

  d_nodes=data.frame(x=c(xdown,xup),
                     y=c(ydown,yup),
                     label = c(labdown,labup))

  # Create a dataframe with links

  plant_positions <- data.frame(plant = labdown,
                                xdown = xdown,
                                ydown = ydown)

  pollinator_positions <- data.frame(pollinator = labup,
                                     xup = xup,
                                     yup = yup)

  link_df <- motif_data %>% left_join(plant_positions, by = "plant") %>% 
    left_join(pollinator_positions, by = "pollinator")

  # Before plotting the motif, we remove letters from the node position id

  d_nodes$label <- gsub("[^0-9.-]", "", d_nodes$label)
  plant_pies_aux <- plant_pies %>% select(-x,-y) %>% mutate(position = as.character(position))

  d_nodes_plant <- d_nodes %>% rename(position = label) %>% 
    left_join(plant_pies_aux, by = "position") %>% as_tibble() %>%
    filter(!is.na(`1`))

  pollinator_pies_aux <- pollinator_pies %>% select(-x,-y) %>% mutate(position = as.character(position))

  d_nodes_pollinator <- d_nodes %>% rename(position = label) %>% 
    left_join(pollinator_pies_aux, by = "position") %>% as_tibble() %>%
    filter(!is.na(Bee))

  # Define the number of colors you want
  nb.cols <- 5
  mycolors <- c(brewer.pal(9, "Paired"),brewer.pal(5, "Dark2"))

  ggplot() +
    scale_shape_identity() +
    geom_segment(data = link_df, aes(x=xdown, y=ydown, xend = xup, yend = yup), 
                 size = 1, color = "grey20")+  geom_scatterpie(aes(x=x, y=y, group=position,r=.45), 
    data = d_nodes_pollinator, cols=as.character(c("Bee","Birds","Coleoptera","Lepidoptera",
     "Lizards","Non-bee-Hymenoptera", "Non-syrphids-diptera","Other_insects",
     "Syrphids"))) +
    geom_scatterpie(aes(x=x, y=y, group=position,r=0.45), 
                    data = d_nodes_plant,
                    cols=as.character(c("1","2","3","4","5")))+
    coord_fixed()+
    scale_fill_manual(values = mycolors,"test",
      breaks=c("Bee","Birds","Coleoptera","Lepidoptera",
               "Lizards","Non-bee-Hymenoptera", "Non-syrphids-diptera","Other_insects",
               "Syrphids","1","2","3","4","5"), labels=c("Bee","Birds","Coleoptera","Lepidoptera",
                                                        "Lizards","Non-bee-Hymenoptera", "Non-syrphids-diptera","Other_insects",
                                                        "Syrphids", 
          "Selfing herbs","Small outcrossing perennials","Self-incomp. perennials\nwith large flowers",
          "Tall plants with small\nunisexual flowers","Short-lived outcrossers with\nlong zygomorphic flowers"))+
    xlim(-1.5,4)+ ylim(-0.5,2.5)+
    theme(panel.background = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())+
    labs(x=NULL,y=NULL)+
    guides(color = FALSE,fill=guide_legend(title="Functional groups"))+
    ggtitle(paste0("Motif ",motif_number))+
    theme(plot.title = element_text(hjust = 0.5),legend.key.size = unit(0.5, 'cm'))+
    theme(plot.margin=unit(c(0,0,0,0),"cm"))
}

motif_number <- 1
p1 <- plot_pie_motif_positions(motif_number)

motif_number <- 2
p2 <- plot_pie_motif_positions(motif_number)

motif_number <- 3
p3 <- plot_pie_motif_positions(motif_number)

motif_number <- 4
p4 <- plot_pie_motif_positions(motif_number)

motif_number <- 5
p5 <- plot_pie_motif_positions(motif_number)

motif_number <- 6
p6 <- plot_pie_motif_positions(motif_number)

motif_number <- 7
p7 <- plot_pie_motif_positions(motif_number)

motif_number <- 8   
p8 <- plot_pie_motif_positions(motif_number)

motif_number <- 9
p9 <- plot_pie_motif_positions(motif_number)

motif_number <- 10
p10 <- plot_pie_motif_positions(motif_number)

motif_number <- 11
p11 <- plot_pie_motif_positions(motif_number)

motif_number <- 12
p12 <- plot_pie_motif_positions(motif_number)

motif_number <- 13
p13 <- plot_pie_motif_positions(motif_number)

motif_number <- 14
p14 <- plot_pie_motif_positions(motif_number)

motif_number <- 15
p15 <- plot_pie_motif_positions(motif_number)

motif_number <- 16
p16 <- plot_pie_motif_positions(motif_number)

motif_number <- 17
p17 <- plot_pie_motif_positions(motif_number)

library(patchwork)

(p1 & ylab("2 species")| p2 & ylab("3 species") | p3  |plot_spacer()|plot_spacer())/ 
  (p4 & ylab("4 species")| p5 | p6 | p7|plot_spacer())/ 
  (p8  & ylab("5 species")| p9 | p10 | p11 | p12) / 
  (p13 & ylab("5 species")| p14 | p15 | p16 | p17) + plot_layout(guides = 'collect') &
  theme(axis.title.y = element_text(color="black", size=14, face="bold"),legend.position = "bottom")

```

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=8, fig.width=12, fig.cap="Figure 5."}

library(tidyverse)
library(RColorBrewer)

equidistribution <- function(number_nodes){

  x_width <- 5.5/(number_nodes+1)

  x_result <- rep(-1.5,number_nodes)

  for(i in 1:number_nodes){

    x_result[i] <- x_result[i] + i*x_width

  }

  return(x_result)

}

plot_motif_label <- function(label){

  label_split <- str_split(label,"_")[[1]]
  motif_number <- as.numeric(label_split[1])
  nodes <- label_split[2:length(label_split)]
  plant_nodes <- nodes[nodes %in% plant_means_reordered$Node_FG]
  pollinator_nodes <- nodes[nodes %in% pollinator_means_reordered$Node_FG]

  motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

  motif_data <- motifs_raw_data %>% filter(motif_id == motif_number)

  # Create dataframe with node positions

  labdown = motif_data$plant %>% unique()

  xdown = equidistribution(length(labdown))

  ydown = rep(0,length(labdown))

  labup = motif_data$pollinator %>% unique()

  xup = equidistribution(length(labup))

  yup = rep(1,length(labup))

  d_nodes=data.frame(x=c(xdown,xup),
                     y=c(ydown,yup),
                     label = c(labdown,labup))

  # Create a dataframe with links

  plant_positions <- data.frame(plant = labdown,
                                xdown = xdown,
                                ydown = ydown)

  pollinator_positions <- data.frame(pollinator = labup,
                                     xup = xup,
                                     yup = yup)

  link_df <- motif_data %>% left_join(plant_positions, by = "plant") %>% 
    left_join(pollinator_positions, by = "pollinator")

  # Add FG IDs to nodes
  FG_IDs <- tibble(label = c(sort(labdown),sort(labup)),
                   FG = c(plant_nodes,pollinator_nodes))

  d_nodes_FG <- d_nodes %>% left_join(FG_IDs, by = "label")
  levels_FG <- c("Bee","Birds","Coleoptera","Lepidoptera",
                 "Lizards","Non-bee-Hymenoptera", "Non-syrphids-diptera","Other_insects",
                 "Syrphids","1","2","3","4","5")

  d_nodes_FG$FG <- factor(d_nodes_FG$FG, levels = levels_FG)

  levels(d_nodes_FG$FG) <- c("Bee","Birds","Coleoptera","Lepidoptera",
                             "Lizards","Non-bee-Hymenoptera", "Non-syrphids-diptera","Other_insects",
                             "Syrphids","1","2","3","4","5")

  # Before plotting the motif, we remove letters from the node position id

  d_nodes_FG$label <- gsub("[^0-9.-]", "", d_nodes_FG$label)

  # Define the number of colors you want
  mycolors <- c(brewer.pal(9, "Paired"),brewer.pal(5, "Dark2"))

  ggplot(data=d_nodes_FG, aes(x, y)) +
    scale_shape_identity() +
    geom_segment(data = link_df, aes(x=xdown, y=ydown, xend = xup, yend = yup), size = 1, color = "grey20")+
    geom_point(aes(color = FG), size=7.5)+
    geom_text(aes(x, y , label = label), size=3)+
    scale_color_manual(drop = FALSE, values = mycolors,
                       breaks=c("Bee","Birds","Coleoptera","Lepidoptera",
                                "Lizards","Non-bee-Hymenoptera", "Non-syrphids-diptera","Other_insects",
                                "Syrphids","1","2","3","4","5"), 
                       labels=c("Bee","Birds","Coleoptera","Lepidoptera",
                                "Lizards","Non-bee-Hymenoptera", "Non-syrphids-diptera",
                                "Other insects", "Syrphids", 
                                "Selfing herbs","Small outcrossing perennials",
                                "Self-incomp. perennials\nwith large flowers",
                                "Tall plants with small\nunisexual flowers",
                                "Short-lived outcrossers with\nlong zygomorphic flowers"))+
    xlim(-1.5,4)+ ylim(-0.5,1.5)+
    theme(panel.background = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())+
    labs(x=NULL,y=NULL)+
    ggtitle(paste0("Motif ",motif_number))+
    theme(plot.title = element_text(hjust = 0.5,size=10))+
    guides(color=guide_legend(title="Functional groups"))

}

# Data with pollinator/plant labels

plant_means_reordered <- read_csv("../Data/Csv/plant_abs_freq_means.csv")
pollinator_means_reordered <- read_csv("../Data/Csv/pollinator_abs_freq_means.csv")

####################################################################################################
#Read data to check top 5 motif of each category of observed propabilities

#Under, No signi, Over-rep
####################################################################################################
# Check top Z-scores of each group (Under, No diff, Over)

#Read data
data <- read.csv("../Data/Csv/motifs_observed_probability_SIMUL_CI.csv")

#Find critical value of Z-score
p <- 0.05 #cutoff probability 95% confidence
critical_value <- qnorm(p/2) #double tail probability divide by 2

data_1 <- data %>%
  mutate(infra_over_represented = case_when(
    z_score < -abs(critical_value) ~ "infra",
    between(z_score, -abs(critical_value), abs(critical_value)) ~ "no_diff",
    z_score > abs(critical_value) ~ "over"
  ))

#data_2 <- filter(data_1, abs(z_score)<2000) 
#str(data_2)

#Select under-represented by the ones with highest porbability of being observed (top10)
under_data <- filter(data_1, infra_over_represented=="infra")

under_top_5 <- filter(under_data, row_number(desc(abs(motif_observed_probability)))<= 5) %>%
  select(motif_functional_ID,motif_observed_probability,counts_observed,z_score)

#Select no statistical difference by the ones with highest porbability of being observed (top5)
no_diff_data <- filter(data_1, infra_over_represented=="no_diff")

no_diff_top_5 <- filter(no_diff_data, row_number(desc(abs(motif_observed_probability)))<= 5) %>%
  select(motif_functional_ID,motif_observed_probability,counts_observed,z_score)

#Select over-represenetd by the ones with highest porbability of being observed (top5)
over_data <- filter(data_1, infra_over_represented=="over")

over_top_5 <- filter(over_data, row_number(desc(abs(motif_observed_probability)))<= 5) %>%
  select(motif_functional_ID,motif_observed_probability,counts_observed,z_score)
####################################################################################################
#Now plot it 
####################################################################################################

library(patchwork)

#Prepare list of 5 plots for each group
#Under
labels_under <- under_top_5$motif_functional_ID
p_under <- list()

for (i in 1:5){

  p_under[[i]] <- plot_motif_label(labels_under[[i]])
}

#No signficance
labels_nodiff <- no_diff_top_5$motif_functional_ID
p_nodiff <- list()

for (i in 1:5){

  p_nodiff[[i]] <- plot_motif_label(labels_nodiff[[i]])
}

#Over
labels_over <- over_top_5$motif_functional_ID
p_over <- list()

for (i in 1:5){

  p_over[[i]] <- plot_motif_label(labels_over[[i]])
}

#Run Script with histogram plot
source("../Scripts/Z-score_percentile_histogram_plot.R")

hist <- ggplot(data_2 %>% filter(round_motif_observed_probability>0), aes(x=z_score, color=infra_over_represented, fill=infra_over_represented)) + 
  geom_histogram(bins = 93, alpha = 0.5, position = "identity",lwd = 0.25)+
  geom_vline(xintercept = -abs(critical_value))+
  geom_vline(xintercept = abs(critical_value))+
  xlim(-60,60) + ylab("Frequency")+  theme_bw() +
  scale_fill_manual(name="Motif frequencies" ,values=c("coral2", "palegreen3", "cyan3"), labels=c("Under-represented",
                                                                                                  "No statistical difference", "Over-represented")) +
  scale_color_manual(name="Motif frequencies" ,values=c("coral2", "palegreen3", "cyan3"), labels=c("Under-represented",
                                                                                                   "No statistical difference", "Over-represented")) + xlab("Z-score")

t_col <- function(color, percent = 50, name = NULL) {
  #      color = color name
  #    percent = % transparency
  #       name = an optional name for the color

  ## Get RGB values for named color
  rgb.val <- col2rgb(color)

  ## Make new color using input color as base and alpha set by transparency
  t.col <- rgb(rgb.val[1], rgb.val[2], rgb.val[3],
               max = 255,
               alpha = (100 - percent) * 255 / 100,
               names = name)

  ## Save the color
  invisible(t.col)
}
## END

#Prepare panel of plots
hist / ((p_under[[1]]/p_under[[2]]/p_under[[3]]/p_under[[4]] & theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill=t_col("coral2")))) | 
          (p_nodiff[[1]]/p_nodiff[[2]]/p_nodiff[[3]]/p_nodiff[[4]]& theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill=t_col("palegreen3")))) | 
          (p_over[[1]]/p_over[[2]]/p_over[[3]]/p_over[[4]] & theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill=t_col("cyan3"))))) + plot_layout(guides = 'collect')

```

##REFERENCES

<div id ="refs"></div>

##SUPPLEMENTARY MATERIAL

**TABLES**

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE}
##############################
#TABLE S1. List of studies
##############################

#Load libraries
library(kableExtra)
library(dplyr)

#Create data.frame with references

#vector with number of netw per study
`Number of networks` <- c("6", #1
                          "2", #2
                          "3",#3
                          "1",#4
                          "1",#5
                          "1",#6
                          "1",#7
                          "2",#8
                          "1",#9
                          "1",#10
                          "8",#11
                          "16",#12
                          "6",#13
                          "2",#14
                          "4",#15
                          "1",#16
                          "3",#17
                          "1"#18
                          )

`First author` <- c("Bartomeus", #1
                    "Dicks",#2
                    "Dupont",#3
                    "Elberling",#4
                    "Fang",#5
                    "Inouye",#6
                    "Lundgren",#7
                    "Olesen",#8
                    "Small",#9
                    "Souza",#10
                    "Kaiser-Bunbury",#11
                    "Bartomeus",#12
                    "Kaiser-Bunbury",#13
                    "Kaiser-Bunbury",#14
                    "Peralta",#15
                    "Burkle",#16
                    "Arroyo-Correa",#17
                    "Bundgaard"#18
                    )

Year <- c(2008,#1
          2002,#2
          2003,#3
          1999,#4
          2008,#5
          1988,#6
          2005,#7
          2002,#8
          1976,#9
          2017,#10
          2017,#11
          2015,#12
          2011,#13
          2010,#14
          2006,#15
          2013,#16
          2019,#17
          2003#18
          )

Country <- c("Spain", #1
               "England", #2
               "Denmark", #3
               "Sweden", #4
               "China", #5
               "United States", #6
               "Denmark (Greenland)", #7
               "Mauritius and Portugal (Azores)", #8
               "Japan", #9
               "Brazil", #10
               "Seychelles", #11
               "Spain", #12
               "Seychelles", #13
               "Mauritius", #14
               "Argentina", # 15
               "United States", #16
               "New Zealand", #17
               "Denmark" #18
               )

DOI <- c("https://doi.org/10.1007/s00442-007-0946-1", #1
         "https://doi.org/10.1046/j.0021-8790.2001.00572.x", #2
         "https://doi.org/10.1111/j.1365-2656.2008.01501.x", #3
         "https://doi.org/10.1111/j.1600-0587.1999.tb00507.x", #4
         "https://doi.org/10.1111/1749-4877.12190", #5
         "https://doi.org/10.1111/j.1442-9993.1988.tb00968.x", #6
         "https://doi.org/10.1657/1523-0430(2005)037[0514:TDAHCW]2.0.CO;2", #7
         "https://doi.org/10.1046/j.1472-4642.2002.00148.x", #8
         "/13960/t4km08d21", #9
         "https://doi.org/10.1111/1365-2745.12978", #10
         "https://doi.org/10.1038/nature21071", #11
         "https://github.com/ibartomeus/BeeFunData", #12
         "https://doi.org/10.1111/j.1365-2745.2010.01732.x", #13
         "https://doi.org/10.1016/j.ppees.2009.04.001", #14
         "https://doi.org/10.1111/ele.13510", #15
         "https://doi.org/10.1126/science.1232728", #16
         "https://doi.org/10.1111/1365-2745.13332", #17
         "Unpublished, Master thesis" #18
         )

references <- data.frame(`First author`, Year, `Number of networks`, Country,   DOI)

colnames(references) <- c("First author", "Year", "Number of networks", "Country",  "DOI")

#Check number of studies
#nrow(references) # 18 studies
#check number of networks
#sum(as.numeric(references$`Number of networks`)) #60

references %>%
  arrange(`First author`)%>%
  kable( longtable = T, booktabs = T,linesep = "\\addlinespace",align = c("cccc"),caption = "Table S1. List of studies ordered by author with the year of publication, number of contributed networks and digital object identifier") %>%
  kable_styling(latex_options = c("repeat_header","striped"), font_size = 12, full_width=F,position = c("center"))

```

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=12, fig.width=12}

library(kableExtra)

library(dplyr)
library(readxl) #read excel file (trait data)
library(dplyr) #data manipulation
library(visdat) #VISUALIZE MISSING DATA
library(naniar)
library(tidyverse)
########################################################################################################################################################
#1) READ TRAIT DATA
########################################################################################################################################################

Trait <- c("Plant height (m)", "Flower width (mm)", "Flower length (mm)", "Inflorescence width (mm)", "Style length (mm)", "Ovules per flower",
           "Flowers per plant",  "Autonomous selfing (fruit set)")

Trait1 <- c("Lifepan", "Life form", "Flower shape", "Flower symmetry", "Autonomous selfing", "Compatibility system", "Breeding system", "")

T.cat <- linebreak(c("Short-lived \n Perennial", "Herb \n Shrub \n Tree", "Brush \n Campanulate \n Capitulum \n Open \n Papilionaceous \n Tube",  "Actinomorphic \n Zygomorphic", "None \n Low \n Medium \n High", "Self-incomp. \n Partially self-comp. \n Self-comp.", "Hermaphrodite \n Monoecious \n Dioecious",""),align = c("l"))

Type <- c("Vegetative", "Floral", "Floral", "Floral", "Floral", "Floral", "Floral","Reproductive")

Type_1 <- c("Vegetative", "Vegetative", "Floral", "Floral", "Reproductive", "Reproductive", "Reproductive", "")

dat <- data.frame(Type, Trait,Type_1, Trait1)

colnames(dat) <- c("Type","Traits","Type", "Traits")

kbl(dat, booktabs = T, linesep = "\\addlinespace", escape = FALSE, align = c("c","l","c","c","l","l","c"),caption = "Table S2. Traits used to delimit the different plant functional groups divided in quantitative and categorical traits.") %>%
kable_styling(full_width = F) %>%
add_header_above(c("Quantitative traits" = 2,"Categorical traits" = 2),bold=T)  %>%
    column_spec(c(1,4), bold = T) %>% row_spec(0, bold=T)

```

**FIGURES**

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=13, fig.cap="Figure S1. Plant functional group composition separated in qualitative and quantitative variables. Panel A) shows the percentage of the different categories within trait represented with different colours for each functional group. Plot B) shows the radar plot of the different quantitative variables standardize on the same scale also coloured with the same patterns of colours as qualitative variables per cluster."}

#CODE TO PLOT FUNCTIONAL GROUPS
#IT CAN BE PLOT WITH MUCH LESS CODE BUT NO TIME FOR CLEANING IT

##########################################
#Visualization of plant functional groups#
##########################################

library(cowplot)
library(ggplot2)
library(tidyr)
library(dplyr)
library(reshape2)
library(data.table)
library(patchwork)

#LOAD DATA
#read unscaled trait data in order to visualize better the clusters
d <- read.csv("../Data/Csv/all_species_imputed_trait_data_forest_data.csv")

dat_1 <- read.csv("../Data/Csv/imputed_trait_data_hclust_5_clusters_forest_data.csv", row.names = "X")

d$Clusters <- dat_1$Clusters

#select columns of interest
t <- d[c("Breeding_system","IMPUTED_Compatibility","Autonomous_selfing_level",
         "Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", "Flowers_per_plant", "Flowers_per_inflorescence",
         "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean", "STYLE_IMPUTED", "OVULES_IMPUTED", "life_form", "lifespan",
         "IMPUTED_plant_height_mean_m", "Nectar_presence_absence", "Clusters")]

##############################################################################################################################
#Prepare table qualitative variables
##############################################################################################################################

##############################################################################################################################
#Breeding system
t$Breeding_system <- factor(t$Breeding_system, levels=c("Dioecious", "Monoecious", "Hermaphrodite"))

#Summary to create table
breeding <- t %>%
  group_by(Clusters,Breeding_system) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

breeding <- breeding[,-3]

breeding <- spread(breeding, Clusters, Percentage)

colnames(breeding)[1] <- "Category"

Trait <- data.frame(Trait= c(rep("Breeding system", 3)))

Breeding_system <- cbind(Trait, breeding)

##############################################################################################################################
#Compatibility
str(t)
t$IMPUTED_Compatibility <- as.character(t$IMPUTED_Compatibility)
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="dioecious"] <- "Unisexual flower"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="monoecious"] <- "Unisexual flower"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_incompatible"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="partially_self_compatible"] <- "Partially self compatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_compatible"] <- "Self compatible"

t$IMPUTED_Compatibility <- factor(t$IMPUTED_Compatibility, levels=c("Unisexual flower", "Self incompatible", "Partially self compatible", "Self compatible"))
#Summary to create table
comp <- t %>%
  group_by(Clusters,IMPUTED_Compatibility) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

comp <- comp[,-3]

comp <- spread(comp, Clusters, Percentage)
#Replace Na's with zeros
comp[is.na(comp)] <- 0
#Set call name os categories per trait
colnames(comp)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Compatibility system", 4)))
#Add column
Compatibility_system <- cbind(Trait, comp)
#Bind rows
bind_rows(Breeding_system, Compatibility_system)

############################################################################################################################################
#Life form
t$life_form <- as.character(t$life_form)
t$life_form[t$life_form =="vine"] <- "Shrub"
t$life_form[t$life_form =="tree"] <- "Tree"
t$life_form[t$life_form =="herb"] <- "Herb"
t$life_form[t$life_form =="shrub"] <- "Shrub"
#Order levels
t$life_form <- factor(t$life_form, levels=c("Tree", "Shrub", "Herb"))
#Summary to create table
life_form <- t %>%
  group_by(Clusters,life_form) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_form <- life_form[,-3]
#convert to wide
life_form <- spread(life_form, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_form)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life form", 3)))
#Add column
Life_form<- cbind(Trait, life_form)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form)

############################################################################################################################################
t$lifespan <- factor(t$lifespan, levels=c("Perennial","Short lived"))
#Summary to create table
life_span <- t %>%
  group_by(Clusters,lifespan) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_span <- life_span[,-3]
#convert to wide
life_span <- spread(life_span, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_span)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life span", 2)))
#Add column
Life_span<- cbind(Trait, life_span)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span)

############################################################################################################################################
t$Flower_morphology <- as.character(t$Flower_morphology)
t$Flower_morphology[t$Flower_morphology=="Funnelform"] <- "Campanulate"
t$Flower_morphology[t$Flower_morphology=="Spike"] <- "Brush"

Flower_morphology <- t %>%
  group_by(Clusters,Flower_morphology) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
Flower_morphology <- Flower_morphology[,-3]
#convert to wide
Flower_morphology <- spread(Flower_morphology, Clusters, Percentage)
#Set call name os categories per trait
colnames(Flower_morphology)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Flower shape", 6)))
#Add column
Flower_morphology<- cbind(Trait, Flower_morphology)
#Replace Na's with zeros
Flower_morphology[is.na(Flower_morphology)] <- 0
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology)

############################################################################################################################################
t$Flower_symmetry <- as.character(t$Flower_symmetry)
t$Flower_symmetry[t$Flower_symmetry =="actinomorphic"] <- "Actinomorphic"
t$Flower_symmetry[t$Flower_symmetry =="zygomorphic"] <- "Zygomorphic"

Flower_symmetry <- t %>%
  group_by(Clusters,Flower_symmetry) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
Flower_symmetry <- Flower_symmetry[,-3]
#convert to wide
Flower_symmetry <- spread(Flower_symmetry, Clusters, Percentage)
#Set call name os categories per trait
colnames(Flower_symmetry)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Flower symmetry", 2)))
#Add column
Flower_symmetry<- cbind(Trait, Flower_symmetry)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology,Flower_symmetry)
############################################################################################################################################
t$Autonomous_selfing_level <- as.character(t$Autonomous_selfing_level)
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="none"] <- "None"
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="low"] <- "Low"
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="medium"] <- "Medium"
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="high"] <- "High"

#Order levels
t$Autonomous_selfing_level <- factor(t$Autonomous_selfing_level, levels=c("High", "Medium", "Low", "None"))

Selfing <- t %>%
  group_by(Clusters,Autonomous_selfing_level) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

#delete column to convert to wide
Selfing <- Selfing[,-3]
#convert to wide
Selfing <- spread(Selfing, Clusters, Percentage)
#Set call name os categories per trait
colnames(Selfing)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Selfing level", 2)))
#Replace Na's with zeros
Selfing[is.na(Selfing)] <- 0
#Add column
Selfing_1 <- cbind(Trait, Selfing)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Selfing_1)

############################################################################################################################################
t$Nectar_presence_absence <- as.character(t$Nectar_presence_absence)
t$Nectar_presence_absence[t$Nectar_presence_absence =="yes"] <- "Yes"
t$Nectar_presence_absence[t$Nectar_presence_absence =="no"] <- "No"

Nectar <- t %>%
  group_by(Clusters,Nectar_presence_absence) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

#delete column to convert to wide
Nectar <- Nectar[,-3]
#convert to wide
Nectar <- spread(Nectar, Clusters, Percentage)
#Set call name os categories per trait
colnames(Nectar)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Nectar", 2)))
#Replace Na's with zeros
Nectar[is.na(Nectar)] <- 0
#Add column
Nectar <- cbind(Trait, Nectar)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Flower_symmetry,Selfing_1, Nectar)

##############################################################################################################################
#Prepare table quantitative variables
##############################################################################################################################

Selfing <-  t %>% group_by(Clusters)%>% 
  select(Autonomous_selfing_level_fruit_set) %>% # select variables to summarise
  summarise_each(funs(min = min,
                     # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                    sd = sd))

s1 <- as.data.frame(t(Selfing))
#set colnames
colnames(s1) <- c("1","2","3","4","5")
#remove the cluster row
s1 <- s1[-1,]
#convert rownames to 1st columns
s1 <- setDT(s1, keep.rownames = "TRUE")[]
colnames(s1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Autonomous selfing", 5)))
#Add column
Selfing <- cbind(Trait, s1)

##############################################################################################################################
flower_number <-  t %>% group_by(Clusters)%>% 
  select(Flowers_per_plant) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

f1 <- as.data.frame(t(flower_number))
#set colnames
colnames(f1) <- c("1","2","3","4","5")
#remove the cluster row
f1 <- f1[-1,]
#convert rownames to 1st columns
f1 <- setDT(f1, keep.rownames = "TRUE")[]
colnames(f1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flower number", 5)))
#Add column
Flower_number <- cbind(Trait, f1)

##############################################################################################################################
Flowers_per_inflorescence <-  t %>% group_by(Clusters)%>% 
  select(Flowers_per_inflorescence) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

f_i_1 <- as.data.frame(t(Flowers_per_inflorescence))
#set colnames
colnames(f_i_1) <- c("1","2","3","4","5")
#remove the cluster row
f_i_1 <- f_i_1[-1,]
#convert rownames to 1st columns
f_i_1 <- setDT(f_i_1, keep.rownames = "TRUE")[]
colnames(f_i_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flowers per inflorescence", 5)))
#Add column
Flower_per_inflo <- cbind(Trait, f_i_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo)

##############################################################################################################################
Floral_unit_width <-  t %>% group_by(Clusters)%>% 
  select(Floral_unit_width) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

f_u_w_1 <- as.data.frame(t(Floral_unit_width))
#set colnames
colnames(f_u_w_1) <- c("1","2","3","4","5")
#remove the cluster row
f_u_w_1 <- f_u_w_1[-1,]
#convert rownames to 1st columns
f_u_w_1 <- setDT(f_u_w_1, keep.rownames = "TRUE")[]
colnames(f_u_w_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Inflorescence width", 5)))
#Add column
Inflo_width <- cbind(Trait, f_u_w_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width)

##############################################################################################################################
Corolla_diameter_mean <-  t %>% group_by(Clusters)%>% 
  select(Corolla_diameter_mean) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

c_w_1 <- as.data.frame(t(Corolla_diameter_mean))
#set colnames
colnames(c_w_1) <- c("1","2","3","4","5")
#remove the cluster row
c_w_1 <- c_w_1[-1,]
#convert rownames to 1st columns
c_w_1 <- setDT(c_w_1, keep.rownames = "TRUE")[]
colnames(c_w_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flower width", 5)))
#Add column
Flower_width <- cbind(Trait, c_w_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width)

##############################################################################################################################
Corolla_length_mean <-  t %>% group_by(Clusters)%>% 
  select(Corolla_length_mean) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

c_l_1 <- as.data.frame(t(Corolla_length_mean))
#set colnames
colnames(c_l_1) <- c("1","2","3","4","5")
#remove the cluster row
c_l_1 <- c_l_1[-1,]
#convert rownames to 1st columns
c_l_1 <- setDT(c_l_1, keep.rownames = "TRUE")[]
colnames(c_l_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flower length", 5)))
#Add column
Flower_length<- cbind(Trait, c_l_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length)

##############################################################################################################################
STYLE_IMPUTED <-  t %>% group_by(Clusters)%>% 
  select(STYLE_IMPUTED) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

s_l_1 <- as.data.frame(t(STYLE_IMPUTED))
#set colnames
colnames(s_l_1) <- c("1","2","3","4","5")
#remove the cluster row
s_l_1 <- s_l_1[-1,]
#convert rownames to 1st columns
s_l_1 <- setDT(s_l_1, keep.rownames = "TRUE")[]
colnames(s_l_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Style length", 5)))
#Add column
Style_length<- cbind(Trait, s_l_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length, Style_length)

##############################################################################################################################

OVULES_IMPUTED <-  t %>% group_by(Clusters)%>% 
  select(OVULES_IMPUTED) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

o_n_1 <- as.data.frame(t(OVULES_IMPUTED))
#set colnames
colnames(o_n_1) <- c("1","2","3","4","5")
#remove the cluster row
o_n_1 <- o_n_1[-1,]
#convert rownames to 1st columns
o_n_1 <- setDT(o_n_1, keep.rownames = "TRUE")[]
colnames(o_n_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Ovule number", 5)))
#Add column
Ovule_number <- cbind(Trait, o_n_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length, Ovule_number)

##############################################################################################################################

IMPUTED_plant_height_mean_m <-  t %>% group_by(Clusters)%>% 
  select(IMPUTED_plant_height_mean_m) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

p_h_1 <- as.data.frame(t(IMPUTED_plant_height_mean_m))
#set colnames
colnames(p_h_1) <- c("1","2","3","4","5")
#remove the cluster row
p_h_1 <- p_h_1[-1,]
#convert rownames to 1st columns
p_h_1 <- setDT(p_h_1, keep.rownames = "TRUE")[]
colnames(p_h_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Plant height", 5)))
#Add column
Plant_height <- cbind(Trait, p_h_1)

#Bind quantitative and qualitative variables
Table_all_traits <- bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Flower_symmetry,Selfing_1, Nectar,
          Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length, Style_length, Ovule_number, Plant_height)

#Select just two decimals
is.num <- sapply(Table_all_traits, is.numeric)
Table_all_traits[is.num] <- lapply(Table_all_traits[is.num], round, 2)

colnames(Table_all_traits) <- c("Trait", "Category", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")

#LOAD LIBRARIES
#devtools::install_github("ricardo-bion/ggradar", dependencies=TRUE)
suppressPackageStartupMessages(library(dplyr))
library(scales)
library(tibble)
library(ggradar)
library(janitor)
library(grid)
library(dplyr)
library(lubridate)
#QUANTITATIVVE VARIABLES

#generate subset of quantitative variables for plotting
quant_median <- subset(Table_all_traits, Category=="median")

#delete col number 2
quant_median <- quant_median[,-2]

#transpose dataframe
final_df <- as.data.frame(t(quant_median))

#rownames to colnames (1st one)
final_df_1 <- final_df %>% row_to_names(row_number = 1)

#Convert all variables to numeric
final_df_2 <- sapply(final_df_1, function(x) as.numeric(as.character(x)))

#set back the missed rownames in the processing
rownames(final_df_2) <- rownames(final_df_1)

# Convert to dataframe
final_df_2 <- as.data.frame(final_df_2)

# Scale values and set rowname as column
final_df_3 <-final_df_2 %>% rownames_to_column( var = "group" ) %>%   mutate_at(vars(-group),funs(rescale))

#Plot Radar
#ggradar(final_df_3)

#QUALITATIVE VARIABLES

#LOAD DATA
#read unscaled trait data in order to visualize better the clusters
d <- read.csv("../Data/Csv/all_species_imputed_trait_data_forest_data.csv")

dat_1 <- read.csv("../Data/Csv/imputed_trait_data_hclust_5_clusters_forest_data.csv", row.names = "X")

d$Clusters <- dat_1$Clusters

#select columns of interest
t <- d[c("Breeding_system","IMPUTED_Compatibility","Autonomous_selfing_level",
         "Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", "Flowers_per_plant", "Flowers_per_inflorescence",
         "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean", "STYLE_IMPUTED", "OVULES_IMPUTED", "life_form", "lifespan",
         "IMPUTED_plant_height_mean_m", "Nectar_presence_absence", "Clusters")]

##############################################################################################################################
#Prepare table qualitative variables
##############################################################################################################################

##############################################################################################################################
#Breeding system
t$Breeding_system <- factor(t$Breeding_system, levels=c("Dioecious", "Monoecious", "Hermaphrodite"))

#Summary to create table
breeding <- t %>%
  group_by(Clusters,Breeding_system) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

breeding <- breeding[,-3]

breeding <- spread(breeding, Clusters, Percentage)

colnames(breeding)[1] <- "Category"

Trait <- data.frame(Trait= c(rep("Breeding system", 3)))

Breeding_system <- cbind(Trait, breeding)

##############################################################################################################################
#Compatibility
str(t)
t$IMPUTED_Compatibility <- as.character(t$IMPUTED_Compatibility)
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="dioecious"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="monoecious"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_incompatible"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="partially_self_compatible"] <- "Partially self compatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_compatible"] <- "Self compatible"

t$IMPUTED_Compatibility <- factor(t$IMPUTED_Compatibility, levels=c("Self incompatible", "Partially self compatible", "Self compatible"))
#Summary to create table
comp <- t %>%
  group_by(Clusters,IMPUTED_Compatibility) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

comp <- comp[,-3]

comp <- spread(comp, Clusters, Percentage)
#Replace Na's with zeros
comp[is.na(comp)] <- 0
#Set call name os categories per trait
colnames(comp)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Compatibility system", 3)))
#Add column
Compatibility_system <- cbind(Trait, comp)
#Bind rows
bind_rows(Breeding_system, Compatibility_system)

############################################################################################################################################
#Life form
t$life_form <- as.character(t$life_form)
t$life_form[t$life_form =="vine"] <- "Shrub"
t$life_form[t$life_form =="tree"] <- "Tree"
t$life_form[t$life_form =="herb"] <- "Herb"
t$life_form[t$life_form =="shrub"] <- "Shrub"
#Order levels
t$life_form <- factor(t$life_form, levels=c("Tree", "Shrub", "Herb"))
#Summary to create table
life_form <- t %>%
  group_by(Clusters,life_form) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_form <- life_form[,-3]
#convert to wide
life_form <- spread(life_form, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_form)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life form", 3)))
#Add column
Life_form<- cbind(Trait, life_form)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form)

############################################################################################################################################
t$lifespan <- factor(t$lifespan, levels=c("Perennial","Short lived"))
#Summary to create table
life_span <- t %>%
  group_by(Clusters,lifespan) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_span <- life_span[,-3]
#convert to wide
life_span <- spread(life_span, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_span)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life span", 2)))
#Add column
Life_span<- cbind(Trait, life_span)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span)

############################################################################################################################################
#t$Flower_morphology <- as.character(t$Flower_morphology)
#t$Flower_morphology[t$Flower_morphology=="Funnelform"] <- "Campanulate"
#t$Flower_morphology[t$Flower_morphology=="Spike"] <- "Brush"
#
#Flower_morphology <- t %>%
#  group_by(Clusters,Flower_morphology) %>%
#  summarise (n = n()) %>%
#  mutate(Percentage = n / sum(n)*100)
##delete column to convert to wide
#Flower_morphology <- Flower_morphology[,-3]
##convert to wide
#Flower_morphology <- spread(Flower_morphology, Clusters, Percentage)
##Set call name os categories per trait
#colnames(Flower_morphology)[1] <- "Category"
##Create new column
#Trait <- data.frame(Trait= c(rep("Flower shape", 6)))
##Add column
#Flower_morphology<- cbind(Trait, Flower_morphology)
##Replace Na's with zeros
#Flower_morphology[is.na(Flower_morphology)] <- 0
##Bind rows
#bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology)
#
############################################################################################################################################
t$Flower_symmetry <- as.character(t$Flower_symmetry)
t$Flower_symmetry[t$Flower_symmetry =="actinomorphic"] <- "Actinomorphic"
t$Flower_symmetry[t$Flower_symmetry =="zygomorphic"] <- "Zygomorphic"

Flower_symmetry <- t %>%
  group_by(Clusters,Flower_symmetry) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
Flower_symmetry <- Flower_symmetry[,-3]
#convert to wide
Flower_symmetry <- spread(Flower_symmetry, Clusters, Percentage)
#Set call name os categories per trait
colnames(Flower_symmetry)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Flower symmetry", 2)))
#Add column
Flower_symmetry<- cbind(Trait, Flower_symmetry)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology,Flower_symmetry)
############################################################################################################################################
#t$Autonomous_selfing_level <- as.character(t$Autonomous_selfing_level)
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="none"] <- "None"
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="low"] <- "Low"
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="medium"] <- "Medium"
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="high"] <- "High"
#
##Order levels
#t$Autonomous_selfing_level <- factor(t$Autonomous_selfing_level, levels=c("High", "Medium", "Low", "None"))
#
#Selfing <- t %>%
#  group_by(Clusters,Autonomous_selfing_level) %>%
#  summarise (n = n()) %>%
#  mutate(Percentage = n / sum(n)*100)
#
##delete column to convert to wide
#Selfing <- Selfing[,-3]
##convert to wide
#Selfing <- spread(Selfing, Clusters, Percentage)
##Set call name os categories per trait
#colnames(Selfing)[1] <- "Category"
##Create new column
#Trait <- data.frame(Trait= c(rep("Selfing level", 2)))
##Replace Na's with zeros
#Selfing[is.na(Selfing)] <- 0
##Add column
#Selfing_1 <- cbind(Trait, Selfing)
##Bind rows
#bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Selfing_1)
#
#############################################################################################################################################
t$Nectar_presence_absence <- as.character(t$Nectar_presence_absence)
t$Nectar_presence_absence[t$Nectar_presence_absence =="yes"] <- "Yes"
t$Nectar_presence_absence[t$Nectar_presence_absence =="no"] <- "No"

Nectar <- t %>%
  group_by(Clusters,Nectar_presence_absence) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

#delete column to convert to wide
Nectar <- Nectar[,-3]
#convert to wide
Nectar <- spread(Nectar, Clusters, Percentage)
#Set call name os categories per trait
colnames(Nectar)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Nectar", 2)))
#Replace Na's with zeros
Nectar[is.na(Nectar)] <- 0
#Add column
Nectar <- cbind(Trait, Nectar)
#Bind rows
data <- bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span, Flower_symmetry, Nectar)

#remove first col
data_1 <- data[,-1]
colnames(data_1) <- c("Category", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")

#transpose dataframe
data_2 <- as.data.frame(t(data_1))

#rownames to colnames (1st one)
data_3 <- data_2 %>% row_to_names(row_number = 1)

#Convert all variables to numeric
data_4 <- sapply(data_3, function(x) as.numeric(as.character(x)))

#set back the missed rownames in the processing
rownames(data_4) <- rownames(data_3)

# Convert to dataframe
data_4 <- as.data.frame(data_4)

# Scale values and set rowname as column
data_5 <-data_4 %>% rownames_to_column( var = "group" )

data_6 <- melt(data_5)

cluster1 <- subset(data_6, group=="Cluster 1")

cluster1$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

#Try to achieve the desire output with plot_grid

c1 <- cluster1 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("sandybrown"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(axis.title.x = element_blank(), axis.text.x=element_blank(),plot.title = element_text(size = 18))+ ylab("")+
  ggtitle("A) Qualitative variables")

cluster2 <- subset(data_6, group=="Cluster 2")

cluster2$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

c2 <- cluster2 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("#7BB0A3"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(strip.background = element_blank(), strip.text = element_blank())+
  theme(axis.title.x = element_blank(), axis.text.x=element_blank())+ylab("")

cluster3 <- subset(data_6, group=="Cluster 3")

cluster3$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

c3 <- cluster3 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("lightsteelblue2"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(axis.title.x = element_blank(), axis.text.x=element_blank())+
  theme(strip.background = element_blank(), strip.text = element_blank())+ylab("Percentage")

cluster4 <- subset(data_6, group=="Cluster 4")

cluster4$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

c4 <- cluster4 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("thistle"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(axis.title.x = element_blank(), axis.text.x=element_blank())+
  theme(strip.background = element_blank(), strip.text = element_blank())+ylab("")

cluster5 <- subset(data_6, group=="Cluster 5")

cluster5$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))
c5 <- cluster5 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("lightgoldenrod2"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 45,vjust = 1, hjust=1))+
  theme(strip.background = element_blank(), strip.text = element_blank())+ylab("")+xlab("")

A <- c1 + c2 + c3 + c4 + c5 + plot_layout(ncol = 1)
B <- ggradar(final_df_3, group.colours=c("sandybrown", "#7BB0A3","lightsteelblue2", "thistle", "lightgoldenrod2"),
             legend.text.size=5, grid.label.size = 0,legend.position = "none",axis.label.size=3)+ 
  guides(shape = guide_legend(override.aes = list(size = 0.75)))+ ggtitle("B) Quantitative variables")+
  theme(plot.title = element_text(size = 18))

patch <- A | B

patch + plot_annotation(title = "Plant functional group composition",theme = theme(plot.title = element_text(size = 20,face="bold")))

```

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=13, fig.cap="Figure S2. Hierarchical clustering dendrogram with the branches coloured by the optimal number of clusters (5). The labels of the subgroup of species (N = 524) used in this study are coloured in black in order to show the evenness of the distribution of the species across clusters. The rest of species labels are omitted for visualization purposes (N = 982)."}

library(ggplot2)
library(dendextend)
library(dplyr)

e.clust_5 <- readRDS("../Data/RData/e.clust_5.rds")

dendro <- as.dendrogram(e.clust_5)
dendro.col <- dendro %>%
  set("branches_k_color", k = 5, value =   c("lightsteelblue2", "thistle", "sandybrown", "#7BB0A3", "lightgoldenrod2")) %>%
  set("branches_lwd", 0.7) %>%
  set("labels_colors", 
      value = c("darkslategray")) %>% 
  set("labels_cex", 0.075)

labels_colors(dendro.col) <- get_leaves_branches_col(dendro.col)

# We use "darkgrey" to color some species
final_d_1 <- read.csv("../Data/Csv/data_for_motifs_analysis_1.csv")

plant_species <- unique(final_d_1$Plant_species)

index_species_out <- which(names(labels_colors(dendro.col)) %in% plant_species)

'%ni%' <- Negate('%in%')

index_species_in <- which(names(labels_colors(dendro.col)) %ni% plant_species)

labels_colors(dendro.col)[index_species_out] <- "black"
labels_colors(dendro.col)[index_species_in] <- "white"

ggd1 <- as.ggdend(dendro.col)

ggd1$labels$angle <- 90
ggd1$labels$vjust <- 1
ggd1$labels$hjust <- 1

ggplot(ggd1,offset_labels=-0.05) + theme(panel.grid.major = element_blank(),
 axis.text = element_blank(),axis.title = element_blank()) +
  ggtitle("Plant functional groups")+
  theme(plot.title = element_text(hjust = 0.5, vjust = -2,size = 20,face="bold"))+
  annotate("text", x=140, y=1.65, label= "Self-incomp. perennials with large flowers ", size=2.5)+
  annotate("text", x=380, y=2.15, label= "Tall plants with small unisexual flowers", size=2.5) +
  annotate("text", x=600, y=3.85, label= "Short-lived selfers", size=2.5) +
  annotate("text", x=1400, y=2.05, label= "Short-lived outcrossers with \n long zygomorphic flowers", size=2.5) +annotate("text", x=950, y=2.05, label= "Small outcrossing perennials", size=2.5)

```

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.cap="Figure S3. Comparison of the motif frequencies between empirical and simulated networks. Average percentages of motifs close to 0 and 100 indicate under- and over- representation in empirical networks, after removing non-robust links, that is, interactions whose frequency was equal to one. The different motifs are coloured by the mean path length as done in @simmons2020.", fig.width=9,fig.height=5}
library(lme4)
library(nlme)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models_NOSING/Motifs_frequency_percentile_NOSING.csv")
d <- add_study_id(d)
#str(d)
#head(d)

#ggplot(d %>% filter(motif != 1), aes(percentil_sizeclass))+
#  geom_histogram(color="black", fill="white")+
#  facet_wrap(~motif)+
#  theme_bw()+
#  labs(x="Sizeclass percentile")
#
# Motifs tend to be over- and under-represensented in real networks

# Mean percentile and SE taking into account the study system
motif_codes <- d %>% filter(motif != 1) %>% # We remove links (motif code: 1)
  select(motif) %>% unique() %>% pull()
motif_means <- tibble(motif = motif_codes)
motif_means$mean <- NA
motif_means$SE <- NA

for(i.motif in 1:length(motif_codes)){

  m <- lmer(percentil_sizeclass ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(d, motif == motif_codes[i.motif]))

  motif_means$mean[i.motif] <- fixed.effects(m) %>% unname()
  motif_means$SE[i.motif] <- sqrt(diag(vcov(m)))

}

#Add categories based on path length set by Simmons 2021 et al., Functional Ecology
d$Broad_categories[d$motif==2 | d$motif==7 | d$motif==17 | d$motif==3|d$motif==4|d$motif==8] <- "Fan"

d$Broad_categories[d$motif==15 | d$motif==11] <- "Medium-weak"

d$Broad_categories[d$motif==6 | d$motif==16 |  d$motif==12] <- "Strong"

d$Broad_categories[d$motif==5 | d$motif==9 |  d$motif==14|  d$motif==13|  d$motif==10] <- "Weak"

#order the categories from short to long path length
d$Broad_categories <- factor(d$Broad_categories, levels = c("Strong", "Fan", "Medium-weak", "Weak"))

ggplot(NULL) + 
  geom_point(data = d %>% filter(motif != 1),
             aes(y=as.factor(motif), x=100*percentil_sizeclass, 
                 color = as.factor(Broad_categories)),
             position = "jitter",alpha=0.5)+
  geom_errorbar(data = motif_means,aes(y = as.factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=1.0,size=1)+
  geom_point(data = motif_means,aes(y = as.factor(motif), x=100*mean), 
                size=2)+
  labs(y="Motif", x = "Percentile (%)",colour="Path length classification")+
  theme_bw() + scale_color_manual(labels = c('Strong' = expression(paste("Complete (", bar(x), " = 1.38)")),'Fan' = expression(paste("Fan (", bar(x), " = 1.48)")), 'Medium-weak' = expression(paste("Asymmetric complete (", bar(x), " = 1.60)")), "Weak" =expression(paste("Core-peripheral (", bar(x), " = 1.85)"))), name = "Path length classification",values = c("grey25","springgreen4","sienna2","goldenrod2"),guide = guide_legend(title.position = "top",nrow = 2)) +
theme(legend.text.align = 0)+
  theme(legend.position="bottom")

#Trying to maintain a gradient from low to high in colors

#Fig 1 Should be a dot plot (or forest plot) of the mean +- SE of the 15 motifs.
#NICE! Can we put motifs in Y axes, and % in X?
```

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=8,fig.width=8,fig.align = 'center', fig.cap="Figure S4. Heatmap indicating under- and over- representation of pollinator and plant functional groups in the different motif positions, after removing non-robust links, that is, interactions whose frequency was equal to one. The different motif positions are dividied by the average path length clasification by @simmons2020."}

library(lme4)
library(nlme)
library(tidyverse)
library(stringi)
source("../Scripts/Scripts_Alfonso/add_study_id.R")

#############################################
# EXTRACT MEAN VALUES
#############################################

all_position_percentiles <- read_csv("../Data/Csv/Motifs_positions_and_null_models_NOSING/GF_positions_frequency_percentile_NOSING.csv")

all_position_percentiles$position <- stri_extract_first_regex(all_position_percentiles$position,
                                                              "[0-9]+") %>% as.numeric()

all_position_percentiles <- add_study_id(all_position_percentiles)
#str(all_position_percentiles)
#head(all_position_percentiles)

list_plants_FG <- as.character(1:10)

plant_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF)) #ME DA ERROR
pollinator_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(!Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF))

# Plants-------------

# Mean percentile and SE taking into account the study system
plant_codes <- plant_position_percentiles_filtered$position %>% unique()
plant_means <- plant_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
plant_means$mean <- NA
plant_means$SE <- NA

for(i.pos in 1:length(plant_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id),
            data = plant_position_percentiles_filtered %>%
              filter(position == plant_codes[i.pos]))

  temp <- fixed.effects(m) %>% unname()
  plant_means$mean[plant_means$position == plant_codes[i.pos]] <- c(temp[1],
                                                                    temp[1]+temp[2],
                                                                    temp[1]+temp[3],
                                                                    temp[1]+temp[4],
                                                                    temp[1]+temp[5]) ## to get real means per FG
  temp <- sqrt(diag(vcov(m)))
  plant_means$SE[plant_means$position == plant_codes[i.pos]] <- c(temp[1], 
                                                                  temp[1]+temp[2],
                                                                  temp[1]+temp[3],
                                                                  temp[1]+temp[4],
                                                                  temp[1]+temp[5])

}

# Pollinators--------

# Mean percentile and SE taking into account the study system
pollinator_codes <- pollinator_position_percentiles_filtered$position %>% unique()
pollinator_means <- pollinator_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
pollinator_means$mean <- NA
pollinator_means$SE <- NA
pollinator_means_reordered <- NULL

for(i.pos in 1:length(pollinator_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = pollinator_position_percentiles_filtered %>%
              filter(position == pollinator_codes[i.pos]),
            control = lmerControl(optimizer ="Nelder_Mead")) # I changed the optimizer because motif 8 (i.pos = 3) caused convergence problems, max gradient = 0.0022 > 0.002.

  temp <- fixed.effects(m) %>% unname()
  pollinator_means_aux <- pollinator_means[pollinator_means$position == pollinator_codes[i.pos],] %>% arrange(Node_FG)

  pollinator_means_aux$mean <- c(temp[1],
                                 temp[1]+temp[2],
                                 temp[1]+temp[3],
                                 temp[1]+temp[4],
                                 temp[1]+temp[5],
                                 temp[1]+temp[6],
                                 temp[1]+temp[7],
                                 temp[1]+temp[8],
                                 temp[1]+temp[9]) ## to get real means per FG

  temp <- sqrt(diag(vcov(m)))
  pollinator_means_aux$SE <- c(temp[1],
                               temp[1]+temp[2],
                               temp[1]+temp[3],
                               temp[1]+temp[4],
                               temp[1]+temp[5],
                               temp[1]+temp[6],
                               temp[1]+temp[7],
                               temp[1]+temp[8],
                               temp[1]+temp[9])

  pollinator_means_reordered <- bind_rows(pollinator_means_reordered,pollinator_means_aux)

}

# There are lower limits (mean - SE) smaller than 0 (see Other insects)
# We set the lower limit of those error bars to zero

pollinator_means_reordered <- pollinator_means_reordered %>% mutate(lower = mean-SE)
pollinator_means_reordered$lower[pollinator_means_reordered$lower < 0] <- 0

#############################################
# ADDING BROAD MOTIF CATEGORIES TO POSITIONS
#############################################

motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

motifs_raw_data$Broad_categories <- NA

motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(5,9,14,13,10)] <- "Core-peripherical"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(6,16,12)] <- "Complete"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(2,3,7,4,17,8)] <- "Fan"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(15,11)] <- "Asymmetric complete"

motifs_raw_data$plant <- gsub("[^0-9.-]", "", motifs_raw_data$plant)
motifs_raw_data$pollinator <- gsub("[^0-9.-]", "", motifs_raw_data$pollinator)
motifs_raw_data$plant <- as.numeric(motifs_raw_data$plant)
motifs_raw_data$pollinator <- as.numeric(motifs_raw_data$pollinator)
motifs_raw_data <- motifs_raw_data %>% filter(!is.na(Broad_categories))

fan_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

pollinator_means_reordered$Broad_categories <- NA

pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% fan_pollinator] <- "Fan" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% asymm_pollinator] <- "Medium-weak" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% complete_pollinator] <- "Strong" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% core_per_pollinator] <- "Weak"

fan_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

plant_means$Broad_categories <- NA

plant_means$Broad_categories[plant_means$position %in% fan_plant] <- "Fan" 
plant_means$Broad_categories[plant_means$position %in% asymm_plant] <- "Medium-weak" 
plant_means$Broad_categories[plant_means$position %in% complete_plant] <- "Strong" 
plant_means$Broad_categories[plant_means$position %in% core_per_plant] <- "Weak"

########################################
# CREATE HEATMAPS
########################################

library(ComplexHeatmap)

# PLANTS-----

# Prepare data for heatmap
heat_plants_aux <- plant_means %>% spread(Node_FG,mean) %>% select(-SE)

heat_plants <- heat_plants_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_plants_fan <- heat_plants %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_fan) <- heat_plants_fan[,1]

heat_plants_Weak <- heat_plants %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Weak) <- heat_plants_Weak[,1]

heat_plants_Strong <- heat_plants %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Strong) <- heat_plants_Strong[,1]

heat_plants_Medium_weak <- heat_plants %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Medium_weak) <- heat_plants_Medium_weak[,1]

# POLLINATORS-----------

filter_rare_groups <- T

if(filter_rare_groups == T){
  heat_pollinators_aux <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds",
                                                                                "Lizards",
                                                                                "Other_insects")) %>% 
    spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee",            
                                      "Coleoptera","Lepidoptera","Non-bee\nHymenoptera", 
                                      "Non-syrphids\ndiptera","Syrphids")
}else{
  heat_pollinators_aux <- pollinator_means_reordered %>% spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee","Birds",            
                                      "Coleoptera","Lepidoptera","Lizards","Non-bee\nHymenoptera", 
                                      "Non-syrphids\ndiptera","Other insects","Syrphids")
}

heat_pollinators <- heat_pollinators_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_pollinators_fan <- heat_pollinators %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_fan) <- heat_pollinators_fan[,1]

heat_pollinators_Weak <- heat_pollinators %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Weak) <- heat_pollinators_Weak[,1]

heat_pollinators_Strong <- heat_pollinators %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Strong) <- heat_pollinators_Strong[,1]

heat_pollinators_Medium_weak <- heat_pollinators %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Medium_weak) <- heat_pollinators_Medium_weak[,1]

##################

#Do not know how to handle this multiplot for creating a panel

# I'm going to create the 4 plots in one, 
# I'm not sure that the grouping of the functional group is correct
# when is done separately

#sorry for the mess of libraries that I do below

 #Prepare first pollinators

library(tidyr)
library(data.table)
long <- melt(setDT(heat_pollinators), id.vars = c("position","Broad_categories"), variable.name = "guilds")

long_1 <- long %>% select(-Broad_categories)

library(reshape2)
long_m <- acast(long_1, position~guilds, value.var="value")

#set manually the divisions, so far this is the best way I have seen to split the rows by categories

r_split <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 4),
             rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 3),rep(c("Asymmetric complete"), 2),
             "Complete", "Fan")

#set order of labels that avoids overlapping
r_split <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

#plat a bit with colors, a bit of a more complex set but looks nicer
library(circlize)
col_fun = colorRamp2(c(0.05644663,0.25, 0.4673256,0.75, 0.8782047), c("blue","cadetblue3", "azure2", "gold","red"))

h_pol <- Heatmap(long_m, name = "Mean\npercentile", row_split=r_split,cluster_row_slices = FALSE, 
                 column_title = "Pollinator functional groups",row_names_gp = gpar(fontsize = 10),
                 col = col_fun,column_names_rot = 45,  column_names_gp = grid::gpar(fontsize = 10),
                 column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                 row_title_gp = gpar(fontsize = 10, fontface = "bold"))

 #Prepare plants

#long format
long_plants <- melt(setDT(heat_plants), id.vars = c("position","Broad_categories"), variable.name = "Functional_groups")

#filter for creating matrix
long_plants_1 <- long_plants %>% select(-Broad_categories)

#set better colnames
long_plants_1$Functional_groups <- as.character(long_plants_1$Functional_groups)

long_plants_1$Functional_groups[long_plants_1$Functional_groups==1] <- "Selfing herbs" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==2] <- "Small outcrossing\nperennials" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==3] <- "Self-incomp. perennials\nwith large flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==4] <- "Tall plants with small\nunisexual flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==5] <- "Short-lived outcrossers with\nlong zygomorphic flowers"

long_plants_1$Functional_groups <- factor(long_plants_1$Functional_groups, levels=c("Selfing herbs",
"Small outcrossing\nperennials" ,"Self-incomp. perennials\nwith large flowers",
"Tall plants with small\nunisexual flowers" ,"Short-lived outcrossers with\nlong zygomorphic flowers" ))

#create matrix with reshape
long_plants_m <- acast(long_plants_1, position~Functional_groups, value.var="value")

r_split_plants <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 3),
                    rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 4),rep(c("Asymmetric complete"), 2),
                    "Complete", "Fan")

r_split_plants <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

h_plan <- Heatmap(long_plants_m, name = "Mean\npercentile", row_split=r_split_plants,cluster_row_slices = F, 
                  column_title = "Plant functional groups",row_names_gp = gpar(fontsize = 10),col = col_fun,column_names_rot = 45,  
                  column_names_gp = grid::gpar(fontsize = 10),column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                  row_title_gp = gpar(fontsize = 10, fontface = "bold"))

library(patchwork)
ht_list <- h_pol+h_plan

draw(ht_list, auto_adjust = FALSE, cluster_rows = F)

```