### 2021-10-05 Draft/Draft.Rmd
Update of abstract

---
title: Plant-pollintor networks worldwide are composed by the same specific building
  blocks
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{caption,setspace}
- \captionsetup[figure]{labelformat=empty}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}

output:
  pdf_document: default
  html_document: default
  word_document: default
always_allow_html: true
bibliography: bibliography.bib
---

__Jose B. Lanuza^1,2^__  | __Alfonso Allen-Perkins^2,3^__ |  __Romina Rader^1^__ | __Ignasi Bartomeus^2^__

1 School of Environmental and Rural Science, University of New England, Armidale, New South Wales 2350, Australia.

2 Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain.

3 Departamento de Ingeniería Eléctrica, Electrónica, Automática y Física Aplicada, ETSIDI, Universidad Politécnica de Madrid, 28040 Madrid, Spain.

\captionsetup[table]{labelformat=empty}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Abstract

Ecological processes leave distinct structural imprints on species interactions shaping the topology of mutualistic networks. Detecting those relationships is not trivial since they go beyond pair-wise interactions, but may get blurred when considering full network descriptors. However, recent work has shown the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of interactions between a small number of species (i.e. motifs) and those constitute the building blocks of the whole network. Here, we have compiled 60 networks from 18 different studies and show that some motifs are consitently under- and over-represented worldwide, suggesting that the building blocks of plant-pollintor networks are not random and are associated depending on the motif position and trophic level to specific fucntional groups. For instance, bees are the predominant group found on all motif positions but for plants there are three functional groups associated with corss-pollination that are similarly found on all motif positions. Hence, the distribution of funtional groups on the different positions is not random either which determine in last instance different probabilities of the overall functional group combinations of motifs.

\doublespacing

##INTRODUCTION

The interaction between plants and pollinators can be studied at different scales, from species level interactions (micro-scale) to the full network structure (macro-scale). Research of the plant-pollinator network structure have proven common invariant structural properties across them including a degree distribution that decays as a power law [@jordano1987], nestedness [@bascompte2003], or modularity in large networks [@olesen2007]. In addition, species phenological overlap, morphological matching and species abundances have been shown to be determinant for the understanding of pairwise plant-pollinator interactions [@bartomeus2016; @stang2006; @peralta2020]. However, both species level information and the holistic view of the full network involves missing relevant information for the understanding of ecological processes [@cirtwill2018; @simmons2019].

Traditionally, plant-pollinator research has focused on direct interactions but overlooked indirect interactions (i.e., the mediated effect between two species by a third species), such as facilitative or competitive interactions between plants for pollinators [@moeller2004; @sargent2008]. Despite the widespread nature of indirect interactions in ecological communities [@strauss1991], plant-pollinator research often fails to finely capture those indirect interactions with the conventional analytical tools that condense the information either by species (e.g., interaction frequency) or in single topological indices (e.g., nestedness). Nonetheless, the emerging framework of network motifs in plant-pollinator research, the building blocks of a network that depict subsets of interactions [@milo2002], allows to consider both direct and indirect interactions [@simmons2019]. The analysis of motifs on this intermediate level of complexity (meso-scale) in plant-pollinator networks have revealed that the different ecological processes that govern species interactions (e.g., species abundances versus trait-matching) can lead to different patterns of indirect interactions [@simmons2020]. Yet, the global patterns of indirect interactions in plant-pollinator networks are still unknown (e.g., over- and under- represented motifs).

Motifs are abstract representations often decoupled from species ecology. Thus, linking the structural properties of the meso-scale with the species ecology can help the understanding of ecological processes. For instance, different motifs can have different ecological meanings [@simmons2019] and the position within a motif can determine the species functional role [@stouffer2012; @baker2015]. However, it is unclear how the species ecology and life history traits determines the species functional role within the network of interactions [@coux2016]. For example, large pollinators can forage larger distances [@greenleaf2007], deposit greater pollen quantities [@foldesi2021] and handle complex zygomorphic flowers in comparison with small pollinators that are restricted to lower floral complexity [@gong2009]. How this different pollinator behaviors translate into their interaction topology is unknown. Similarly, recent empirical findings indicate that the meso-scale is the best descriptor of plant reproductive success [@allen2021], but little is known on how plants reproductive strategies shape their position within the network of interactions. Although some studies have evaluated plant reproductive strategies in plant-pollinator networks [@tur2013; @lazaro2020], they are often overlooked in a community context [@devaux2014] and rarely incorporated into plant-pollinator network studies. Hence, exploring how the main plant reproductive strategies integrate with the emergent motif framework can shed light on key aspects of ecosystem functioning.

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks. Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) the overall motif patterns (over and under-represented motifs); (ii) over and under-represented functional groups in the different motif positions; (iii) the probability of finding plant and floral visitor functional groups in the different motif positions; and, (iv) over- and under- represented functional group combinations within the different motifs.

---
# THIS IS A COMMENT
# Each os this points (i to iv) corresponds to a figure
# I think that ideally we can have that structure on methods and results (excluding functional groups)
# so far we have (i), (ii) and (iv) on methods
# we lack an explicit mention to (iii) 
# maybe merge it with another research question and just have 3?
# What do you think? 
---

##METHODS

**Plant-pollinator studies**

We have compiled 60 plant-pollinator networks from 18 different studies (**Table S1**). All studies sampled plant-pollinator interactions in natural systems and were selected based on wide geographical coverage and presence of interaction frequency as a measure of interaction strength. In total, there were 503 plant species, 1,111 floral visitors species and 6248 pairwise interactions registered. For ease of data manipulation, plant and floral visitors species names were standardize with the help of the package _taxize_ version 0.9.99 [@chamberlain2020].

**Plant and floral visitors functional groups**

First, plant species were grouped through hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. We used the trait dataset collated in Lanuza et al., (unpublished) that comprised 1,506 plant species and contained the same 60 weighted plant-pollinator networks used in this study plus 4 additional large unweighted metawebs (see **Table S1** Lanuza et al., unpublished). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30%; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. Previous to the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because the main orders of floral visitors differed in form and behaviour and had lower superior taxonomical complexity (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies of the possible pollinators: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi) Coleoptera. However, a minor set of species belonged to other functional groups that were considered in analyses but not discussed further because of their low representation (3.55% of the total interactions recorded in the full set of networks). These functional groups were 'lizards', 'birds' and 'other insects'. This last group was formed by a mix of uncommon insect taxa on the full set of networks.

**Overall motif patterns**

Following previous work [@simmons2019; @simmons2020], we broke down the plant-pollinator networks into their constituent motifs. Prior to analyses, we turned the quantitative networks into qualitative (or binary) ones, where interactions are present or absent.

We calculated the frequency of all motifs up to five nodes (see **Figure 1**) for each empirical network, by using the _bmotif_ package [@bmotifsimmons2019]. To control for variation in network size, motif frequencies were normalised as a proportion of the total number of motifs within each motif class (i.e., the number of nodes a motif contains). Like @simmons2020, we just used five-node motifs in our analyses for visualisation, interpretation and computational reasons. In addition, we also excluded two-node motifs (or links) from our analyses because their normalised frequencies would always equal one.

```{r, echo=FALSE, message=FALSE, cache=FALSE, fig.pos="t", warning=FALSE,out.width="75%",fig.align="center", fig.cap="\\textbf{Figure 1.} Adapted figure of Simmons et al., 2019 with all the possible motifs from two to five species in bipartite networks. There is a total of 17 possible motifs with 46 different positions denoted within each node.", fig.width=10,fig.height=8}
library(tidyverse)
library(RColorBrewer)
source("../Scripts/Scripts_Alfonso/plot_motif_positionsV2.R")

motif_number <- 1
p1 <- plot_motif_positions(motif_number)

motif_number <- 2
p2 <- plot_motif_positions(motif_number)

motif_number <- 3
p3 <- plot_motif_positions(motif_number)

motif_number <- 4
p4 <- plot_motif_positions(motif_number)

motif_number <- 5
p5 <- plot_motif_positions(motif_number)

motif_number <- 6
p6 <- plot_motif_positions(motif_number)

motif_number <- 7
p7 <- plot_motif_positions(motif_number)

motif_number <- 8   
p8 <- plot_motif_positions(motif_number)

motif_number <- 9
p9 <- plot_motif_positions(motif_number)

motif_number <- 10
p10 <- plot_motif_positions(motif_number)

motif_number <- 11
p11 <- plot_motif_positions(motif_number)

motif_number <- 12
p12 <- plot_motif_positions(motif_number)

motif_number <- 13
p13 <- plot_motif_positions(motif_number)

motif_number <- 14
p14 <- plot_motif_positions(motif_number)

motif_number <- 15
p15 <- plot_motif_positions(motif_number)

motif_number <- 16
p16 <- plot_motif_positions(motif_number)

motif_number <- 17
p17 <- plot_motif_positions(motif_number)

library(patchwork)

(p1 & ylab("2 species")| p2 & ylab("3 species") | p3 )/ (p4 & ylab("4 species")| p5 | p6 | p7)/ (p8  & ylab("5 species")| p9 | p10 | p11 | p12) / (p13 & ylab("5 species")| p14 | p15 | p16 | p17) & theme(axis.title.y = element_text(color="black", size=14, face="bold")) #+
 #plot_annotation(tag_levels = '1')

```

To assess the significance of the observed frequencies, we created 1,000 simulated networks for each binary network using the _nullmodel_ function and the _vaznull_ model in the _bipartite_ package [@dormann2009]. Generated networks had the same number of plants and floral visitors, as well as the same connectance of their corresponding empirical networks. After extracting the motif frequencies from the simulated networks, for each motif type and empirical network, we calculated the percentage of simulated networks whose frequencies were smaller than the ones observed, that is, we estimated the percentile of the observed motif frequencies. Motifs whose percentile is close to 0 or 100 are under- or over-represented in the empirical networks, respectively, and thus they cannot be predicted by connectance and the number of species alone. To summarize general patterns across networks, we used an intercept-only linear mixed model (LMM) per motif, where the response variable was the observed motif percentile per network. In these models, we used the study identifiers in **Table S1** as a random intercept. By doing so, we obtained estimates of the average motif frequency, in which we controlled the variation at the study level.

**Over and under-represented functional groups on motif positions**

We calculated which functional groups were over or under-represented in different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. We extracted the position frequencies of all motifs from three to five nodes for each network and species by using the _bmotif_ package [@bmotifsimmons2019]. To estimate the position frequencies of each functional group in a given network, we added the frequencies of those species that belong to the group, and then, we normalised the resulting frequencies by dividing the position measure [Sum of position frequencies?]() for each group by the total number of times that a group appears in any position within the same motif size class. Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did with motif frequencies. To outline the general patterns of position frequencies across networks and functional groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the functional group identifier as an explanatory variable and the study identifiers as a random intercept. By adding the functional group estimates to the model intercept, we assessed the average motif frequency, after controlling the variation at the study level. Finally, we visualized with the help of the package ComplexHeatmap version 2.6.2 [@gu2016] over- and under- representation of plant and pollinator functional groups on the different motif positions.

To test the effect of singletones on the motif frequencies and motif functional groups' positions, we applied the previous analyses to the same set of networks (**Table S1**) but considering only interactions whose frequency was greater than one (64.98% of interactions; see **Figures S1 and S2**).

**Over- and under-represented funtional group combinations of motifs**

Finally, we studied which motif combinations of functional groups (up to five nodes) are over or under-represented in 95% of our empirical networks (57 out of 60 arrangements, due to computational limitations to identify all the nodes in the motifs of the three networks with the highest number of links). To do so, for each of the 53,250 possible motif combinations, we estimated the observed and the expected probability of finding those combination in empirical networks, respectively. Then, we determined whether the observed probabilities are likely to come from the expected probabilities or not. To calculate the observed probability ($p_i^O$) of functional groups within a motif ($i$; e.g., motif 3, bee + bee + functional group 2), we divided the number of times that $i$ appears in our set of empirical networks ($n_i^O$) by the sum of the number of times that each possible combination appeared: $p_i^O = n_i^O / \sum_{k=1}^{53,250} n_k^O$. To estimate the expected probability of a given motif combination $i$, $p_i^E$, firstly, we calculated the probability of finding a given functional group $x$ in the position $\alpha$ of $i$, $p_i(x,\alpha)$. Then, by assuming the independency of $p_i(x,\alpha)$, we computed the expected probability of the combination $i$ as the product of the probability of its pairs $(x,\alpha)$, that is, $p_i^E = \prod _{(x,\alpha )\in i} p_i(x,\alpha)$. To obtain $p_i(x,\alpha)$, we proceeded as follows. First, we calculated the number of times that the functional group $x$ appears in the position $\alpha$ in each empirical network $\eta$, denoted as $n_i^{\eta}(x,\alpha)$. Then, by controlling the variation at the network level, we assessed the average value of the absolute frequency of functional group $x$ appears in the position $\alpha$, $n_i(x,\alpha) = E\left [ n_i^{\eta}(x,\alpha) \right ]$, and estimated $p_i(x,\alpha)$ as $p_i(x,\alpha) = n_i(x,\alpha)/\sum_{k}n_i(k,\alpha)$. To obtain $n_i(x,\alpha)$, we fitted a LMM per motif position, where the response variable was the number of times that a given position was observed per network, the explanatory variable was functional group identifier, and the random intercept was given by network identifiers nested within the study identifiers. We used the ‘lmer’ package [@bates2015] to fit the LMMs models in our analyses.

Once we obtained $p_i^O$ and $p_i^E$, we used simulation to determine whether the former is likely to come from the latter or not, since the large number of possible motif combinations and the small probabilities for some of them advise against using an exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contained 10 million elements and, for each combination, the probability of being selected was equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, denoted as $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. According to the usual interpretation of z-scores, combinations with $z_i^O > 1.96$ are over-represented, whereas those with $z_i^O < -1.96$ are under-represented, at the 95%  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).

##RESULTS

**Plant and floral visitor functional groups**

The hierarchical cluster analysis divided the dataset into five different clusters with different and overlapping characteristics (**Figure S3** and **Figure S4**). Cluster 1 referred to as 'selfing herbs' consisted on herbs with hermaphroditic flowers with high levels of autonomous selfing. Cluster 2 or 'small outcrossing perennials' had small perennial species with a mixed of life forms (i.e., trees shrubs and herbs) with outcrossing hermaphroditic flowers. Cluster 3 referred to as 'self-incompatible perennials with large flowers' comprised perennial species with a mixed of life forms and large self-incompatible hermaphroditic flowers with high number of ovules. Cluster 4 or 'tall plants with small unisexual flowers' had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding systems, small flowers and the highest numbers of flowers and inflorescences per plant. Finally, cluster 5 or 'short-lived outcrossers with long zygomorphic flowers' consisted on small perennial and short-lived herbs with long self-compatible zygomorphic flowers that were unable to self-pollinate.

In total, there were 1126 species of floral visitor with 6325 interactions recorded with plants. There were 2256 bee interactions with plants, 1768 non-syrphid Diptera interactions, 845  syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee Hymenoptera interactions.

**Overall motif patterns**

Most motifs were under- and over-represented (close to the 1st and 99th percentile, respectively) in the comparison between empirical and simulated networks (**Figure 2**). Motifs 3, 5, 9, 10 and 14 were under-represented in empirical networks, that is, all were close to the 1st percentile and under the 25th percentile. Further, four out of five of these motifs belonged to the largest path length classification (i.e., core-peripheral). In addition, motifs 2, 6, 7, 16 and 17 were over-represented, all over the 75th percentile and thus the closest to the 99th percentile. In contrast to the under-represented motifs, over-represented motifs belonged to the two shortest path length groups (i.e., complete and fan). The remaining motifs (i.e., 4, 8, 11, 12, 13, and 15) were between the 25th and 75th percentile but closest to over-representation. Notably, the exclusion of singletones from this analysis did not change the observed results (**Figure S1**).

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=4.5,fig.height=6.5, fig.cap="**Figure 2.** Comparison of motif frequencies between empirical and simulated networks grouped by average path length (plots a, b, c and d) as determined in Simmons et al. (2020). This is shown with the mean percentage of motif frequencies in empirical networks that were over the motif frequencies of the simulated ones (percentiles). This was done by network (light blue dots) and then averaged for all networks (black dots with error bars that correspond to the standard deviation)."}
library(lme4)
library(nlme)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models/Motifs_frequency_percentile.csv")
d <- add_study_id(d)
#str(d)
#head(d)

#ggplot(d %>% filter(motif != 1), aes(percentil_sizeclass))+
#  geom_histogram(color="black", fill="white")+
#  facet_wrap(~motif)+
#  theme_bw()+
#  labs(x="Sizeclass percentile")
#
# Motifs tend to be over- and under-represensented in real networks

# Mean percentile and SE taking into account the study system
motif_codes <- d %>% filter(motif != 1) %>% # We remove links (motif code: 1)
  select(motif) %>% unique() %>% pull()
motif_means <- tibble(motif = motif_codes)
motif_means$mean <- NA
motif_means$SE <- NA

for(i.motif in 1:length(motif_codes)){

  m <- lmer(percentil_sizeclass ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(d, motif == motif_codes[i.motif]))

  motif_means$mean[i.motif] <- fixed.effects(m) %>% unname()
  motif_means$SE[i.motif] <- sqrt(diag(vcov(m)))

}

#Add categories based on path length set by Simmons 2021 et al., Functional Ecology
d$Broad_categories[d$motif==2 | d$motif==7 | d$motif==17 | d$motif==3|d$motif==4|d$motif==8] <- "Fan"

d$Broad_categories[d$motif==15 | d$motif==11] <- "Medium-weak"

d$Broad_categories[d$motif==6 | d$motif==16 |  d$motif==12] <- "Strong"

d$Broad_categories[d$motif==5 | d$motif==9 |  d$motif==14|  d$motif==13|  d$motif==10] <- "Weak"

#order the categories from short to long path length
d$Broad_categories <- factor(d$Broad_categories, levels = c("Strong", "Fan", "Medium-weak", "Weak"))

#Ffilter out two node motifs
d_1 <- d %>% filter(motif != 1)
#Merge dataframes to bl able to create facets
d_2 <- merge(d_1, motif_means, by="motif")

arr <- list('Strong' = expression(bold(paste("(a) Complete (strong, ", bar(x), " = 1.38)"))),
             'Fan' = expression(bold(paste("(b) Fan (medium-strong, ", bar(x), " = 1.48)"))), 
             'Medium-weak' = expression(bold(paste("(c) Asymmetric complete (medium-weak, ", bar(x), " = 1.60)"))), 
             "Weak" =expression(bold(paste("(d) Core-peripheral (weak, ", bar(x), " = 1.85)"))))
mylabel <- function(val) { return(lapply(val, function(x) arr[x])) }

library(ggforce) #workaround with ggforce to have facets with same sizes

p <- ggplot()+geom_point(data = d_2, aes(y=factor(motif), 100*percentil_sizeclass), shape=21,color="black",fill="azure2",size=1.75, alpha=0.3)+
  geom_errorbar(data = d_2,aes(y = factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=0.8,size=0.8,color="black") +theme_bw()+
  geom_point(data = d_2,aes(y = as.factor(motif), x=100*mean), 
             size=2,color="black")+labs(y="Two to five node motifs", x = "Percentile")+
  theme( strip.background = element_blank())+
  theme(strip.text = element_text(face="bold", size=10, hjust=0),
        panel.border=element_rect(color="black",size=1.2))

p + ggforce::facet_col(vars(Broad_categories), scales = 'free', space = 'free',labeller=mylabel)

#Trying to maintain a gradient from low to high in colors

#Fig 1 Should be a dot plot (or forest plot) of the mean +- SE of the 15 motifs.
#NICE! Can we put motifs in Y axes, and % in X?
```

**Over and under-represented functional groups on motif positions**

The comparison of the frequencies between the positions of motifs of empirical and simulated networks showed that motifs positions determined over- and under- representation independently of the identity of functional groups for most cases (rows in **Figure 3**). However, we also found differences in the average percentile within floral visitors and plant functional groups on the different motif positions (columns in **Figure 3**). These differences were more marked for floral visitors functional groups, and thus the groupings by average path length were less informative for floral visitors than for plants. In other words, for plants motifs by themselves were more descriptive than for floral visitors. From most over- to under-represented floral visitors functional groups, we found: bees, non-syrphids Diptera, syrphids, Coleoptera, non-bee hymenoptera and Lepidoptera. Although plant functional groups showed less differences between them, there were also more represented functional groups than others, thus from most over- to under-represented groups on the different motif positions we found: self incompatible perennials with large flowers, small outcrossing perennials, tall plants with unisexual flowers, selfing herbs and short lived outcrossers with long zygomorphic flowers.

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=8,fig.width=8,fig.align = 'center', out.width="90%",fig.cap="\\textbf{Figure 3.} Heatmap indicating under- and over- representation of pollinator and plant functional groups in the different motif positions. The different motif positions are dividied by the average path length clasification determined by Simmons et al. (2020)."}

library(lme4)
library(nlme)
library(tidyverse)
library(stringi)
source("../Scripts/Scripts_Alfonso/add_study_id.R")

#############################################
# EXTRACT MEAN VALUES
#############################################

all_position_percentiles <- read_csv("../Data/Csv/Motifs_positions_and_null_models/GF_positions_frequency_percentile.csv")

all_position_percentiles$position <- stri_extract_first_regex(all_position_percentiles$position,
                                                              "[0-9]+") %>% as.numeric()

all_position_percentiles <- add_study_id(all_position_percentiles)
#str(all_position_percentiles)
#head(all_position_percentiles)

list_plants_FG <- as.character(1:10)

plant_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF)) #ME DA ERROR
pollinator_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(!Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF))

# Plants-------------

# Mean percentile and SE taking into account the study system
plant_codes <- plant_position_percentiles_filtered$position %>% unique()
plant_means <- plant_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
plant_means$mean <- NA
plant_means$SE <- NA

for(i.pos in 1:length(plant_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id),
            data = plant_position_percentiles_filtered %>%
              filter(position == plant_codes[i.pos]))

  temp <- fixed.effects(m) %>% unname()
  plant_means$mean[plant_means$position == plant_codes[i.pos]] <- c(temp[1],
                                                                    temp[1]+temp[2],
                                                                    temp[1]+temp[3],
                                                                    temp[1]+temp[4],
                                                                    temp[1]+temp[5]) ## to get real means per FG
  temp <- sqrt(diag(vcov(m)))
  plant_means$SE[plant_means$position == plant_codes[i.pos]] <- c(temp[1], 
                                                                  temp[1]+temp[2],
                                                                  temp[1]+temp[3],
                                                                  temp[1]+temp[4],
                                                                  temp[1]+temp[5])

}

# Pollinators--------

# Mean percentile and SE taking into account the study system
pollinator_codes <- pollinator_position_percentiles_filtered$position %>% unique()
pollinator_means <- pollinator_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
pollinator_means$mean <- NA
pollinator_means$SE <- NA
pollinator_means_reordered <- NULL

for(i.pos in 1:length(pollinator_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = pollinator_position_percentiles_filtered %>%
              filter(position == pollinator_codes[i.pos]),
            control = lmerControl(optimizer ="Nelder_Mead")) # I changed the optimizer because motif 8 (i.pos = 3) caused convergence problems, max gradient = 0.0022 > 0.002.

  temp <- fixed.effects(m) %>% unname()
  pollinator_means_aux <- pollinator_means[pollinator_means$position == pollinator_codes[i.pos],] %>% arrange(Node_FG)

  pollinator_means_aux$mean <- c(temp[1],
                                 temp[1]+temp[2],
                                 temp[1]+temp[3],
                                 temp[1]+temp[4],
                                 temp[1]+temp[5],
                                 temp[1]+temp[6],
                                 temp[1]+temp[7],
                                 temp[1]+temp[8],
                                 temp[1]+temp[9]) ## to get real means per FG

  temp <- sqrt(diag(vcov(m)))
  pollinator_means_aux$SE <- c(temp[1],
                               temp[1]+temp[2],
                               temp[1]+temp[3],
                               temp[1]+temp[4],
                               temp[1]+temp[5],
                               temp[1]+temp[6],
                               temp[1]+temp[7],
                               temp[1]+temp[8],
                               temp[1]+temp[9])

  pollinator_means_reordered <- bind_rows(pollinator_means_reordered,pollinator_means_aux)

}

# There are lower limits (mean - SE) smaller than 0 (see Other insects)
# We set the lower limit of those error bars to zero

pollinator_means_reordered <- pollinator_means_reordered %>% mutate(lower = mean-SE)
pollinator_means_reordered$lower[pollinator_means_reordered$lower < 0] <- 0

#############################################
# ADDING BROAD MOTIF CATEGORIES TO POSITIONS
#############################################

motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

motifs_raw_data$Broad_categories <- NA

motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(5,9,14,13,10)] <- "Core-peripherical"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(6,16,12)] <- "Complete"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(2,3,7,4,17,8)] <- "Fan"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(15,11)] <- "Asymmetric complete"

motifs_raw_data$plant <- gsub("[^0-9.-]", "", motifs_raw_data$plant)
motifs_raw_data$pollinator <- gsub("[^0-9.-]", "", motifs_raw_data$pollinator)
motifs_raw_data$plant <- as.numeric(motifs_raw_data$plant)
motifs_raw_data$pollinator <- as.numeric(motifs_raw_data$pollinator)
motifs_raw_data <- motifs_raw_data %>% filter(!is.na(Broad_categories))

fan_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

pollinator_means_reordered$Broad_categories <- NA

pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% fan_pollinator] <- "Fan" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% asymm_pollinator] <- "Medium-weak" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% complete_pollinator] <- "Strong" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% core_per_pollinator] <- "Weak"

fan_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

plant_means$Broad_categories <- NA

plant_means$Broad_categories[plant_means$position %in% fan_plant] <- "Fan" 
plant_means$Broad_categories[plant_means$position %in% asymm_plant] <- "Medium-weak" 
plant_means$Broad_categories[plant_means$position %in% complete_plant] <- "Strong" 
plant_means$Broad_categories[plant_means$position %in% core_per_plant] <- "Weak"

########################################
# CREATE HEATMAPS
########################################

library(ComplexHeatmap)

# PLANTS-----

# Prepare data for heatmap
heat_plants_aux <- plant_means %>% spread(Node_FG,mean) %>% select(-SE)

heat_plants <- heat_plants_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_plants_fan <- heat_plants %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_fan) <- heat_plants_fan[,1]

heat_plants_Weak <- heat_plants %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Weak) <- heat_plants_Weak[,1]

heat_plants_Strong <- heat_plants %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Strong) <- heat_plants_Strong[,1]

heat_plants_Medium_weak <- heat_plants %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Medium_weak) <- heat_plants_Medium_weak[,1]

# POLLINATORS-----------

filter_rare_groups <- T

if(filter_rare_groups == T){
  heat_pollinators_aux <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds",
                                                                                "Lizards",
                                                                                "Other_insects")) %>% 
    spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee",            
                                      "Coleoptera","Lepidoptera","Non-bee\nHymenoptera", 
                                      "Non-syrphids\nDiptera","Syrphids")
}else{
  heat_pollinators_aux <- pollinator_means_reordered %>% spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee","Birds",            
                                      "Coleoptera","Lepidoptera","Lizards","Non-bee\nHymenoptera", 
                                      "Non-syrphids\ndiptera","Other insects","Syrphids")
}

heat_pollinators <- heat_pollinators_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_pollinators_fan <- heat_pollinators %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_fan) <- heat_pollinators_fan[,1]

heat_pollinators_Weak <- heat_pollinators %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Weak) <- heat_pollinators_Weak[,1]

heat_pollinators_Strong <- heat_pollinators %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Strong) <- heat_pollinators_Strong[,1]

heat_pollinators_Medium_weak <- heat_pollinators %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Medium_weak) <- heat_pollinators_Medium_weak[,1]

##################

#Do not know how to handle this multiplot for creating a panel

# I'm going to create the 4 plots in one, 
# I'm not sure that the grouping of the functional group is correct
# when is done separately

#sorry for the mess of libraries that I do below

#Prepare first pollinators

library(tidyr)
library(data.table)
long <- melt(setDT(heat_pollinators), id.vars = c("position","Broad_categories"), variable.name = "guilds")

long_1 <- long %>% select(-Broad_categories)

library(reshape2)
long_m <- acast(long_1, position~guilds, value.var="value")

#set manually the divisions, so far this is the best way I have seen to split the rows by categories

r_split <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 4),
             rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 3),rep(c("Asymmetric complete"), 2),
             "Complete", "Fan")

#set order of labels that avoids overlapping
r_split <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

#plat a bit with colors, a bit of a more complex set but looks nicer
library(circlize)
col_fun = colorRamp2(c(0.05644663,0.25, 0.4673256,0.75, 0.8782047), c("blue","cadetblue3", "azure2", "gold","red"))

h_pol <- Heatmap(long_m, name = "Mean\npercentile", row_split=r_split,cluster_row_slices = FALSE, 
                 column_title = "Pollinator functional groups",row_names_gp = gpar(fontsize = 10),
                 col = col_fun,column_names_rot = 45,  column_names_gp = grid::gpar(fontsize = 10),
                 column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                 row_title_gp = gpar(fontsize = 10, fontface = "bold"))

#Prepare plants

#long format
long_plants <- melt(setDT(heat_plants), id.vars = c("position","Broad_categories"), variable.name = "Functional_groups")

#filter for creating matrix
long_plants_1 <- long_plants %>% select(-Broad_categories)

#set better colnames
long_plants_1$Functional_groups <- as.character(long_plants_1$Functional_groups)

long_plants_1$Functional_groups[long_plants_1$Functional_groups==1] <- "Selfing herbs" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==2] <- "Small outcrossing\nperennials" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==3] <- "Self-incomp. perennials\nwith large flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==4] <- "Tall plants with small\nunisexual flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==5] <- "Short-lived outcrossers with\nlong zygomorphic flowers"

long_plants_1$Functional_groups <- factor(long_plants_1$Functional_groups, levels=c("Selfing herbs",
"Small outcrossing\nperennials" ,"Self-incomp. perennials\nwith large flowers",
"Tall plants with small\nunisexual flowers" ,"Short-lived outcrossers with\nlong zygomorphic flowers" ))

#create matrix with reshape
long_plants_m <- acast(long_plants_1, position~Functional_groups, value.var="value")

r_split_plants <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 3),
                    rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 4),rep(c("Asymmetric complete"), 2),
                    "Complete", "Fan")

r_split_plants <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

h_plan <- Heatmap(long_plants_m, name = "Mean\npercentile", row_split=r_split_plants,cluster_row_slices = F, 
                  column_title = "Plant functional groups",row_names_gp = gpar(fontsize = 10),col = col_fun,column_names_rot = 45,  
                  column_names_gp = grid::gpar(fontsize = 10),column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                  row_title_gp = gpar(fontsize = 10, fontface = "bold"))

library(patchwork)
ht_list <- h_pol+h_plan

draw(ht_list, auto_adjust = FALSE, cluster_rows = F)

```

The different observed frequencies on the different motif positions showed a hirarchical order of probabilities for floral visitors (**Figure 4**). That is, the functional group of 'bees' was the most abundant one on all motif positions followed by non-syrphid Diptera which was after bees, the second most frequent functional group on all positions. Then, the group of syrphids was the most abundant in all positions but two (positions 28 and 31 from motifs 11 and 12, respectively). After syrphids, Coleoptera was the most frequent group on the different positions followed by non-bee Hymenoptera and finally, Lepidoptera which was the less frequent group on all positions but two (positions 12 and 22 from motifs 5 and 9, respectively). Remarkably, plant functional groups were more variable on the different motif positions. However, there were also predominant functional groups found across the different motif positions. The three most frequent groups were 'tall plants with small unisexual flowers', 'self-incompatible perennials with large flowers' and 'small outcrossing perennials'. On the contrary, the groups of 'selfing herbs' and 'short lived outcrossers with long zygomorphic flowers' had the lowest probability to be present on the different motifs plant positions.

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=8, fig.width=12, fig.cap="**Figure 4.** Graphical representation of the probability of finding a given functional group $x$ in the position $\\alpha$ of motif $i$, $p_i(x,\\alpha)$, for all the possible motifs from two to five species in bipartite networks. The slices in the nodes for a given functional group $x$ are proportional to the corresponding value of $p_i(x,\\alpha)$."}

library(ggplot2)
library(scatterpie)
library(RColorBrewer)
library(tidyverse)

plant_means_reordered <- read_csv("../Data/Csv/plant_abs_freq_means.csv")
pollinator_means_reordered <- read_csv("../Data/Csv/pollinator_abs_freq_means.csv")
#Filter out in order to have just 6 FG'S
pollinator_means_reordered <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds", "Lizards", "Other_insects"))

# Pollinator--------
pollinator_pies <- pollinator_means_reordered %>% select(-mean, -SE) %>% 
  spread(Node_FG,mean_natural_units)

pollinator_pies$x <- 0
pollinator_pies$y <- 0

#ggplot() + geom_scatterpie(aes(x=x, y=y, group=position,r=5), 
#                           data = pollinator_pies,
#                           cols=as.character(c("Bee","Birds","Coleoptera","Lepidoptera",
#                                               "Lizards","Non-bee-Hymenoptera",
#                                               "Non-syrphids-diptera","Other_insects",
#                                               "Syrphids")))+
#  coord_equal()+
#  scale_fill_brewer(palette = "Paired")+
#  facet_wrap(~position,nrow = 3)+
#  theme_bw()+
#    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#          panel.background = element_blank(),
#          axis.title.x=element_blank(),axis.text.x=element_blank(),
#          axis.ticks.x=element_blank(),axis.title.y=element_blank(),
#          axis.text.y=element_blank(),axis.ticks.y=element_blank(),
#          legend.position="bottom")+
#  labs(title = "Pollinator positions",fill=NULL)
#    
# plant--------
plant_pies <- plant_means_reordered %>% select(-mean, -SE) %>% 
  spread(Node_FG,mean_natural_units)

plant_pies$x <- 0
plant_pies$y <- 0

#ggplot() + geom_scatterpie(aes(x=x, y=y, group=position,r=5), 
#                           data = plant_pies,
#                           cols=as.character(c("1","2","3","4","5")))+
#  coord_equal()+
#  scale_fill_brewer(palette = "Dark2")+
#  facet_wrap(~position,nrow = 3)+
#  theme_bw()+
#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
#        axis.title.x=element_blank(),axis.text.x=element_blank(),
#        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
#        axis.text.y=element_blank(),axis.ticks.y=element_blank(),
#        legend.position="bottom")+
#  labs(title = "Plant positions",fill=NULL)
#

#################################
# CREATE PIE-MOTIFS
#################################

equidistribution <- function(number_nodes){

  x_width <- 5.5/(number_nodes+1)

  x_result <- rep(-1.5,number_nodes)

  for(i in 1:number_nodes){

    x_result[i] <- x_result[i] + i*x_width

  }

  return(x_result)

}

plot_pie_motif_positions <- function(motif_number){
  motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

  motif_data <- motifs_raw_data %>% filter(motif_id == motif_number)

  # Create dataframe with node positions

  labdown = motif_data$plant %>% unique()

  xdown = equidistribution(length(labdown))

  ydown = rep(0,length(labdown))

  labup = motif_data$pollinator %>% unique()

  xup = equidistribution(length(labup))

  yup = rep(2,length(labup))

  d_nodes=data.frame(x=c(xdown,xup),
                     y=c(ydown,yup),
                     label = c(labdown,labup))

  # Create a dataframe with links

  plant_positions <- data.frame(plant = labdown,
                                xdown = xdown,
                                ydown = ydown)

  pollinator_positions <- data.frame(pollinator = labup,
                                     xup = xup,
                                     yup = yup)

  link_df <- motif_data %>% left_join(plant_positions, by = "plant") %>% 
    left_join(pollinator_positions, by = "pollinator")

  # Before plotting the motif, we remove letters from the node position id

  d_nodes$label <- gsub("[^0-9.-]", "", d_nodes$label)
  plant_pies_aux <- plant_pies %>% select(-x,-y) %>% mutate(position = as.character(position))

  d_nodes_plant <- d_nodes %>% rename(position = label) %>% 
    left_join(plant_pies_aux, by = "position") %>% as_tibble() %>%
    filter(!is.na(`1`))

  pollinator_pies_aux <- pollinator_pies %>% select(-x,-y) %>% mutate(position = as.character(position))

  d_nodes_pollinator <- d_nodes %>% rename(position = label) %>% 
    left_join(pollinator_pies_aux, by = "position") %>% as_tibble() %>%
    filter(!is.na(Bee))

  # Define the number of colors you want
 mycolors <- c("#A6CEE3", "chocolate2",  "gold2", "#1F78B4","#B2DF8A",
                     "gray58", "lightgoldenrod2","thistle2", "burlywood","lightcoral", "olivedrab3")

  ggplot() +
    scale_shape_identity() +
    geom_segment(data = link_df, aes(x=xdown, y=ydown, xend = xup, yend = yup), 
                 size = 1, color = "grey20")+  geom_scatterpie(aes(x=x, y=y, group=position,r=.45), 
    data = d_nodes_pollinator, cols=as.character(c("Bee","Coleoptera","Lepidoptera","Non-bee-Hymenoptera", "Non-syrphids-diptera","Syrphids"))) +
    geom_scatterpie(aes(x=x, y=y, group=position,r=0.45), 
                    data = d_nodes_plant,
                    cols=as.character(c("1","2","3","4","5")))+
    coord_fixed()+
    scale_fill_manual(values = mycolors,"test",
      breaks=c("Bee","Coleoptera","Lepidoptera",
               "Non-bee-Hymenoptera", "Non-syrphids-diptera",
               "Syrphids","1","2","3","4","5"), labels=c("Bee","Coleoptera","Lepidoptera",
"Non-bee-Hymenoptera", "Non-syrphids-diptera","Syrphids", 
          "Selfing herbs","Small outcrossing perennials","Self-incomp. perennials\nwith large flowers",
          "Tall plants with small\nunisexual flowers","Short-lived outcrossers with\nlong zygomorphic flowers"))+
    xlim(-1.5,4)+ ylim(-0.5,2.5)+
    theme(panel.background = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          legend.title = element_text(size=12, face="bold"),legend.background = element_rect(size=0.5, linetype="solid",colour ="black"))+
    labs(x=NULL,y=NULL)+
    guides(color = FALSE,fill=guide_legend(title="Functional groups"))+
    ggtitle(paste0("Motif ",motif_number))+
    theme(plot.title = element_text(hjust = 0.5),legend.key.size = unit(0.5, 'cm'))+
    theme(plot.margin=unit(c(0,0,0,0),"cm"))
}

motif_number <- 1
p1 <- plot_pie_motif_positions(motif_number)

motif_number <- 2
p2 <- plot_pie_motif_positions(motif_number)

motif_number <- 3
p3 <- plot_pie_motif_positions(motif_number)

motif_number <- 4
p4 <- plot_pie_motif_positions(motif_number)

motif_number <- 5
p5 <- plot_pie_motif_positions(motif_number)

motif_number <- 6
p6 <- plot_pie_motif_positions(motif_number)

motif_number <- 7
p7 <- plot_pie_motif_positions(motif_number)

motif_number <- 8   
p8 <- plot_pie_motif_positions(motif_number)

motif_number <- 9
p9 <- plot_pie_motif_positions(motif_number)

motif_number <- 10
p10 <- plot_pie_motif_positions(motif_number)

motif_number <- 11
p11 <- plot_pie_motif_positions(motif_number)

motif_number <- 12
p12 <- plot_pie_motif_positions(motif_number)

motif_number <- 13
p13 <- plot_pie_motif_positions(motif_number)

motif_number <- 14
p14 <- plot_pie_motif_positions(motif_number)

motif_number <- 15
p15 <- plot_pie_motif_positions(motif_number)

motif_number <- 16
p16 <- plot_pie_motif_positions(motif_number)

motif_number <- 17
p17 <- plot_pie_motif_positions(motif_number)

library(patchwork)

(p1 & ylab("2 species")| p2 & ylab("3 species") | p3  |plot_spacer()|plot_spacer())/ 
  (p4 & ylab("4 species")| p5 | p6 | p7|plot_spacer())/ 
  (p8  & ylab("5 species")| p9 | p10 | p11 | p12) / 
  (p13 & ylab("5 species")| p14 | p15 | p16 | p17) + plot_layout(guides = 'collect') &
  theme(axis.title.y = element_text(color="black", size=14, face="bold"),legend.position = "bottom")

```

\newpage

**Over- and under-represented funtional group combinations of motifs**

The statistical comparison with z-scores between the observed functional group combinations and the expected ones showed a modified Gaussian distribution (Johnson's $S_U$) with 65% of under-represented group combinations, 20% of no statistical difference and 15% of under-represented ones (**Figure 5**). Remarkably, motifs with small node combinations (i.e., 2 and 3 nodes) only appeared as under-represented. Motifs with 4 and 5 node combinations appeared in the three statistical categories but 4 node motifs had a higher porportion in the under-represented category (10% in comparison with the 1% and the 3% of the no statiscal difference and over-represented categories, respectively). Finally, 5 node motifs were in similar prportions in the no statistical difference and over-represented categories (99% and 97%, respectively) but they appeared in lower porportion in the under-represented category (89%). When, we explore the identity on the different positions of the most probable motif combinations we find as expected that the most common functional groups rule these positions on the most probable motif combiantions (lower panel **Figure 5**).

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=8, fig.width=12, fig.cap="**Figure 5.** Z-score comparison (95% confidence level) of the observed and expected frequencies motif combinations with four specific graphical examples of the most probable motifs of being observed for each category delimited by the z-score critical values: under-represented, no statistical difference and over-represented."}

library(tidyverse)
library(RColorBrewer)

equidistribution <- function(number_nodes){

  x_width <- 5.5/(number_nodes+1)

  x_result <- rep(-1.5,number_nodes)

  for(i in 1:number_nodes){

    x_result[i] <- x_result[i] + i*x_width

  }

  return(x_result)

}

plot_motif_label <- function(label){

  label_split <- str_split(label,"_")[[1]]
  motif_number <- as.numeric(label_split[1])
  nodes <- label_split[2:length(label_split)]
  plant_nodes <- nodes[nodes %in% plant_means_reordered$Node_FG]
  pollinator_nodes <- nodes[nodes %in% pollinator_means_reordered$Node_FG]

  motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

  motif_data <- motifs_raw_data %>% filter(motif_id == motif_number)

  # Create dataframe with node positions

  labdown = motif_data$plant %>% unique()

  xdown = equidistribution(length(labdown))

  ydown = rep(0,length(labdown))

  labup = motif_data$pollinator %>% unique()

  xup = equidistribution(length(labup))

  yup = rep(1,length(labup))

  d_nodes=data.frame(x=c(xdown,xup),
                     y=c(ydown,yup),
                     label = c(labdown,labup))

  # Create a dataframe with links

  plant_positions <- data.frame(plant = labdown,
                                xdown = xdown,
                                ydown = ydown)

  pollinator_positions <- data.frame(pollinator = labup,
                                     xup = xup,
                                     yup = yup)

  link_df <- motif_data %>% left_join(plant_positions, by = "plant") %>% 
    left_join(pollinator_positions, by = "pollinator")

  # Add FG IDs to nodes
  FG_IDs <- tibble(label = c(sort(labdown),sort(labup)),
                   FG = c(plant_nodes,pollinator_nodes))

  d_nodes_FG <- d_nodes %>% left_join(FG_IDs, by = "label")
  levels_FG <- c("Bee","Coleoptera","Lepidoptera",
                 "Non-bee-Hymenoptera", "Non-syrphids-diptera",
                 "Syrphids","1","2","3","4","5")

  d_nodes_FG$FG <- factor(d_nodes_FG$FG, levels = levels_FG)

  levels(d_nodes_FG$FG) <- c("Bee","Coleoptera","Lepidoptera",
                             "Non-bee-Hymenoptera", "Non-syrphids-diptera",
                             "Syrphids","1","2","3","4","5")

  # Before plotting the motif, we remove letters from the node position id

  d_nodes_FG$label <- gsub("[^0-9.-]", "", d_nodes_FG$label)

  # Define the number of colors you want
 mycolors <- c("#A6CEE3", "chocolate2",  "gold2", "#1F78B4","#B2DF8A",
                     "gray58", "lightgoldenrod2","thistle2", "burlywood","lightcoral", "olivedrab3")

  ggplot(data=d_nodes_FG, aes(x, y)) +
    scale_shape_identity() +
    geom_segment(data = link_df, aes(x=xdown, y=ydown, xend = xup, yend = yup), size = 1, color = "grey20")+
    geom_point(aes(color = FG), size=7.5)+
    geom_text(aes(x, y , label = label), size=3)+
    scale_color_manual(drop = FALSE, values = mycolors,
                       breaks=c("Bee","Coleoptera","Lepidoptera",
                                "Non-bee-Hymenoptera", "Non-syrphids-diptera",
                                "Syrphids","1","2","3","4","5"), 
                       labels=c("Bee","Coleoptera","Lepidoptera",
                                "Non-bee-Hymenoptera", "Non-syrphids-diptera",
                                "Syrphids", 
                                "Selfing herbs","Small outcrossing perennials",
                                "Self-incomp. perennials\nwith large flowers",
                                "Tall plants with small\nunisexual flowers",
                                "Short-lived outcrossers with\nlong zygomorphic flowers"))+
    xlim(-1.5,4)+ ylim(-0.5,1.5)+
    theme(panel.background = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())+
    labs(x=NULL,y=NULL)+
    ggtitle(paste0("Motif ",motif_number))+
    theme(plot.title = element_text(hjust = 0.5,size=10))+
    guides(color=guide_legend(title="Functional groups"))

}

# Data with pollinator/plant labels

plant_means_reordered <- read_csv("../Data/Csv/plant_abs_freq_means.csv")
pollinator_means_reordered <- read_csv("../Data/Csv/pollinator_abs_freq_means.csv")
#Filter out in order to have just 6 FG'S
pollinator_means_reordered <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds", "Lizards", "Other_insects"))

####################################################################################################
#Read data to check top 5 motif of each category of observed propabilities

#Under, No signi, Over-rep
####################################################################################################
# Check top Z-scores of each group (Under, No diff, Over)

#Read data
data <- read.csv("../Data/Csv/motifs_observed_probability_SIMUL_CI.csv")

#Find critical value of Z-score
p <- 0.05 #cutoff probability 95% confidence
critical_value <- qnorm(p/2) #double tail probability divide by 2

data_1 <- data %>%
  mutate(infra_over_represented = case_when(
    z_score < -abs(critical_value) ~ "infra",
    between(z_score, -abs(critical_value), abs(critical_value)) ~ "no_diff",
    z_score > abs(critical_value) ~ "over"
  ))

#data_2 <- filter(data_1, abs(z_score)<2000) 
#str(data_2)

#Select under-represented by the ones with highest porbability of being observed (top10)
under_data <- filter(data_1, infra_over_represented=="infra")

under_top_5 <- filter(under_data, row_number(desc(abs(motif_observed_probability)))<= 5) %>%
  select(motif_functional_ID,motif_observed_probability,counts_observed,z_score)

#Select no statistical difference by the ones with highest porbability of being observed (top5)
no_diff_data <- filter(data_1, infra_over_represented=="no_diff")

no_diff_top_5 <- filter(no_diff_data, row_number(desc(abs(motif_observed_probability)))<= 5) %>%
  select(motif_functional_ID,motif_observed_probability,counts_observed,z_score)

#Select over-represenetd by the ones with highest porbability of being observed (top5)
over_data <- filter(data_1, infra_over_represented=="over")

over_top_5 <- filter(over_data, row_number(desc(abs(motif_observed_probability)))<= 5) %>%
  select(motif_functional_ID,motif_observed_probability,counts_observed,z_score)
####################################################################################################
#Now plot it 
####################################################################################################

library(patchwork)

#Prepare list of 5 plots for each group
#Under
labels_under <- under_top_5$motif_functional_ID
p_under <- list()

for (i in 1:5){

  p_under[[i]] <- plot_motif_label(labels_under[[i]])
}

#No signficance
labels_nodiff <- no_diff_top_5$motif_functional_ID
p_nodiff <- list()

for (i in 1:5){

  p_nodiff[[i]] <- plot_motif_label(labels_nodiff[[i]])
}

#Over
labels_over <- over_top_5$motif_functional_ID
p_over <- list()

for (i in 1:5){

  p_over[[i]] <- plot_motif_label(labels_over[[i]])
}

#Run Script with histogram plot
source("../Scripts/Z-score_percentile_histogram_plot.R")

hist <- ggplot(data_2 %>% filter(round_motif_observed_probability>0), aes(x=z_score, color=infra_over_represented, fill=infra_over_represented)) + 
  geom_histogram(bins = 93, alpha = 0.5, position = "identity",lwd = 0.25)+
  geom_vline(xintercept = -abs(critical_value))+
  geom_vline(xintercept = abs(critical_value))+
  xlim(-60,60) + ylab("Frequency")+  theme_bw() +
  scale_fill_manual(name="Motif frequencies" ,values=c("coral2", "palegreen3", "cyan3"), labels=c("Under-represented",
                                                                                                  "No statistical difference", "Over-represented")) +
  scale_color_manual(name="Motif frequencies" ,values=c("coral2", "palegreen3", "cyan3"), labels=c("Under-represented",
                                                                                                   "No statistical difference", "Over-represented")) + xlab("Z-score")

t_col <- function(color, percent = 50, name = NULL) {
  #      color = color name
  #    percent = % transparency
  #       name = an optional name for the color

  ## Get RGB values for named color
  rgb.val <- col2rgb(color)

  ## Make new color using input color as base and alpha set by transparency
  t.col <- rgb(rgb.val[1], rgb.val[2], rgb.val[3],
               max = 255,
               alpha = (100 - percent) * 255 / 100,
               names = name)

  ## Save the color
  invisible(t.col)
}
## END

#Prepare panel of plots
hist / ((p_under[[1]]/p_under[[2]]/p_under[[3]]/p_under[[4]] & theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill=t_col("coral2")))) | 
          (p_nodiff[[1]]/p_nodiff[[2]]/p_nodiff[[3]]/p_nodiff[[4]]& theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill=t_col("palegreen3")))) | 
          (p_over[[1]]/p_over[[2]]/p_over[[3]]/p_over[[4]] & theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill=t_col("cyan3"))))) + plot_layout(guides = 'collect')

```

##DISCUSSION

##REFERENCES

<div id ="refs"></div>

\newpage

##SUPPLEMENTARY MATERIAL

**TABLES**

**Table S1** List of plant-pollinator network studies.

**Table S2** Type of plant traits used in this study.

**FIGURES**

**Figure S1** Over and under-estimated motif frequencies.

**Figure S2** Over and under-estimated functional groups on motif positions.

**Figure S3** Plant trait composition of functional groups.

**Figure S4** Plant dendrogram with functional groups.

\blandscape

```{r, echo=FALSE, message=FALSE, cache=FALSE,fig.width=12,  warning=FALSE}
##############################
#TABLE S1. List of studies
##############################

#Load libraries
library(kableExtra)
library(dplyr)

#Create data.frame with references

#vector with number of netw per study
`Number of networks` <- c("6", #1
                          "2", #2
                          "3",#3
                          "1",#4
                          "1",#5
                          "1",#6
                          "1",#7
                          "2",#8
                          "1",#9
                          "1",#10
                          "8",#11
                          "16",#12
                          "6",#13
                          "2",#14
                          "4",#15
                          "1",#16
                          "3",#17
                          "1"#18
                          )

`First author` <- c("Bartomeus", #1
                    "Dicks",#2
                    "Dupont",#3
                    "Elberling",#4
                    "Fang",#5
                    "Inouye",#6
                    "Lundgren",#7
                    "Olesen",#8
                    "Small",#9
                    "Souza",#10
                    "Kaiser-Bunbury",#11
                    "Bartomeus",#12
                    "Kaiser-Bunbury",#13
                    "Kaiser-Bunbury",#14
                    "Peralta",#15
                    "Burkle",#16
                    "Arroyo-Correa",#17
                    "Bundgaard"#18
                    )

Year <- c(2008,#1
          2002,#2
          2003,#3
          1999,#4
          2008,#5
          1988,#6
          2005,#7
          2002,#8
          1976,#9
          2017,#10
          2017,#11
          2015,#12
          2011,#13
          2010,#14
          2006,#15
          2013,#16
          2019,#17
          2003#18
          )

Country <- c("Spain", #1
               "England", #2
               "Denmark", #3
               "Sweden", #4
               "China", #5
               "United States", #6
               "Denmark (Greenland)", #7
               "Mauritius and Portugal (Azores)", #8
               "Japan", #9
               "Brazil", #10
               "Seychelles", #11
               "Spain", #12
               "Seychelles", #13
               "Mauritius", #14
               "Argentina", # 15
               "United States", #16
               "New Zealand", #17
               "Denmark" #18
               )

DOI <- c("https://doi.org/10.1007/s00442-007-0946-1", #1
         "https://doi.org/10.1046/j.0021-8790.2001.00572.x", #2
         "https://doi.org/10.1111/j.1365-2656.2008.01501.x", #3
         "https://doi.org/10.1111/j.1600-0587.1999.tb00507.x", #4
         "https://doi.org/10.1111/1749-4877.12190", #5
         "https://doi.org/10.1111/j.1442-9993.1988.tb00968.x", #6
         "https://doi.org/10.1657/1523-0430(2005)037[0514:TDAHCW]2.0.CO;2", #7
         "https://doi.org/10.1046/j.1472-4642.2002.00148.x", #8
         "/13960/t4km08d21", #9
         "https://doi.org/10.1111/1365-2745.12978", #10
         "https://doi.org/10.1038/nature21071", #11
         "https://github.com/ibartomeus/BeeFunData", #12
         "https://doi.org/10.1111/j.1365-2745.2010.01732.x", #13
         "https://doi.org/10.1016/j.ppees.2009.04.001", #14
         "https://doi.org/10.1111/ele.13510", #15
         "https://doi.org/10.1126/science.1232728", #16
         "https://doi.org/10.1111/1365-2745.13332", #17
         "Unpublished, Master thesis" #18
         )

references <- data.frame(`First author`, Year, `Number of networks`, Country,   DOI)

colnames(references) <- c("First author", "Year", "Number of networks", "Country",  "DOI")

#Check number of studies
#nrow(references) # 18 studies
#check number of networks
#sum(as.numeric(references$`Number of networks`)) #60

references %>%
  arrange(`First author`)%>%
  kable( longtable = T, booktabs = T,linesep = "\\addlinespace",align = c("cccc"),escape = FALSE, caption= "\\textbf{Table S1.} List of studies ordered by author with the year of publication, number of contributed networks and digital object identifier") %>%
  kable_styling(latex_options = c("repeat_header","striped"),  full_width=F,position = c("center"))

```

\elandscape

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=16, fig.width=12}

library(kableExtra)

library(dplyr)
library(readxl) #read excel file (trait data)
library(dplyr) #data manipulation
library(visdat) #VISUALIZE MISSING DATA
library(naniar)
library(tidyverse)
########################################################################################################################################################
#1) READ TRAIT DATA
########################################################################################################################################################

Trait <- c("Plant height (m)", "Flower width (mm)", "Flower length (mm)", "Inflorescence width (mm)", "Style length (mm)", "Ovules per flower",
           "Flowers per plant",  "Autonomous selfing (fruit set)")

Trait1 <- c("Lifepan", "Life form", "Flower shape", "Flower symmetry", "Autonomous selfing", "Compatibility system", "Breeding system", "")

T.cat <- linebreak(c("Short-lived \n Perennial", "Herb \n Shrub \n Tree", "Brush \n Campanulate \n Capitulum \n Open \n Papilionaceous \n Tube",  "Actinomorphic \n Zygomorphic", "None \n Low \n Medium \n High", "Self-incomp. \n Partially self-comp. \n Self-comp.", "Hermaphrodite \n Monoecious \n Dioecious",""),align = c("l"))

Type <- c("Vegetative", "Floral", "Floral", "Floral", "Floral", "Floral", "Floral","Reproductive")

Type_1 <- c("Vegetative", "Vegetative", "Floral", "Floral", "Reproductive", "Reproductive", "Reproductive", "")

dat <- data.frame(Type, Trait,Type_1, Trait1)

colnames(dat) <- c("Type","Traits","Type", "Traits")

kbl(dat, booktabs = T, linesep = "\\addlinespace", escape = FALSE, align = c("c","l","c","c","l","l","c"),caption = "\\textbf{Table S2.} Traits used to delimit the different plant functional groups divided in quantitative and categorical traits.") %>%
kable_styling(full_width = T,latex_options = c("striped")) %>%
add_header_above(c("Quantitative traits" = 2,"Categorical traits" = 2),bold=T)  %>%
    column_spec(c(1,4), bold = T) %>% row_spec(0, bold=T)

```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.width=4.5,fig.height=6.5, fig.cap="**Figure S1.** Comparison of motif frequencies between empirical and simulated networks grouped by average path length (plots a, b, c and d) as determined in Simmons et al. (2020) without considering singletones. This is shown with the mean percentage of motif frequencies in empirical networks that were over the motif frequencies of the simulated ones (percentiles). This was done by network (light blue dots) and then averaged for all networks (black dots with error bars that correspond to the standard deviation)."}
library(lme4)
library(nlme)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models_NOSING/Motifs_frequency_percentile_NOSING.csv")
d <- add_study_id(d)
#str(d)
#head(d)

#ggplot(d %>% filter(motif != 1), aes(percentil_sizeclass))+
#  geom_histogram(color="black", fill="white")+
#  facet_wrap(~motif)+
#  theme_bw()+
#  labs(x="Sizeclass percentile")
#
# Motifs tend to be over- and under-represensented in real networks

# Mean percentile and SE taking into account the study system
motif_codes <- d %>% filter(motif != 1) %>% # We remove links (motif code: 1)
  select(motif) %>% unique() %>% pull()
motif_means <- tibble(motif = motif_codes)
motif_means$mean <- NA
motif_means$SE <- NA

for(i.motif in 1:length(motif_codes)){

  m <- lmer(percentil_sizeclass ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(d, motif == motif_codes[i.motif]))

  motif_means$mean[i.motif] <- fixed.effects(m) %>% unname()
  motif_means$SE[i.motif] <- sqrt(diag(vcov(m)))

}

#Add categories based on path length set by Simmons 2021 et al., Functional Ecology
d$Broad_categories[d$motif==2 | d$motif==7 | d$motif==17 | d$motif==3|d$motif==4|d$motif==8] <- "Fan"

d$Broad_categories[d$motif==15 | d$motif==11] <- "Medium-weak"

d$Broad_categories[d$motif==6 | d$motif==16 |  d$motif==12] <- "Strong"

d$Broad_categories[d$motif==5 | d$motif==9 |  d$motif==14|  d$motif==13|  d$motif==10] <- "Weak"

#order the categories from short to long path length
d$Broad_categories <- factor(d$Broad_categories, levels = c("Strong", "Fan", "Medium-weak", "Weak"))

#Ffilter out two node motifs
d_1 <- d %>% filter(motif != 1)
#Merge dataframes to bl able to create facets
d_2 <- merge(d_1, motif_means, by="motif")

arr <- list('Strong' = expression(bold(paste("Complete (strong, ", bar(x), " = 1.38)"))),
             'Fan' = expression(bold(paste("Fan (medium-strong, ", bar(x), " = 1.48)"))), 
             'Medium-weak' = expression(bold(paste("Asymmetric complete (medium-weak, ", bar(x), " = 1.60)"))), 
             "Weak" =expression(bold(paste("Core-peripheral (weak, ", bar(x), " = 1.85)"))))
mylabel <- function(val) { return(lapply(val, function(x) arr[x])) }

library(ggforce) #workaround with ggforce to have facets with same sizes

p <- ggplot()+geom_point(data = d_2, aes(y=factor(motif), 100*percentil_sizeclass), shape=21,color="black",fill="azure2",size=2, alpha=1)+
  geom_errorbar(data = d_2,aes(y = factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=1.0,size=1,color="grey20") +theme_bw()+
  geom_point(data = d_2,aes(y = as.factor(motif), x=100*mean), 
             size=2,color="grey20")+labs(y="Two to five node motifs", x = "Percentile")+
  theme( strip.background = element_blank())+
  theme(strip.text = element_text(face="bold", size=10, hjust=0),
        panel.border=element_rect(color="black",size=1.2))

p + ggforce::facet_col(vars(Broad_categories), scales = 'free', space = 'free',labeller=mylabel)

#Trying to maintain a gradient from low to high in colors

#Fig 1 Should be a dot plot (or forest plot) of the mean +- SE of the 15 motifs.
#NICE! Can we put motifs in Y axes, and % in X?
```

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.height=8,fig.width=8,fig.align = 'center', fig.cap="\\textbf{Figure S2.} Heatmap indicating under- and over- representation of pollinator and plant functional groups in the different motif positions excluding singletones, after removing non-robust links, that is, interactions whose frequency was equal to one. The different motif positions are dividied by the average path length clasification by Simmons et al. (2020)"}

library(lme4)
library(nlme)
library(tidyverse)
library(stringi)
source("../Scripts/Scripts_Alfonso/add_study_id.R")

#############################################
# EXTRACT MEAN VALUES
#############################################

all_position_percentiles <- read_csv("../Data/Csv/Motifs_positions_and_null_models_NOSING/GF_positions_frequency_percentile_NOSING.csv")

all_position_percentiles$position <- stri_extract_first_regex(all_position_percentiles$position,
                                                              "[0-9]+") %>% as.numeric()

all_position_percentiles <- add_study_id(all_position_percentiles)
#str(all_position_percentiles)
#head(all_position_percentiles)

list_plants_FG <- as.character(1:10)

plant_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF)) #ME DA ERROR
pollinator_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(!Node_FG %in% list_plants_FG, !position %in% c(1,2),
         !is.na(expected_freq_its_GF))

# Plants-------------

# Mean percentile and SE taking into account the study system
plant_codes <- plant_position_percentiles_filtered$position %>% unique()
plant_means <- plant_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
plant_means$mean <- NA
plant_means$SE <- NA

for(i.pos in 1:length(plant_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id),
            data = plant_position_percentiles_filtered %>%
              filter(position == plant_codes[i.pos]))

  temp <- fixed.effects(m) %>% unname()
  plant_means$mean[plant_means$position == plant_codes[i.pos]] <- c(temp[1],
                                                                    temp[1]+temp[2],
                                                                    temp[1]+temp[3],
                                                                    temp[1]+temp[4],
                                                                    temp[1]+temp[5]) ## to get real means per FG
  temp <- sqrt(diag(vcov(m)))
  plant_means$SE[plant_means$position == plant_codes[i.pos]] <- c(temp[1], 
                                                                  temp[1]+temp[2],
                                                                  temp[1]+temp[3],
                                                                  temp[1]+temp[4],
                                                                  temp[1]+temp[5])

}

# Pollinators--------

# Mean percentile and SE taking into account the study system
pollinator_codes <- pollinator_position_percentiles_filtered$position %>% unique()
pollinator_means <- pollinator_position_percentiles_filtered %>% select(position,Node_FG) %>% unique()
pollinator_means$mean <- NA
pollinator_means$SE <- NA
pollinator_means_reordered <- NULL

for(i.pos in 1:length(pollinator_codes)){

  m <- lmer(percentil_its_GF ~ Node_FG + (1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = pollinator_position_percentiles_filtered %>%
              filter(position == pollinator_codes[i.pos]),
            control = lmerControl(optimizer ="Nelder_Mead")) # I changed the optimizer because motif 8 (i.pos = 3) caused convergence problems, max gradient = 0.0022 > 0.002.

  temp <- fixed.effects(m) %>% unname()
  pollinator_means_aux <- pollinator_means[pollinator_means$position == pollinator_codes[i.pos],] %>% arrange(Node_FG)

  pollinator_means_aux$mean <- c(temp[1],
                                 temp[1]+temp[2],
                                 temp[1]+temp[3],
                                 temp[1]+temp[4],
                                 temp[1]+temp[5],
                                 temp[1]+temp[6],
                                 temp[1]+temp[7],
                                 temp[1]+temp[8],
                                 temp[1]+temp[9]) ## to get real means per FG

  temp <- sqrt(diag(vcov(m)))
  pollinator_means_aux$SE <- c(temp[1],
                               temp[1]+temp[2],
                               temp[1]+temp[3],
                               temp[1]+temp[4],
                               temp[1]+temp[5],
                               temp[1]+temp[6],
                               temp[1]+temp[7],
                               temp[1]+temp[8],
                               temp[1]+temp[9])

  pollinator_means_reordered <- bind_rows(pollinator_means_reordered,pollinator_means_aux)

}

# There are lower limits (mean - SE) smaller than 0 (see Other insects)
# We set the lower limit of those error bars to zero

pollinator_means_reordered <- pollinator_means_reordered %>% mutate(lower = mean-SE)
pollinator_means_reordered$lower[pollinator_means_reordered$lower < 0] <- 0

#############################################
# ADDING BROAD MOTIF CATEGORIES TO POSITIONS
#############################################

motifs_raw_data <- read_csv("../Data/Data_processing/Motifs_connections/motif_pattern_connections.csv")

motifs_raw_data$Broad_categories <- NA

motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(5,9,14,13,10)] <- "Core-peripherical"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(6,16,12)] <- "Complete"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(2,3,7,4,17,8)] <- "Fan"
motifs_raw_data$Broad_categories[motifs_raw_data$motif_id %in% c(15,11)] <- "Asymmetric complete"

motifs_raw_data$plant <- gsub("[^0-9.-]", "", motifs_raw_data$plant)
motifs_raw_data$pollinator <- gsub("[^0-9.-]", "", motifs_raw_data$pollinator)
motifs_raw_data$plant <- as.numeric(motifs_raw_data$plant)
motifs_raw_data$pollinator <- as.numeric(motifs_raw_data$pollinator)
motifs_raw_data <- motifs_raw_data %>% filter(!is.na(Broad_categories))

fan_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_pollinator <- motifs_raw_data$pollinator[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

pollinator_means_reordered$Broad_categories <- NA

pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% fan_pollinator] <- "Fan" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% asymm_pollinator] <- "Medium-weak" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% complete_pollinator] <- "Strong" 
pollinator_means_reordered$Broad_categories[pollinator_means_reordered$position %in% core_per_pollinator] <- "Weak"

fan_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Fan"] %>% unique()
core_per_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Core-peripherical"] %>% unique()
complete_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Complete"] %>% unique()
asymm_plant <- motifs_raw_data$plant[motifs_raw_data$Broad_categories == "Asymmetric complete"] %>% unique()

plant_means$Broad_categories <- NA

plant_means$Broad_categories[plant_means$position %in% fan_plant] <- "Fan" 
plant_means$Broad_categories[plant_means$position %in% asymm_plant] <- "Medium-weak" 
plant_means$Broad_categories[plant_means$position %in% complete_plant] <- "Strong" 
plant_means$Broad_categories[plant_means$position %in% core_per_plant] <- "Weak"

########################################
# CREATE HEATMAPS
########################################

library(ComplexHeatmap)

# PLANTS-----

# Prepare data for heatmap
heat_plants_aux <- plant_means %>% spread(Node_FG,mean) %>% select(-SE)

heat_plants <- heat_plants_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_plants_fan <- heat_plants %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_fan) <- heat_plants_fan[,1]

heat_plants_Weak <- heat_plants %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Weak) <- heat_plants_Weak[,1]

heat_plants_Strong <- heat_plants %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Strong) <- heat_plants_Strong[,1]

heat_plants_Medium_weak <- heat_plants %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_plants_Medium_weak) <- heat_plants_Medium_weak[,1]

# POLLINATORS-----------

filter_rare_groups <- T

if(filter_rare_groups == T){
  heat_pollinators_aux <- pollinator_means_reordered %>% filter(!Node_FG %in% c("Birds",
                                                                                "Lizards",
                                                                                "Other_insects")) %>% 
    spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee",            
                                      "Coleoptera","Lepidoptera","Non-bee\nHymenoptera", 
                                      "Non-syrphids\ndiptera","Syrphids")
}else{
  heat_pollinators_aux <- pollinator_means_reordered %>% spread(Node_FG,mean) %>% select(-SE,-lower)

  colnames(heat_pollinators_aux) <- c("position","Broad_categories","Bee","Birds",            
                                      "Coleoptera","Lepidoptera","Lizards","Non-bee\nHymenoptera", 
                                      "Non-syrphids\ndiptera","Other insects","Syrphids")
}

heat_pollinators <- heat_pollinators_aux %>%
  group_by(position,Broad_categories) %>% 
  summarise_all(sum,na.rm = T)

heat_pollinators_fan <- heat_pollinators %>% filter(Broad_categories == "Fan") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_fan) <- heat_pollinators_fan[,1]

heat_pollinators_Weak <- heat_pollinators %>% filter(Broad_categories == "Weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Weak) <- heat_pollinators_Weak[,1]

heat_pollinators_Strong <- heat_pollinators %>% filter(Broad_categories == "Strong") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Strong) <- heat_pollinators_Strong[,1]

heat_pollinators_Medium_weak <- heat_pollinators %>% filter(Broad_categories == "Medium-weak") %>% select(-Broad_categories) %>% 
  as.matrix()

rownames(heat_pollinators_Medium_weak) <- heat_pollinators_Medium_weak[,1]

##################

#Do not know how to handle this multiplot for creating a panel

# I'm going to create the 4 plots in one, 
# I'm not sure that the grouping of the functional group is correct
# when is done separately

#sorry for the mess of libraries that I do below

 #Prepare first pollinators

library(tidyr)
library(data.table)
long <- melt(setDT(heat_pollinators), id.vars = c("position","Broad_categories"), variable.name = "guilds")

long_1 <- long %>% select(-Broad_categories)

library(reshape2)
long_m <- acast(long_1, position~guilds, value.var="value")

#set manually the divisions, so far this is the best way I have seen to split the rows by categories

r_split <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 4),
             rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 3),rep(c("Asymmetric complete"), 2),
             "Complete", "Fan")

#set order of labels that avoids overlapping
r_split <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

#plat a bit with colors, a bit of a more complex set but looks nicer
library(circlize)
col_fun = colorRamp2(c(0.05644663,0.25, 0.4673256,0.75, 0.8782047), c("blue","cadetblue3", "azure2", "gold","red"))

h_pol <- Heatmap(long_m, name = "Mean\npercentile", row_split=r_split,cluster_row_slices = FALSE, 
                 column_title = "Pollinator functional groups",row_names_gp = gpar(fontsize = 10),
                 col = col_fun,column_names_rot = 45,  column_names_gp = grid::gpar(fontsize = 10),
                 column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                 row_title_gp = gpar(fontsize = 10, fontface = "bold"))

 #Prepare plants

#long format
long_plants <- melt(setDT(heat_plants), id.vars = c("position","Broad_categories"), variable.name = "Functional_groups")

#filter for creating matrix
long_plants_1 <- long_plants %>% select(-Broad_categories)

#set better colnames
long_plants_1$Functional_groups <- as.character(long_plants_1$Functional_groups)

long_plants_1$Functional_groups[long_plants_1$Functional_groups==1] <- "Selfing herbs" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==2] <- "Small outcrossing\nperennials" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==3] <- "Self-incomp. perennials\nwith large flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==4] <- "Tall plants with small\nunisexual flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==5] <- "Short-lived outcrossers with\nlong zygomorphic flowers"

long_plants_1$Functional_groups <- factor(long_plants_1$Functional_groups, levels=c("Selfing herbs",
"Small outcrossing\nperennials" ,"Self-incomp. perennials\nwith large flowers",
"Tall plants with small\nunisexual flowers" ,"Short-lived outcrossers with\nlong zygomorphic flowers" ))

#create matrix with reshape
long_plants_m <- acast(long_plants_1, position~Functional_groups, value.var="value")

r_split_plants <- c(rep(c("Fan"), 3),rep(c("Core-peripheral"), 2),"Complete",rep(c("Fan"),2),rep(c("Core-peripheral"), 3),
                    rep(c("Asymmetric complete"), 2), "Complete", rep(c("Core-peripheral"), 4),rep(c("Asymmetric complete"), 2),
                    "Complete", "Fan")

r_split_plants <- factor(r_split, levels=c("Core-peripheral","Complete","Fan","Asymmetric complete"))

h_plan <- Heatmap(long_plants_m, name = "Mean\npercentile", row_split=r_split_plants,cluster_row_slices = F, 
                  column_title = "Plant functional groups",row_names_gp = gpar(fontsize = 10),col = col_fun,column_names_rot = 45,  
                  column_names_gp = grid::gpar(fontsize = 10),column_title_gp = gpar(fontsize = 15, fontface = "bold"),
                  row_title_gp = gpar(fontsize = 10, fontface = "bold"))

library(patchwork)
ht_list <- h_pol+h_plan

draw(ht_list, auto_adjust = FALSE, cluster_rows = F)

```

\newpage

\blandscape

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=13, fig.cap="**Figure S3.** Plant functional group composition separated in qualitative and quantitative variables. Panel A) shows the percentage of the different categories within trait represented with different colours for each functional group. Plot B) shows the radar plot of the different quantitative variables standardize on the same scale also coloured with the same patterns of colours as qualitative variables per cluster."}

#CODE TO PLOT FUNCTIONAL GROUPS
#IT CAN BE PLOT WITH MUCH LESS CODE BUT NO TIME FOR CLEANING IT

##########################################
#Visualization of plant functional groups#
##########################################

library(cowplot)
library(ggplot2)
library(tidyr)
library(dplyr)
library(reshape2)
library(data.table)
library(patchwork)

#LOAD DATA
#read unscaled trait data in order to visualize better the clusters
d <- read.csv("../Data/Csv/all_species_imputed_trait_data_forest_data.csv")

dat_1 <- read.csv("../Data/Csv/imputed_trait_data_hclust_5_clusters_forest_data.csv", row.names = "X")

d$Clusters <- dat_1$Clusters

#select columns of interest
t <- d[c("Breeding_system","IMPUTED_Compatibility","Autonomous_selfing_level",
         "Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", "Flowers_per_plant", "Flowers_per_inflorescence",
         "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean", "STYLE_IMPUTED", "OVULES_IMPUTED", "life_form", "lifespan",
         "IMPUTED_plant_height_mean_m", "Nectar_presence_absence", "Clusters")]

##############################################################################################################################
#Prepare table qualitative variables
##############################################################################################################################

##############################################################################################################################
#Breeding system
t$Breeding_system <- factor(t$Breeding_system, levels=c("Dioecious", "Monoecious", "Hermaphrodite"))

#Summary to create table
breeding <- t %>%
  group_by(Clusters,Breeding_system) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

breeding <- breeding[,-3]

breeding <- spread(breeding, Clusters, Percentage)

colnames(breeding)[1] <- "Category"

Trait <- data.frame(Trait= c(rep("Breeding system", 3)))

Breeding_system <- cbind(Trait, breeding)

##############################################################################################################################
#Compatibility
str(t)
t$IMPUTED_Compatibility <- as.character(t$IMPUTED_Compatibility)
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="dioecious"] <- "Unisexual flower"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="monoecious"] <- "Unisexual flower"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_incompatible"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="partially_self_compatible"] <- "Partially self compatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_compatible"] <- "Self compatible"

t$IMPUTED_Compatibility <- factor(t$IMPUTED_Compatibility, levels=c("Unisexual flower", "Self incompatible", "Partially self compatible", "Self compatible"))
#Summary to create table
comp <- t %>%
  group_by(Clusters,IMPUTED_Compatibility) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

comp <- comp[,-3]

comp <- spread(comp, Clusters, Percentage)
#Replace Na's with zeros
comp[is.na(comp)] <- 0
#Set call name os categories per trait
colnames(comp)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Compatibility system", 4)))
#Add column
Compatibility_system <- cbind(Trait, comp)
#Bind rows
bind_rows(Breeding_system, Compatibility_system)

############################################################################################################################################
#Life form
t$life_form <- as.character(t$life_form)
t$life_form[t$life_form =="vine"] <- "Shrub"
t$life_form[t$life_form =="tree"] <- "Tree"
t$life_form[t$life_form =="herb"] <- "Herb"
t$life_form[t$life_form =="shrub"] <- "Shrub"
#Order levels
t$life_form <- factor(t$life_form, levels=c("Tree", "Shrub", "Herb"))
#Summary to create table
life_form <- t %>%
  group_by(Clusters,life_form) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_form <- life_form[,-3]
#convert to wide
life_form <- spread(life_form, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_form)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life form", 3)))
#Add column
Life_form<- cbind(Trait, life_form)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form)

############################################################################################################################################
t$lifespan <- factor(t$lifespan, levels=c("Perennial","Short lived"))
#Summary to create table
life_span <- t %>%
  group_by(Clusters,lifespan) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_span <- life_span[,-3]
#convert to wide
life_span <- spread(life_span, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_span)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life span", 2)))
#Add column
Life_span<- cbind(Trait, life_span)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span)

############################################################################################################################################
t$Flower_morphology <- as.character(t$Flower_morphology)
t$Flower_morphology[t$Flower_morphology=="Funnelform"] <- "Campanulate"
t$Flower_morphology[t$Flower_morphology=="Spike"] <- "Brush"

Flower_morphology <- t %>%
  group_by(Clusters,Flower_morphology) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
Flower_morphology <- Flower_morphology[,-3]
#convert to wide
Flower_morphology <- spread(Flower_morphology, Clusters, Percentage)
#Set call name os categories per trait
colnames(Flower_morphology)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Flower shape", 6)))
#Add column
Flower_morphology<- cbind(Trait, Flower_morphology)
#Replace Na's with zeros
Flower_morphology[is.na(Flower_morphology)] <- 0
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology)

############################################################################################################################################
t$Flower_symmetry <- as.character(t$Flower_symmetry)
t$Flower_symmetry[t$Flower_symmetry =="actinomorphic"] <- "Actinomorphic"
t$Flower_symmetry[t$Flower_symmetry =="zygomorphic"] <- "Zygomorphic"

Flower_symmetry <- t %>%
  group_by(Clusters,Flower_symmetry) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
Flower_symmetry <- Flower_symmetry[,-3]
#convert to wide
Flower_symmetry <- spread(Flower_symmetry, Clusters, Percentage)
#Set call name os categories per trait
colnames(Flower_symmetry)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Flower symmetry", 2)))
#Add column
Flower_symmetry<- cbind(Trait, Flower_symmetry)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology,Flower_symmetry)
############################################################################################################################################
t$Autonomous_selfing_level <- as.character(t$Autonomous_selfing_level)
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="none"] <- "None"
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="low"] <- "Low"
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="medium"] <- "Medium"
t$Autonomous_selfing_level[t$Autonomous_selfing_level =="high"] <- "High"

#Order levels
t$Autonomous_selfing_level <- factor(t$Autonomous_selfing_level, levels=c("High", "Medium", "Low", "None"))

Selfing <- t %>%
  group_by(Clusters,Autonomous_selfing_level) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

#delete column to convert to wide
Selfing <- Selfing[,-3]
#convert to wide
Selfing <- spread(Selfing, Clusters, Percentage)
#Set call name os categories per trait
colnames(Selfing)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Selfing level", 2)))
#Replace Na's with zeros
Selfing[is.na(Selfing)] <- 0
#Add column
Selfing_1 <- cbind(Trait, Selfing)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Selfing_1)

############################################################################################################################################
t$Nectar_presence_absence <- as.character(t$Nectar_presence_absence)
t$Nectar_presence_absence[t$Nectar_presence_absence =="yes"] <- "Yes"
t$Nectar_presence_absence[t$Nectar_presence_absence =="no"] <- "No"

Nectar <- t %>%
  group_by(Clusters,Nectar_presence_absence) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

#delete column to convert to wide
Nectar <- Nectar[,-3]
#convert to wide
Nectar <- spread(Nectar, Clusters, Percentage)
#Set call name os categories per trait
colnames(Nectar)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Nectar", 2)))
#Replace Na's with zeros
Nectar[is.na(Nectar)] <- 0
#Add column
Nectar <- cbind(Trait, Nectar)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Flower_symmetry,Selfing_1, Nectar)

##############################################################################################################################
#Prepare table quantitative variables
##############################################################################################################################

Selfing <-  t %>% group_by(Clusters)%>% 
  select(Autonomous_selfing_level_fruit_set) %>% # select variables to summarise
  summarise_each(funs(min = min,
                     # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                    sd = sd))

s1 <- as.data.frame(t(Selfing))
#set colnames
colnames(s1) <- c("1","2","3","4","5")
#remove the cluster row
s1 <- s1[-1,]
#convert rownames to 1st columns
s1 <- setDT(s1, keep.rownames = "TRUE")[]
colnames(s1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Autonomous selfing", 5)))
#Add column
Selfing <- cbind(Trait, s1)

##############################################################################################################################
flower_number <-  t %>% group_by(Clusters)%>% 
  select(Flowers_per_plant) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

f1 <- as.data.frame(t(flower_number))
#set colnames
colnames(f1) <- c("1","2","3","4","5")
#remove the cluster row
f1 <- f1[-1,]
#convert rownames to 1st columns
f1 <- setDT(f1, keep.rownames = "TRUE")[]
colnames(f1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flower number", 5)))
#Add column
Flower_number <- cbind(Trait, f1)

##############################################################################################################################
Flowers_per_inflorescence <-  t %>% group_by(Clusters)%>% 
  select(Flowers_per_inflorescence) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

f_i_1 <- as.data.frame(t(Flowers_per_inflorescence))
#set colnames
colnames(f_i_1) <- c("1","2","3","4","5")
#remove the cluster row
f_i_1 <- f_i_1[-1,]
#convert rownames to 1st columns
f_i_1 <- setDT(f_i_1, keep.rownames = "TRUE")[]
colnames(f_i_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flowers per inflorescence", 5)))
#Add column
Flower_per_inflo <- cbind(Trait, f_i_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo)

##############################################################################################################################
Floral_unit_width <-  t %>% group_by(Clusters)%>% 
  select(Floral_unit_width) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

f_u_w_1 <- as.data.frame(t(Floral_unit_width))
#set colnames
colnames(f_u_w_1) <- c("1","2","3","4","5")
#remove the cluster row
f_u_w_1 <- f_u_w_1[-1,]
#convert rownames to 1st columns
f_u_w_1 <- setDT(f_u_w_1, keep.rownames = "TRUE")[]
colnames(f_u_w_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Inflorescence width", 5)))
#Add column
Inflo_width <- cbind(Trait, f_u_w_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width)

##############################################################################################################################
Corolla_diameter_mean <-  t %>% group_by(Clusters)%>% 
  select(Corolla_diameter_mean) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

c_w_1 <- as.data.frame(t(Corolla_diameter_mean))
#set colnames
colnames(c_w_1) <- c("1","2","3","4","5")
#remove the cluster row
c_w_1 <- c_w_1[-1,]
#convert rownames to 1st columns
c_w_1 <- setDT(c_w_1, keep.rownames = "TRUE")[]
colnames(c_w_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flower width", 5)))
#Add column
Flower_width <- cbind(Trait, c_w_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width)

##############################################################################################################################
Corolla_length_mean <-  t %>% group_by(Clusters)%>% 
  select(Corolla_length_mean) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

c_l_1 <- as.data.frame(t(Corolla_length_mean))
#set colnames
colnames(c_l_1) <- c("1","2","3","4","5")
#remove the cluster row
c_l_1 <- c_l_1[-1,]
#convert rownames to 1st columns
c_l_1 <- setDT(c_l_1, keep.rownames = "TRUE")[]
colnames(c_l_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Flower length", 5)))
#Add column
Flower_length<- cbind(Trait, c_l_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length)

##############################################################################################################################
STYLE_IMPUTED <-  t %>% group_by(Clusters)%>% 
  select(STYLE_IMPUTED) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

s_l_1 <- as.data.frame(t(STYLE_IMPUTED))
#set colnames
colnames(s_l_1) <- c("1","2","3","4","5")
#remove the cluster row
s_l_1 <- s_l_1[-1,]
#convert rownames to 1st columns
s_l_1 <- setDT(s_l_1, keep.rownames = "TRUE")[]
colnames(s_l_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Style length", 5)))
#Add column
Style_length<- cbind(Trait, s_l_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length, Style_length)

##############################################################################################################################

OVULES_IMPUTED <-  t %>% group_by(Clusters)%>% 
  select(OVULES_IMPUTED) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

o_n_1 <- as.data.frame(t(OVULES_IMPUTED))
#set colnames
colnames(o_n_1) <- c("1","2","3","4","5")
#remove the cluster row
o_n_1 <- o_n_1[-1,]
#convert rownames to 1st columns
o_n_1 <- setDT(o_n_1, keep.rownames = "TRUE")[]
colnames(o_n_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Ovule number", 5)))
#Add column
Ovule_number <- cbind(Trait, o_n_1)

bind_rows(Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length, Ovule_number)

##############################################################################################################################

IMPUTED_plant_height_mean_m <-  t %>% group_by(Clusters)%>% 
  select(IMPUTED_plant_height_mean_m) %>% # select variables to summarise
  summarise_each(funs(min = min,
                      # q25 = quantile(., 0.25), 
                      median = median, 
                      #q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd))

p_h_1 <- as.data.frame(t(IMPUTED_plant_height_mean_m))
#set colnames
colnames(p_h_1) <- c("1","2","3","4","5")
#remove the cluster row
p_h_1 <- p_h_1[-1,]
#convert rownames to 1st columns
p_h_1 <- setDT(p_h_1, keep.rownames = "TRUE")[]
colnames(p_h_1)[1] <- "Category"
#Create data frame to add grouping variable
Trait <- data.frame(Trait= c(rep("Plant height", 5)))
#Add column
Plant_height <- cbind(Trait, p_h_1)

#Bind quantitative and qualitative variables
Table_all_traits <- bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Flower_symmetry,Selfing_1, Nectar,
          Selfing, Flower_number, Flower_per_inflo, Inflo_width, Flower_width, Flower_length, Style_length, Ovule_number, Plant_height)

#Select just two decimals
is.num <- sapply(Table_all_traits, is.numeric)
Table_all_traits[is.num] <- lapply(Table_all_traits[is.num], round, 2)

colnames(Table_all_traits) <- c("Trait", "Category", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")

#LOAD LIBRARIES
#devtools::install_github("ricardo-bion/ggradar", dependencies=TRUE)
suppressPackageStartupMessages(library(dplyr))
library(scales)
library(tibble)
library(ggradar)
library(janitor)
library(grid)
library(dplyr)
library(lubridate)
#QUANTITATIVVE VARIABLES

#generate subset of quantitative variables for plotting
quant_median <- subset(Table_all_traits, Category=="median")

#delete col number 2
quant_median <- quant_median[,-2]

#transpose dataframe
final_df <- as.data.frame(t(quant_median))

#rownames to colnames (1st one)
final_df_1 <- final_df %>% row_to_names(row_number = 1)

#Convert all variables to numeric
final_df_2 <- sapply(final_df_1, function(x) as.numeric(as.character(x)))

#set back the missed rownames in the processing
rownames(final_df_2) <- rownames(final_df_1)

# Convert to dataframe
final_df_2 <- as.data.frame(final_df_2)

# Scale values and set rowname as column
final_df_3 <-final_df_2 %>% rownames_to_column( var = "group" ) %>%   mutate_at(vars(-group),funs(rescale))

#Plot Radar
#ggradar(final_df_3)

#QUALITATIVE VARIABLES

#LOAD DATA
#read unscaled trait data in order to visualize better the clusters
d <- read.csv("../Data/Csv/all_species_imputed_trait_data_forest_data.csv")

dat_1 <- read.csv("../Data/Csv/imputed_trait_data_hclust_5_clusters_forest_data.csv", row.names = "X")

d$Clusters <- dat_1$Clusters

#select columns of interest
t <- d[c("Breeding_system","IMPUTED_Compatibility","Autonomous_selfing_level",
         "Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", "Flowers_per_plant", "Flowers_per_inflorescence",
         "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean", "STYLE_IMPUTED", "OVULES_IMPUTED", "life_form", "lifespan",
         "IMPUTED_plant_height_mean_m", "Nectar_presence_absence", "Clusters")]

##############################################################################################################################
#Prepare table qualitative variables
##############################################################################################################################

##############################################################################################################################
#Breeding system
t$Breeding_system <- factor(t$Breeding_system, levels=c("Dioecious", "Monoecious", "Hermaphrodite"))

#Summary to create table
breeding <- t %>%
  group_by(Clusters,Breeding_system) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

breeding <- breeding[,-3]

breeding <- spread(breeding, Clusters, Percentage)

colnames(breeding)[1] <- "Category"

Trait <- data.frame(Trait= c(rep("Breeding system", 3)))

Breeding_system <- cbind(Trait, breeding)

##############################################################################################################################
#Compatibility
str(t)
t$IMPUTED_Compatibility <- as.character(t$IMPUTED_Compatibility)
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="dioecious"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="monoecious"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_incompatible"] <- "Self incompatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="partially_self_compatible"] <- "Partially self compatible"
t$IMPUTED_Compatibility[t$IMPUTED_Compatibility=="self_compatible"] <- "Self compatible"

t$IMPUTED_Compatibility <- factor(t$IMPUTED_Compatibility, levels=c("Self incompatible", "Partially self compatible", "Self compatible"))
#Summary to create table
comp <- t %>%
  group_by(Clusters,IMPUTED_Compatibility) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

comp <- comp[,-3]

comp <- spread(comp, Clusters, Percentage)
#Replace Na's with zeros
comp[is.na(comp)] <- 0
#Set call name os categories per trait
colnames(comp)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Compatibility system", 3)))
#Add column
Compatibility_system <- cbind(Trait, comp)
#Bind rows
bind_rows(Breeding_system, Compatibility_system)

############################################################################################################################################
#Life form
t$life_form <- as.character(t$life_form)
t$life_form[t$life_form =="vine"] <- "Shrub"
t$life_form[t$life_form =="tree"] <- "Tree"
t$life_form[t$life_form =="herb"] <- "Herb"
t$life_form[t$life_form =="shrub"] <- "Shrub"
#Order levels
t$life_form <- factor(t$life_form, levels=c("Tree", "Shrub", "Herb"))
#Summary to create table
life_form <- t %>%
  group_by(Clusters,life_form) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_form <- life_form[,-3]
#convert to wide
life_form <- spread(life_form, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_form)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life form", 3)))
#Add column
Life_form<- cbind(Trait, life_form)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form)

############################################################################################################################################
t$lifespan <- factor(t$lifespan, levels=c("Perennial","Short lived"))
#Summary to create table
life_span <- t %>%
  group_by(Clusters,lifespan) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
life_span <- life_span[,-3]
#convert to wide
life_span <- spread(life_span, Clusters, Percentage)
#Set call name os categories per trait
colnames(life_span)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Life span", 2)))
#Add column
Life_span<- cbind(Trait, life_span)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span)

############################################################################################################################################
#t$Flower_morphology <- as.character(t$Flower_morphology)
#t$Flower_morphology[t$Flower_morphology=="Funnelform"] <- "Campanulate"
#t$Flower_morphology[t$Flower_morphology=="Spike"] <- "Brush"
#
#Flower_morphology <- t %>%
#  group_by(Clusters,Flower_morphology) %>%
#  summarise (n = n()) %>%
#  mutate(Percentage = n / sum(n)*100)
##delete column to convert to wide
#Flower_morphology <- Flower_morphology[,-3]
##convert to wide
#Flower_morphology <- spread(Flower_morphology, Clusters, Percentage)
##Set call name os categories per trait
#colnames(Flower_morphology)[1] <- "Category"
##Create new column
#Trait <- data.frame(Trait= c(rep("Flower shape", 6)))
##Add column
#Flower_morphology<- cbind(Trait, Flower_morphology)
##Replace Na's with zeros
#Flower_morphology[is.na(Flower_morphology)] <- 0
##Bind rows
#bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology)
#
############################################################################################################################################
t$Flower_symmetry <- as.character(t$Flower_symmetry)
t$Flower_symmetry[t$Flower_symmetry =="actinomorphic"] <- "Actinomorphic"
t$Flower_symmetry[t$Flower_symmetry =="zygomorphic"] <- "Zygomorphic"

Flower_symmetry <- t %>%
  group_by(Clusters,Flower_symmetry) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)
#delete column to convert to wide
Flower_symmetry <- Flower_symmetry[,-3]
#convert to wide
Flower_symmetry <- spread(Flower_symmetry, Clusters, Percentage)
#Set call name os categories per trait
colnames(Flower_symmetry)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Flower symmetry", 2)))
#Add column
Flower_symmetry<- cbind(Trait, Flower_symmetry)
#Bind rows
bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology,Flower_symmetry)
############################################################################################################################################
#t$Autonomous_selfing_level <- as.character(t$Autonomous_selfing_level)
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="none"] <- "None"
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="low"] <- "Low"
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="medium"] <- "Medium"
#t$Autonomous_selfing_level[t$Autonomous_selfing_level =="high"] <- "High"
#
##Order levels
#t$Autonomous_selfing_level <- factor(t$Autonomous_selfing_level, levels=c("High", "Medium", "Low", "None"))
#
#Selfing <- t %>%
#  group_by(Clusters,Autonomous_selfing_level) %>%
#  summarise (n = n()) %>%
#  mutate(Percentage = n / sum(n)*100)
#
##delete column to convert to wide
#Selfing <- Selfing[,-3]
##convert to wide
#Selfing <- spread(Selfing, Clusters, Percentage)
##Set call name os categories per trait
#colnames(Selfing)[1] <- "Category"
##Create new column
#Trait <- data.frame(Trait= c(rep("Selfing level", 2)))
##Replace Na's with zeros
#Selfing[is.na(Selfing)] <- 0
##Add column
#Selfing_1 <- cbind(Trait, Selfing)
##Bind rows
#bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span,Flower_morphology, Selfing_1)
#
#############################################################################################################################################
t$Nectar_presence_absence <- as.character(t$Nectar_presence_absence)
t$Nectar_presence_absence[t$Nectar_presence_absence =="yes"] <- "Yes"
t$Nectar_presence_absence[t$Nectar_presence_absence =="no"] <- "No"

Nectar <- t %>%
  group_by(Clusters,Nectar_presence_absence) %>%
  summarise (n = n()) %>%
  mutate(Percentage = n / sum(n)*100)

#delete column to convert to wide
Nectar <- Nectar[,-3]
#convert to wide
Nectar <- spread(Nectar, Clusters, Percentage)
#Set call name os categories per trait
colnames(Nectar)[1] <- "Category"
#Create new column
Trait <- data.frame(Trait= c(rep("Nectar", 2)))
#Replace Na's with zeros
Nectar[is.na(Nectar)] <- 0
#Add column
Nectar <- cbind(Trait, Nectar)
#Bind rows
data <- bind_rows(Breeding_system, Compatibility_system, Life_form, Life_span, Flower_symmetry, Nectar)

#remove first col
data_1 <- data[,-1]
colnames(data_1) <- c("Category", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")

#transpose dataframe
data_2 <- as.data.frame(t(data_1))

#rownames to colnames (1st one)
data_3 <- data_2 %>% row_to_names(row_number = 1)

#Convert all variables to numeric
data_4 <- sapply(data_3, function(x) as.numeric(as.character(x)))

#set back the missed rownames in the processing
rownames(data_4) <- rownames(data_3)

# Convert to dataframe
data_4 <- as.data.frame(data_4)

# Scale values and set rowname as column
data_5 <-data_4 %>% rownames_to_column( var = "group" )

data_6 <- melt(data_5)

cluster1 <- subset(data_6, group=="Cluster 1")

cluster1$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

#Try to achieve the desire output with plot_grid

c1 <- cluster1 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("sandybrown"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(axis.title.x = element_blank(), axis.text.x=element_blank(),plot.title = element_text(size = 18))+ ylab("")+
  ggtitle("A) Qualitative variables")

cluster2 <- subset(data_6, group=="Cluster 2")

cluster2$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

c2 <- cluster2 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("#7BB0A3"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(strip.background = element_blank(), strip.text = element_blank())+
  theme(axis.title.x = element_blank(), axis.text.x=element_blank())+ylab("")

cluster3 <- subset(data_6, group=="Cluster 3")

cluster3$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

c3 <- cluster3 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("lightsteelblue2"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(axis.title.x = element_blank(), axis.text.x=element_blank())+
  theme(strip.background = element_blank(), strip.text = element_blank())+ylab("Percentage")

cluster4 <- subset(data_6, group=="Cluster 4")

cluster4$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))

c4 <- cluster4 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("thistle"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+scale_y_continuous(expand = c(0,0)) +
  theme(axis.title.x = element_blank(), axis.text.x=element_blank())+
  theme(strip.background = element_blank(), strip.text = element_blank())+ylab("")

cluster5 <- subset(data_6, group=="Cluster 5")

cluster5$traits <- c(rep("Breeding system",3), rep("Compatibility",3), rep("Life form",3),
                     rep("Life span",2), rep("Flower symmetry",2), rep("Nectar",2))
c5 <- cluster5 %>% 
  ggplot(aes(x = variable,  y = value, fill = group))  +  
  geom_col(position = "dodge") +   scale_fill_manual(name = NULL,values=c("lightgoldenrod2"))+
  facet_grid(~traits, scales = "free_x", space = "free_x")+theme_bw()+
  scale_y_continuous(expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 45,vjust = 1, hjust=1))+
  theme(strip.background = element_blank(), strip.text = element_blank())+ylab("")+xlab("")

A <- c1 + c2 + c3 + c4 + c5 + plot_layout(ncol = 1)
B <- ggradar(final_df_3, group.colours=c("sandybrown", "#7BB0A3","lightsteelblue2", "thistle", "lightgoldenrod2"),
             legend.text.size=5, grid.label.size = 0,legend.position = "none",axis.label.size=3)+ 
  guides(shape = guide_legend(override.aes = list(size = 0.75)))+ ggtitle("B) Quantitative variables")+
  theme(plot.title = element_text(size = 18))

patch <- A | B

patch + plot_annotation(title = "Plant functional group composition",theme = theme(plot.title = element_text(size = 20,face="bold")))

```

\newpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=7, fig.width=13, fig.cap="**Figure S4.** Hierarchical clustering dendrogram with the branches coloured by the optimal number of clusters (5). The labels of the subgroup of species (N = 524) used in this study are coloured in black in order to show the evenness of the distribution of the species across clusters. The rest of species labels are omitted for visualization purposes (N = 982)."}

library(ggplot2)
library(dendextend)
library(dplyr)

e.clust_5 <- readRDS("../Data/RData/e.clust_5.rds")

dendro <- as.dendrogram(e.clust_5)
dendro.col <- dendro %>%
  set("branches_k_color", k = 5, value =   c("lightsteelblue2", "thistle", "sandybrown", "#7BB0A3", "lightgoldenrod2")) %>%
  set("branches_lwd", 0.7) %>%
  set("labels_colors", 
      value = c("darkslategray")) %>% 
  set("labels_cex", 0.075)

labels_colors(dendro.col) <- get_leaves_branches_col(dendro.col)

# We use "darkgrey" to color some species
final_d_1 <- read.csv("../Data/Csv/data_for_motifs_analysis_1.csv")

plant_species <- unique(final_d_1$Plant_species)

index_species_out <- which(names(labels_colors(dendro.col)) %in% plant_species)

'%ni%' <- Negate('%in%')

index_species_in <- which(names(labels_colors(dendro.col)) %ni% plant_species)

labels_colors(dendro.col)[index_species_out] <- "black"
labels_colors(dendro.col)[index_species_in] <- "white"

ggd1 <- as.ggdend(dendro.col)

ggd1$labels$angle <- 90
ggd1$labels$vjust <- 1
ggd1$labels$hjust <- 1

ggplot(ggd1,offset_labels=-0.05) + theme(panel.grid.major = element_blank(),
 axis.text = element_blank(),axis.title = element_blank()) +
  ggtitle("Plant functional groups")+
  theme(plot.title = element_text(hjust = 0.5, vjust = -2,size = 20,face="bold"))+
  annotate("text", x=140, y=1.65, label= "Self-incomp. perennials with large flowers ", size=2.5)+
  annotate("text", x=380, y=2.15, label= "Tall plants with small unisexual flowers", size=2.5) +
  annotate("text", x=600, y=3.85, label= "Short-lived selfers", size=2.5) +
  annotate("text", x=1400, y=2.05, label= "Short-lived outcrossers with \n long zygomorphic flowers", size=2.5) +annotate("text", x=950, y=2.05, label= "Small outcrossing perennials", size=2.5)

```

\elandscape