### 2022-02-08 Manuscript/Supplementary_Material/specificity_VS_position_frequencies.Rmd
Specificity plot ready for supplementary

---
title: ""
output:
  pdf_document: default
  html_document: default
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=12, fig.width=12}

library(ggplot2)
library(patchwork)

pollinator_means_reordered_s_1 <- read.csv("../../Data/Csv/pollinator_means_reordered_s_1.csv")

plant_means_s <- read.csv("../../Data/Csv/plant_means_s.csv")

p1 <- ggplot(pollinator_means_reordered_s_1,aes(x=s,y=mean))+
  geom_point(alpha=0.5, color="black")+
  geom_smooth(method = "lm",color="black")+
  facet_wrap(~Node_FG)+
  labs(x="Specificty",y="Average frequency per position") +
  theme_bw() + ggtitle("Floral visitors") +
  theme(plot.title = element_text(hjust = 0.5, size = 25),
        axis.text=element_text(size=12),
        axis.title=element_text(size=18),
        strip.text = element_text(size=20))

p2 <- ggplot(plant_means_s,aes(x=s,y=mean))+
  geom_point(alpha=0.5, color="black")+
  geom_smooth(method = "lm", color="black")+
  facet_wrap(~Node_FG)+
  labs(x="Specificty",y="Average frequency per position")+
  theme_bw() + ggtitle("Plants") +
  theme(plot.title = element_text(hjust = 0.5, size = 25),
        axis.text=element_text(size=12),
        axis.title=element_text(size=18),
        strip.text = element_text(size=20))

p1/p2

```