### 2022-03-08 Manuscript/Main_text/Title_page_Oikos.Rmd

---
title: "The non-random assembly of functional motifs in plant-pollinator networks worldwide"
output: pdf_document
---

__Jose B. Lanuza^1,2^__  | __Alfonso Allen-Perkins^2,3^__ | __Ignasi Bartomeus^2^__

1 School of Environmental and Rural Science, University of New England, Armidale, New South Wales 2350, Australia.

2 Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain.

3 Departamento de Ingeniería Eléctrica, Electrónica, Automática y Física Aplicada, ETSIDI, Universidad Politécnica de Madrid, 28040 Madrid, Spain.