### 2022-03-08 Manuscript/Supplementary_Material/Table_plant_species.Rmd

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE,fig.width=12,  warning=FALSE}
#Load libraries
library(tidyverse)
library(kableExtra)

#Read data
networks <- read_csv("../Data/Csv/data_for_motifs_analysis_1.csv")

#Subset by interaction greater than 0
d <- subset(networks, Interaction>0)
#Create dataframe
plant_species <- as.data.frame(unique(d$Plant_species))
#Divide dataset in two
x <- plant_species[1:168,]
y <- plant_species[169:336,]
z <- plant_species[337:503,]

n <- max(length(x), length(y),length(z))
length(x) <- n                      
length(y) <- n
length(z) <- n

#cbind with max length
d <- as.data.frame(cbind(x, y,z))
#Convert NA to blank spaces
d[is.na(d)] <- ""
#Create table
kbl(d,booktabs = T,col.names = c("", "List of plant species",""),longtable = T)%>%column_spec(1,italic = T)%>%column_spec(2,italic = T) %>%column_spec(3,italic = T)

```