### 2022-03-16 Manuscript/Supplementary_Material/Indirect_direct_plants.Rmd
Update on ms and supporting info

---
title: ""
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, fig.pos="t", warning=FALSE,out.width="75%",fig.align="center", fig.width=10,fig.height=12}

library(ggplot2)
library(scatterpie)
library(RColorBrewer)
library(tidyverse)

direct_indirect <- read_csv("../../Data/Data_processing/Motifs_connections/motif_interactions.csv")
direct_indirect <- direct_indirect %>% select(motif_id,position,direct_interactions, indirect_interactions,ind_same_type_over_directed_int)

plant_means_reordered <- read_csv("../../Data/Csv/plant_abs_freq_means.csv")

#Dplyr merge
joined_tibble <- left_join(direct_indirect, plant_means_reordered)

#1 Selfing herbs
one <- joined_tibble %>% filter(Node_FG=="1")
one$ind_same_type_over_directed_int <- round(one$ind_same_type_over_directed_int, digits = 1)
#Convert to factor
one$position <- as.factor(one$position)
one$indirect_interactions <- as.factor(one$indirect_interactions)
one_plot <- ggplot(one,aes(position,mean_natural_units))+
  geom_col(position = "dodge")+
  facet_grid(.~indirect_interactions,scale='free_x', space = "free_x") + ggtitle("Selfing herbs")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Frequency")

#2 Small outcrossing perennials
two <- joined_tibble %>% filter(Node_FG=="2")
two$ind_same_type_over_directed_int <- round(two$ind_same_type_over_directed_int, digits = 1)
#Convert to factor
two$position <- as.factor(two$position)
two$indirect_interactions <- as.factor(two$indirect_interactions)
two_plot <- ggplot(two,aes(position,mean_natural_units))+
  geom_col(position = "dodge")+
  facet_grid(.~indirect_interactions,scale='free_x', space = "free_x") + ggtitle("Small outcrossing perennials")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Frequency")

#3 Self-incomp. perennials with large flowers
three <- joined_tibble %>% filter(Node_FG=="3")
three$ind_same_type_over_directed_int <- round(three$ind_same_type_over_directed_int, digits = 1)
#Convert to factor
three$position <- as.factor(three$position)
three$indirect_interactions <- as.factor(three$indirect_interactions)
three_plot <- ggplot(three,aes(position,mean_natural_units))+
  geom_col(position = "dodge")+
  facet_grid(.~indirect_interactions,scale='free_x', space = "free_x") + ggtitle("Self-incomp. perennials with large flowers") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Frequency")

#4 Tall plants with small unisexual flowers
four <- joined_tibble %>% filter(Node_FG=="4")
four$ind_same_type_over_directed_int <- round(four$ind_same_type_over_directed_int, digits = 2)
#Convert to factor
four$position <- as.factor(four$position)
four$indirect_interactions <- as.factor(four$indirect_interactions)
four_plot <- ggplot(four,aes(position,mean_natural_units))+
  geom_col(position = "dodge")+
  facet_grid(.~indirect_interactions,scale='free_x', space = "free_x") + ggtitle("Tall plants with small unisexual flowers") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  ylab("Frequency")

#5 Short-lived outcrossers with long zygomorphic flowers
five <- joined_tibble %>% filter(Node_FG=="5")
five$ind_same_type_over_directed_int <- round(five$ind_same_type_over_directed_int, digits = 1)
#Convert to factor
five$position <- as.factor(five$position)
five$indirect_interactions <- as.factor(five$indirect_interactions)
five_plot <- ggplot(five,aes(position,mean_natural_units))+
  geom_col(position = "dodge")+
  facet_grid(.~indirect_interactions,scale='free_x', space = "free_x") + ggtitle("Short-lived outcrossers with long zygomorphic flowers")+
  ylab("Frequency")

library(patchwork)
one_plot / two_plot / three_plot / four_plot / five_plot + plot_annotation(
  title = 'Number of indirect interactions (0, 1, 2, 3)',
  theme = theme(plot.title = element_text(size = 18, face="bold"))) & theme_bw()

```